---
title: "Mutational analysis: differential expression + ORA"
author: "Camilo Posso"
date: "02/16/2022"
output: 
  html_document:
    code_folding: hide
    toc: true
editor_options: 
  chunk_output_type: inline
---

## Goal


The goal of this markdown is to make the ORA and heatmap plots coming from the
differential expression for selected mutations.


```{r include=FALSE}
library(clusterProfiler)
library(dplyr)
library(ggplot2)
library(kableExtra)

source("../util/synapseUtil.R")
source("../util/loading_data.R")
source("../util/differential_expression_helper.R")


#load("../Misc/load.combined.data 1-20-2022.RData")
load.combined.data()
metadata <- load.metadata()
samples <- global.data %>%
  pull(Barcode.ID) %>% unique()

total_mutations <- WES.data %>%
  pull(Gene) %>% unique()

## load clinical summary
clinical_syn <- "syn26642974"
syn <- synapseLogin()

clinical_summary <- read.table(syn$get(clinical_syn)$path)

IDH1_IDH2 <- "IDH1+IDH2"
KRAS_NRAS <- "KRAS+NRAS"

```



```{r}

## Mutations with at least 4 patients having said mutation
WES_reduced <- WES.data %>%
  filter(Barcode.ID %in% samples) %>% 
  group_by(Barcode.ID, Gene) %>%
  dplyr::slice(1) %>%
  ungroup(Barcode.ID, Gene)

mutation_count <- WES_reduced %>%
  group_by(Gene) %>%
  summarise(n = n())

mutations <-  mutation_count %>%
  filter(n > 3) %>%
  pull(Gene)

WES_samples <- WES_reduced %>%
  group_by(Gene) %>%
  mutate(n = n()) %>%
  arrange(-n) %>%
  pull(Barcode.ID) %>%
  unique()

WES_reduced <- WES_reduced %>%
  mutate(value = 1) %>%
  dplyr::select(Barcode.ID, Gene, value)


m.WES <- make.msnset(WES_reduced, "Gene", "Barcode.ID", "value", meta)  
mat <- exprs(m.WES)[, WES_samples]
mat[is.na(mat)] <- 0

```


Of the `r length(total_mutations)` mutations recorded in at least one sample, `r length(mutations)`
of them are present in at least 4 samples. Below we summarize the mutational data for these
`r length(mutations)` mutations using a heatmap. Note the columns represent samples, while the
light beige color denotes mutation.


```{r}
pheatmap(mat[mutations, ], color = c("#16182A", "#F5E9D9"), 
         legend = FALSE, show_colnames = FALSE)
```


## Differential Expression

Focusing on any one of the `r length(mutations)`, say \(\mu\), we can ask. Which phospho-sites 
are differentially expressed in \(\mu\) mutated samples vs non mutated samples? What if we use
global or RNA data? Below we collect this information into a table.


```{r include=FALSE}

m.RNA <- make.msnset(RNA.data, feature.col = "Gene",
                     value.col = "RNA counts", metadata = meta)
m.global <- make.msnset(global.data, feature.col = "Gene",
                        value.col = "LogRatio", metadata = meta)
m.phospho <- make.msnset(phospho.data, feature.col = "SiteID",
                         value.col = "LogRatio", metadata = meta)


datasets <- list(m.global, m.phospho, m.RNA)
names(datasets) <- c("global", "phospho", "RNA")

special_mutations <- WES.data %>%
  filter(Gene %in% c("IDH1", "IDH2", "KRAS", "NRAS")) %>%
  select(Gene, Barcode.ID) %>%
  mutate(mutation = case_when(grepl("IDH", Gene) ~ IDH1_IDH2,
                              TRUE ~ KRAS_NRAS)) %>%
  select(Barcode.ID, mutation) %>%
  dplyr::rename(Gene = mutation) %>%
  unique()

special_mutations_FLT3 <- metadata %>%
  filter(FLT3.ITD == "TRUE") %>%
  select(Barcode.ID) %>%
  mutate(Gene = "FLT3.ITD")

special_mutation_NPM1 <- clinical_summary %>%
  select(labId, NPM1Call) %>% 
  mutate(NPM1 = grepl("^Positive", NPM1Call),
         Barcode.ID = labId) %>%
  filter(NPM1) %>%
  select(Barcode.ID) %>%
  mutate(Gene = "NPM1_clinical")

special_mutations <- special_mutations %>%
  rbind(special_mutation_NPM1) %>%
  rbind(special_mutations_FLT3)

all_mutation_data <- WES.data %>%
  select(Barcode.ID, Gene) %>%
  rbind(special_mutations)


mutations <- c(mutations, unique(special_mutations$Gene))
## NPM1 mutations are made from clinical summary data used instead of the WES data
## so we remove NPM1 here
NPM1_WES_index <- which(mutations == "NPM1")
mutations <- mutations[-NPM1_WES_index]

differential_expression_results <- lapply(mutations, check_diff_exp_individual, 
                                          all_mutation_data, datasets)

summary_table <- lapply(differential_expression_results, function(xx){xx[[1]]}) %>% 
  do.call("rbind", .)

names(differential_expression_results) = sapply(differential_expression_results, function(xx){
  xx[[1]]$mutation[[1]]
})

```


```{r}
plot_df <- summary_table %>%
  select(-`Number.mutated`) %>%
  pivot_longer(-mutation) %>%
  mutate(name = sub("^.*in ", "", name)) %>%
  mutate(individual_bar = paste(mutation, name)) %>%
  arrange(-value) %>%
  mutate(mutation = factor(mutation, levels = unique(mutation)))

p <- ggplot(plot_df, aes(y = mutation, x = value, fill = name)) + 
  geom_bar(stat = 'identity') + xlab("Features differentially expressed")

ggsave(filename = "mutational_differential_expression_barplot_count.png", plot = p)

p

```



```{r}
library(DT)
DT::datatable(summary_table)
```


```{r include=FALSE}
### Run this to make all the heatmaps. A few will fail, due to the exactly one feature
### being selected to plot.** or failure to compute distances due to many NA
dataset_names <- names(datasets)

for (i in 1:length(differential_expression_results)){
  mutation = names(differential_expression_results)[[i]]
  diff_exp <- differential_expression_results[[i]][[2]]
  
  
  features <- lapply(1:3, function(j){
    diff_exp %>%
      filter(datatype == dataset_names[[j]]) %>%
      filter(adj.P.Val < 0.05) %>%
      head(20) %>%
      pull(feature)
  })
  
  for (j in 1:3){
    if (length(features[[j]]) > 0){
      try(plot_diffexp_heatmap(mutation, all_mutation_data, datasets[j], features[[j]], 
                               folder = "./plots/differential_expression_heatmaps/", 
                               suffix = dataset_names[[j]]))
    }
  }
}

```



```{r}
## Here we compute the enrichment results
library(org.Hs.eg.db)

selected_mutations <- c("IDH1", "IDH2", "TP53", "NRAS", "KRAS", "FLT3.ITD", "NPM1_clinical", 
                        IDH1_IDH2, KRAS_NRAS)


## need same order as datasets here
universes <- list(unique(global.data$Gene), unique(phospho.data$Gene), 
                  unique(RNA.data$Gene))
names(universes) <- names(datasets)

gsea_ora_results <- lapply(selected_mutations, function(mutation){
  print(mutation)
  gsea_all = gsea_global = gsea_rna <- data.frame(Description = c(), 
                                                  GeneRation = c(), 
                                                  BgRatio = c(), 
                                                  pvalue = c(), 
                                                  p.adjust = c(), 
                                                  Count = c())
  
  genes <- differential_expression_results[[mutation]][[2]] %>%
    filter(adj.P.Val < 0.05) %>%
    filter(datatype == "global") %>%
    pull(feature)
  
  if (length(genes) > 0){
    gsea_global <- enrichGO(genes, keyType = "SYMBOL", OrgDb = org.Hs.eg.db, ont = "BP",
                            universe = universes[[1]])@result %>%
      dplyr::select(Description, GeneRatio, BgRatio, pvalue, p.adjust, Count) %>%
      mutate(mutation = mutation, data_type = "global")
  }
  
  genes <- differential_expression_results[[mutation]][[2]] %>%
    filter(adj.P.Val < 0.05) %>%
    filter(datatype == "RNA") %>%
    pull(feature)

  if (length(genes) > 0){
    gsea_rna <- enrichGO(genes, keyType = "SYMBOL", OrgDb = org.Hs.eg.db, ont = "BP",
                         universe = universes[[3]])@result %>%
      dplyr::select(Description, GeneRatio, BgRatio, pvalue, p.adjust, Count) %>%
      mutate(mutation = mutation, data_type = "rna")
  }
  
  gsea_all <- rbind(gsea_all, gsea_global) %>%
    rbind(gsea_rna) %>%
    mutate(mutation = mutation)
  
  return(gsea_all)
}) 

gsea_ora_results_df <- do.call("rbind", gsea_ora_results)

```



```{r, ora plots for the selected mutations}
## Here we make the plots using the enrichment results
to_plot <- gsea_ora_results_df %>%
  filter(p.adjust < 0.05) 

categories_plot <- to_plot %>%
  dplyr::select(mutation, data_type) %>%
  unique()
rownames(categories_plot) <- 1:nrow(categories_plot)

lapply(1:nrow(categories_plot), function(i){
  plot_mutation <- categories_plot$mutation[[i]]
  plot_data_type <- categories_plot$data_type[[i]]
  plot_title <- paste(plot_mutation, plot_data_type)
  
  plot_path <- "./plots/over_representation_analysis/ORA_top_sig"
  plot_path <- paste(plot_path, plot_mutation, plot_data_type, sep = "_") %>%
    paste0(".png")
  plot_df <- to_plot %>%
    filter(mutation == plot_mutation) %>%
    filter(data_type == plot_data_type) %>%
    dplyr::select(Description, Count, p.adjust) %>%
    dplyr::rename(enrichment = Count,
                  pathway = Description,
                  adj_p_val = p.adjust) %>%
    head(20)
  
  png(file = plot_path, width = 1200, height = 800)
  plot_enrichment_result(plot_df, "Count", enrichment_title = plot_title)
  dev.off()
  
  "done"
})

```


