---
title: "Mutational analysis: GSEA and KSEA"
author: "Camilo Posso"
date: "02/16/2022"
output: 
  html_document:
    code_folding: hide
    toc: true
editor_options: 
  chunk_output_type: inline
---

## Goal


The goal of this markdown is to make the GSEA and KSEA plots comparing mutated to non mutated
samples for various selected mutations.


```{r include=FALSE}
library(clusterProfiler)
library(dplyr)
library(ggplot2)
library(kableExtra)

source("../util/synapseUtil.R")
source("../util/loading_data.R")
source("../util/differential_expression_helper.R")
source("../util/make_plots_util.R")


load("../Misc/load.combined.data 1-20-2022.RData")
# load.combined.data()
metadata <- load.metadata()
samples <- global.data %>%
  pull(Barcode.ID) %>% unique()

total_mutations <- WES.data %>%
  pull(Gene) %>% unique()

## load clinical summary
clinical_syn <- "syn26642974"
syn <- synapseLogin()

clinical_summary <- read.table(syn$get(clinical_syn)$path)

IDH1_IDH2 <- "IDH1+IDH2"
KRAS_NRAS <- "KRAS+NRAS"

```



```{r}

## Mutations with at least 4 patients having said mutation
WES_reduced <- WES.data %>%
  filter(Barcode.ID %in% samples) %>% 
  group_by(Barcode.ID, Gene) %>%
  dplyr::slice(1) %>%
  ungroup(Barcode.ID, Gene)

mutation_count <- WES_reduced %>%
  group_by(Gene) %>%
  summarise(n = n())

mutations <-  mutation_count %>%
  filter(n > 3) %>%
  pull(Gene)

WES_samples <- WES_reduced %>%
  group_by(Gene) %>%
  mutate(n = n()) %>%
  arrange(-n) %>%
  pull(Barcode.ID) %>%
  unique()

WES_reduced <- WES_reduced %>%
  mutate(value = 1) %>%
  dplyr::select(Barcode.ID, Gene, value)


m.WES <- make.msnset(WES_reduced, "Gene", "Barcode.ID", "value", meta)  
mat <- exprs(m.WES)[, WES_samples]
mat[is.na(mat)] <- 0

```


Of the `r length(total_mutations)` mutations recorded in at least one sample, `r length(mutations)`
of them are present in at least 4 samples. Below we summarize the mutational data for these
`r length(mutations)` mutations using a heatmap. Note the columns represent samples, while the
light beige color denotes mutation.


```{r}
pheatmap(mat[mutations, ], color = c("#16182A", "#F5E9D9"), 
         legend = FALSE, show_colnames = FALSE)
```



```{r include=FALSE}
## Here we run differential expression for convenience (to not rewrite code),
## in the end we are only interested in the log fold changes for the rank based GSEA
m.RNA <- make.msnset(RNA.data, feature.col = "Gene",
                     value.col = "RNA counts", metadata = meta)
m.global <- make.msnset(global.data, feature.col = "Gene",
                        value.col = "LogRatio", metadata = meta)
m.phospho <- make.msnset(phospho.data, feature.col = "SiteID",
                         value.col = "LogRatio", metadata = meta)


datasets <- list(m.global, m.phospho, m.RNA)
names(datasets) <- c("global", "phospho", "RNA")

special_mutations <- WES.data %>%
  filter(Gene %in% c("IDH1", "IDH2", "KRAS", "NRAS")) %>%
  select(Gene, Barcode.ID) %>%
  mutate(mutation = case_when(grepl("IDH", Gene) ~ IDH1_IDH2,
                              TRUE ~ KRAS_NRAS)) %>%
  select(Barcode.ID, mutation) %>%
  dplyr::rename(Gene = mutation) %>%
  unique()

special_mutations_FLT3 <- metadata %>%
  filter(FLT3.ITD == "TRUE") %>%
  select(Barcode.ID) %>%
  mutate(Gene = "FLT3.ITD")

special_mutation_NPM1 <- clinical_summary %>%
  select(labId, NPM1Call) %>% 
  mutate(NPM1 = grepl("^Positive", NPM1Call),
         Barcode.ID = labId) %>%
  filter(NPM1) %>%
  select(Barcode.ID) %>%
  mutate(Gene = "NPM1_clinical")

special_mutations <- special_mutations %>%
  rbind(special_mutation_NPM1) %>%
  rbind(special_mutations_FLT3)

all_mutation_data <- WES.data %>%
  select(Barcode.ID, Gene) %>%
  rbind(special_mutations)


mutations <- c(mutations, unique(special_mutations$Gene))
## NPM1 mutations are made from clinical summary data used instead of the WES data
## so we remove NPM1 here
NPM1_WES_index <- which(mutations == "NPM1")
mutations <- mutations[-NPM1_WES_index]

differential_expression_results <- lapply(mutations, check_diff_exp_individual, 
                                          all_mutation_data, datasets)

summary_table <- lapply(differential_expression_results, function(xx){xx[[1]]}) %>% 
  do.call("rbind", .)

names(differential_expression_results) = sapply(differential_expression_results, function(xx){
  xx[[1]]$mutation[[1]]
})

```


```{r}
library(clusterProfiler)
library(msigdbr)

t2g_hallmark <- msigdbr(species = "Homo sapiens", category = "H") %>% 
  dplyr::select(gs_name, gene_symbol)
## 50 unique terms
# unique(t2g_hallmark$gs_name)

t2g_oncogenic <- msigdbr(species = "Homo sapiens", category = "C6") %>% 
  dplyr::select(gs_name, gene_symbol)
## 189 unique terms
# unique(t2g_oncogenic$gs_name)

t2g_gobp <- msigdbr(species = "Homo sapiens", category = "C5", subcategory = "BP") %>% 
  dplyr::select(gs_name, gene_symbol) %>%
  mutate(gs_name = gsub("_", " ", gs_name))

```



```{r, rank based gsea}
library(org.Hs.eg.db)
library(KSEAapp)

selected_mutations <- c("IDH1", "IDH2", "TP53", "NRAS", "KRAS", "FLT3.ITD", "NPM1_clinical", 
                        IDH1_IDH2, KRAS_NRAS)

### KSEA pathway table
KSDB <- read.csv(system.file('PSP&NetworKIN_Kinase_Substrate_Dataset_July2016.csv',
                             package='amlresistancenetworks'),stringsAsFactors = FALSE)

gsea_rank_results <- lapply(selected_mutations, function(mutation){
  print(mutation)
  
  ## logFC in the differential expression table is calculated as follows:
  ## mean(mutated) - mean(non mutated)
  
  diff_exp <- differential_expression_results[[mutation]][[2]] %>%
    filter(datatype == "global") %>%
    arrange(-logFC)
  gene_list <- diff_exp$logFC
  names(gene_list) <- diff_exp$feature
  
  
  gsea_global_gobp <- GSEA(geneList = gene_list, eps = 1e-30, 
                           minGSSize = 10, pvalueCutoff = 0.05, TERM2GENE = t2g_gobp)@result %>%
    compress_enrichment(colname = "p.adjust", descending = FALSE) %>%
    dplyr::select(Description, setSize, p.adjust, NES) %>%
    dplyr::rename(pathway = Description, enrichment = NES, 
                  adj_p_val = p.adjust, set_size = setSize) %>%
    dplyr::mutate(data_type = "global",
                  mutation = mutation,
                  DB = "GO_BP")
  
  
  gsea_global_hallmark <- GSEA(geneList = gene_list, eps = 1e-30,
                               minGSSize = 10, pvalueCutoff = 0.05, TERM2GENE = t2g_hallmark)@result %>%
    dplyr::select(Description, setSize, p.adjust, NES) %>%
    dplyr::rename(pathway = Description, enrichment = NES, 
                  adj_p_val = p.adjust, set_size = setSize) %>%
    dplyr::mutate(data_type = "global",
                  mutation = mutation,
                  DB = "hallmark")
  
  
  gsea_global_oncogenic <- GSEA(geneList = gene_list, eps = 1e-30,
                                minGSSize = 10, pvalueCutoff = 0.05, TERM2GENE = t2g_oncogenic)@result %>%
    dplyr::select(Description, setSize, p.adjust, NES) %>%
    dplyr::rename(pathway = Description, enrichment = NES, 
                  adj_p_val = p.adjust, set_size = setSize) %>%
    dplyr::mutate(data_type = "global",
                  mutation = mutation,
                  DB = "oncogenic")
  
  gsea_global <- rbind(gsea_global_gobp, gsea_global_hallmark, gsea_global_oncogenic)
  

  diff_exp <- differential_expression_results[[mutation]][[2]] %>%
    filter(datatype == "RNA") %>%
    arrange(-logFC)
  gene_list <- diff_exp$logFC
  names(gene_list) <- diff_exp$feature
  
  gsea_rna_gobp <- GSEA(geneList = gene_list, eps = 1e-30, 
                        minGSSize = 10, pvalueCutoff = 0.05, TERM2GENE = t2g_gobp)@result %>%
    compress_enrichment(colname = "p.adjust", descending = FALSE) %>%
    dplyr::select(Description, setSize, p.adjust, NES) %>%
    dplyr::rename(pathway = Description, enrichment = NES, 
                  adj_p_val = p.adjust, set_size = setSize) %>%
    dplyr::mutate(data_type = "RNA",
                  mutation = mutation,
                  DB = "GO_BP")
  
  
  gsea_rna_hallmark <- GSEA(geneList = gene_list, eps = 1e-30,
                            minGSSize = 10, pvalueCutoff = 0.05, TERM2GENE = t2g_hallmark)@result %>%
    dplyr::select(Description, setSize, p.adjust, NES) %>%
    dplyr::rename(pathway = Description, enrichment = NES, 
                  adj_p_val = p.adjust, set_size = setSize) %>%
    dplyr::mutate(data_type = "RNA",
                  mutation = mutation,
                  DB = "hallmark")
  
  
  gsea_rna_oncogenic <- GSEA(geneList = gene_list, eps = 1e-30,
                             minGSSize = 10, pvalueCutoff = 0.05, TERM2GENE = t2g_oncogenic)@result %>%
    dplyr::select(Description, setSize, p.adjust, NES) %>%
    dplyr::rename(pathway = Description, enrichment = NES, 
                  adj_p_val = p.adjust, set_size = setSize) %>%
    dplyr::mutate(data_type = "RNA",
                  mutation = mutation,
                  DB = "oncogenic")
  
  gsea_rna <- rbind(gsea_rna_gobp, gsea_rna_hallmark, gsea_rna_oncogenic)
  
  
  
  diff_exp <- differential_expression_results[[mutation]][[2]] %>%
    filter(datatype == "phospho") %>%
    arrange(-logFC) %>%
    ## Making the necessary columns for KSEA
    mutate(Gene = sub("-.*$", "", feature),
           Residue.Both = sub("^.*-", "", feature)) %>%
    mutate(Residue.Both = gsub("[a-z]", ";", Residue.Both)) %>%
    mutate(Residue.Both = sub(";$", "", Residue.Both))
  
  PX <- data.frame(Protein = "NULL", Gene = diff_exp$Gene,
                   Peptide = "NULL", Residue.Both = diff_exp$Residue.Both,
                   p = "NULL", FC = diff_exp$logFC)
  
  ksea_phospho <- KSEA.Scores(KSDB, PX, NetworKIN = TRUE, NetworKIN.cutoff = 5) %>%
    dplyr::select(Kinase.Gene, m, FDR, z.score) %>%
    dplyr::rename(pathway = Kinase.Gene, enrichment = z.score,
                  adj_p_val = FDR, set_size = m) %>%
    dplyr::mutate(data_type = "phospho",
                  mutation = mutation,
                  DB = "KSDB") %>%
    filter(set_size > 3)
  
  gsea_all <- rbind(gsea_global, gsea_rna, ksea_phospho)
  
  return(gsea_all)
}) 

gsea_rank_results_df <- do.call("rbind", gsea_rank_results)


```




```{r}
to_plot <- gsea_rank_results_df %>%
  filter(adj_p_val < 0.05) 

categories_plot <- to_plot %>%
  dplyr::select(mutation, data_type, DB) %>%
  unique()
rownames(categories_plot) <- 1:nrow(categories_plot)

lapply(1:nrow(categories_plot), function(i){
  plot_mutation <- categories_plot$mutation[[i]]
  plot_data_type <- categories_plot$data_type[[i]]
  plot_DB <- categories_plot$DB[[i]]
  if (plot_data_type %in% c("global", "RNA")){
    x_label <- "NES"
    plot_title <- paste(plot_DB, plot_mutation, "GSEA in", plot_data_type)
  } else {
    x_label <- "z-score"
    plot_title <- paste(plot_DB, plot_mutation, "KSEA in", plot_data_type)
  }
  
  plot_path <- "./plots/gsea_rank_analysis/G(K)SEA"
  plot_path <- paste(plot_path, plot_DB, plot_mutation, plot_data_type, sep = "_") %>%
    paste0(".png")
  plot_df <- to_plot %>%
    filter(mutation == plot_mutation) %>%
    filter(data_type == plot_data_type) %>%
    filter(DB == plot_DB) %>%
    dplyr::select(pathway, enrichment, adj_p_val) %>%
    head(20)
  
  png(file = plot_path, width = 1200, height = 800)
  plot_enrichment_result(plot_df, x_label, enrichment_title = plot_title)
  dev.off()
  
  upload.plot(plot_path, parentId = "syn27220699")
  
  "done"
})


```


```{r}




```



