---
title: "Figure 1 Data Summary"
author: "Sara Gosline"
date: "03/30/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Get Data
This calls the standard data loading function that exists at `../beatAMLdata.R`. Once it is loaded it should not have to be reloaded. 

```{r setup, include=FALSE,message=FALSE}
library(amlresistancenetworks)
library(dplyr)
library(ggplot2)
source("../util/synapseUtil.R")
source("../util/loading_data.R")

load.combined.data()

```

## Data Summary

The goal fo this markdown is to summarize the BeatAML data that we have so far. Patient summary, summary of AUC value, and general proteomics/protein data. Here we have the data for

```{r histogram and heatmaps drug level, message=FALSE}
## This defines an analogous auc.dat table to the Pilot manuscript.
## The function plotAllAUCs needs the dat.summ table. 
## It makes dat.summ normally from passed data,
## But it's easier to pass it here instead. This needs a very 
## small modification of plotAllAUCs. Other ways of getting around 
## this require changing the original function too much. 
## The current change is minimal and should not break any previous code.
auc.dat <- functional.data %>%
  dplyr::rename(`AML sample` = Barcode.ID,
                Condition = Inhibitor)
dat.summ <- sample.summary %>%
  dplyr::rename(`AML sample` = Barcode.ID,
                mutations = WES,
                proteins = Global,
                phosphoSite = Phospho)

res <- auc.dat %>%
  plotAllAUCs(pat.data,drug.metric='AUC',drug.column='Condition', 
              dat.summ = dat.summ)

full.drg<-res%>%
  tibble::rownames_to_column("AML sample")%>%
  left_join(auc.dat)%>%
  select(`AML sample`,Drug='Condition',proteins,RNA,mutations)%>%
  distinct()%>%
  group_by(Drug)%>%
  summarize(`Patients with proteomics`=length(which(proteins==TRUE)),
            `Patients with mRNA`=length(which(RNA==TRUE)),
            `Patients with Mutations`=length(which(mutations==TRUE)))
  
knitr::kable(full.drg)
write.csv(full.drg,'table1.csv')
write.csv(res,'supp_table1.csv')
  
```

Since many of the patients are incomplete, we save a second version of the data that has only those 'complete' datasets.

```{r full patients}
res<-plotAllPatients(auc.dat,pat.data,pat.phos)

## full pats is the set of samples with all types of data (here total of 4, same as now).
full.pats<<-res%>%
  rowwise()%>%
  mutate(fullData=(as.character(RNA)=="TRUE" && as.character(proteins)=="TRUE"
                   && as.character(mutations)=="TRUE" && as.character(phosphoSites)=="TRUE"))%>%
  subset(fullData==TRUE)%>%
  dplyr::select(`AML sample`)
  
  
red.auc<-auc.dat%>%subset(`AML sample`%in%full.pats$`AML sample`)
red.dat<-pat.data%>%subset(`AML sample`%in%full.pats$`AML sample`)
red.phos<-pat.phos%>%subset(Sample%in%full.pats$`AML sample`)

knitr::kable(res[full.pats$`AML sample`,])
```

## Drug Family summary

We can also collate information by drug family to increase power. 


```{r drug family, message=FALSE}
res<-auc.dat.fam%>%select(-c('FLT3-ITD','FLT3-MUT','NPM1'))%>%
  plotAllAUCs(.,pat.data,drug.metric='AUC',drug.column='family')

full.drg<-res%>%
  tibble::rownames_to_column("AML sample")%>%
  left_join(auc.dat.fam)%>%
  select(`AML sample`,Drug='family',proteins,RNA,mutations)%>%
  distinct()%>%
  group_by(Drug)%>%
  summarize(`Patients with proteomics`=length(which(proteins==TRUE)),`Patients with mRNA`=length(which(RNA==TRUE)),`Patients with
            Mutations`=length(which(mutations==TRUE)))

knitr::kable(full.drg)
write.csv(full.drg,'fam_table1.csv')
write.csv(res,'fam_supp_table1.csv')

```


## Protein/mRNA/phosphosite correlation values
how well do these samples correlate across all samples?


### mRNA/protein


```{r, prot cors,warning=FALSE,message=FALSE}

library(dplyr)
library(ggplot2)

## These mRNA levels are not normalized, ours are. Ask about whether we should change anything here.
mp<-combined%>%
  dplyr::select('AML sample','proteinLevels','mRNALevels','Gene')%>%
  distinct()%>%
  subset(mRNALevels>1)%>%
  group_by(`AML sample`)%>%
  summarize(corval=cor(proteinLevels,mRNALevels,method='spearman',use='pairwise.complete.obs'))

DT::datatable(mp)

mpp<-combined%>%
  dplyr::select('AML sample','proteinLevels','mRNALevels','Gene','binaryMutations')%>%
  distinct()%>%
  subset(mRNALevels>1)%>%
  subset(proteinLevels!=0)%>%
  mutate(mRNALevels=log10(0.01+mRNALevels))%>%
  ggplot(aes(x=mRNALevels,y=proteinLevels,col=`AML sample`,
             shape=as.factor(binaryMutations)))+geom_point()+scale_color_viridis_d()+ggtitle('mRNA/Protein correlation')


```


No real correlation here

### protein/phosphosite

```{r prot phospho cor, warning=FALSE}
library(dplyr)
mp<-combined%>%
  dplyr::select('AML sample','proteinLevels','Phosphosite','Gene')%>%
  distinct()%>%
  group_by(`AML sample`)%>%
  summarize(corval=cor(proteinLevels,Phosphosite,method='spearman',use='pairwise.complete.obs'))

DT::datatable(mp)

mpp<-combined%>%
  dplyr::select('AML sample','proteinLevels','Phosphosite','Gene','binaryMutations')%>%
  distinct()%>%
 # subset(mRNALevels>1)%>%
  subset(proteinLevels!=0)%>%
#  mutate(mRNALevels=log10(0.01+mRNALevels))%>%
  ggplot(aes(x=Phosphosite,y=proteinLevels,col=`AML sample`,shape=as.factor(binaryMutations)))+geom_point()+scale_color_viridis_d()+ggtitle('Protein/phosphosite correlation')


```
### gene/mRNA

### gene/protein

## FLT3 pathway visualization

Let's see how each data type is represented across the FLT3 pathway

```{r flt3pathway, message=F,warning=F}
library(ggplot2)
flt3prots<-c("FLT3","FYN","GAB2","GRB2","HRAS",'HCK','KRAS','NRAS','PIK3R1','PTPN11','SOS1')


quiz.auc<-auc.dat%>%
#  subset(`AML sample`%in%flt3.pats$`AML sample`)%>%
  dplyr::select(Condition,AUC,'AML sample')%>%
  subset(Condition=='Quizartinib (AC220)')

flt3.pats<-combined%>%
  subset(Gene%in%flt3prots)%>%
  inner_join(quiz.auc)
  #subset(`AML sample`%in%auc.dat$`AML sample`)

genecors<-flt3.pats%>%
  group_by(Gene)%>%
  summarize(tcor=cor(mRNALevels,proteinLevels,use='pairwise.complete.obs',method='spearman'))

phoscors<-flt3.pats%>%
  
  group_by(Gene)%>%
  summarize(pcor=cor(Phosphosite,proteinLevels,use='pairwise.complete.obs',method='spearman'))

genecors<-genecors%>%left_join(phoscors)
write.csv(genecors,file='flt3CorVals.csv')

DT::datatable(genecors)

cor.genes<-genecors$Gene#$subset(genecors,abs(tcor)>0.25)$Gene

dotplots<-lapply(cor.genes,function(x)
  subset(flt3.pats,Gene==x)%>%
    mutate(mRNALevels=log10(0.01+mRNALevels))%>%
    mutate(mutated=as.factor(binaryMutations))%>%
    ggplot(aes(x=proteinLevels,y=mRNALevels,shape=mutated,color=AUC))+geom_point()+ggtitle(x)
)

p<-cowplot::plot_grid(plotlist=dotplots, ncol=2)
p

ggsave('flt3_correlatedDotPlots.pdf',p,width=12,height=24)

annotes<-quiz.auc%>%
  pivot_wider(values_from='AUC',names_from='Condition',values_fn=list(AUC=mean))%>%
  tibble::column_to_rownames('AML sample')

pat.order<-arrange(annotes, `Quizartinib (AC220)`)%>%rownames()
##get prot
prot.mat<-flt3.pats%>%
  dplyr::select(Gene,`AML sample`,proteinLevels)%>%
  pivot_wider(values_from = proteinLevels,names_from=`AML sample`,values_fill=0.0,
              values_fn=list(proteinLevels=mean))%>%
  mutate(Gene=paste(Gene,'prot'))%>%
  tibble::column_to_rownames('Gene')

##get phospho
ph.mat<-pat.phos%>%
    subset(Sample%in%auc.dat$`AML sample`)%>%
  subset(Gene%in%flt3prots)%>%
  dplyr::select(site,Sample,LogFoldChange)%>%
  pivot_wider(values_from='LogFoldChange',names_from='Sample',values_fill=0.0,
              values_fn=list(LogFoldChange=mean))%>%
  tibble::column_to_rownames('site')

## get mrna
t.mat<-flt3.pats%>%
  dplyr::select(Gene,`AML sample`,mRNALevels)%>%
  pivot_wider(values_from = mRNALevels,names_from=`AML sample`,values_fill=0.0,
              values_fn=list(mRNALevels=mean))%>%
  mutate(Gene=paste(Gene,'mRNA'))%>%
  tibble::column_to_rownames('Gene')
t.mat<-log10(0.01+t.mat)
missing=which(apply(t.mat,2,var)==0)

t.mat<-t.mat[,-missing]

##get VAF
v.mat<-combined%>%
  subset(Gene%in%c("FLT3","NRAS",'KRAS'))%>%
  dplyr::select(Gene,`AML sample`,binaryMutations)%>%
  pivot_wider(values_from = binaryMutations,names_from=`AML sample`,values_fill=0.0,
              values_fn=list(binaryMutations=max))%>%
  mutate(Gene=paste(Gene,'Gene'))%>%
  tibble::column_to_rownames('Gene')


flt3.vars<-t(v.mat) ##asave this for later so twe can add as ribbon to next figure
pat.order<-intersect(pat.order,colnames(t.mat))
fullmat<-rbind(prot.mat[,pat.order],t.mat[,pat.order],ph.mat[,pat.order])

fannotes<-cbind(Quizartinib=annotes[colnames(fullmat),],t(v.mat)[colnames(fullmat),])
#allz<-which(apply(fannotes,2,var)==0)
#fannotes<-fannotes[,-allz]

pheatmap(fullmat,cluster_rows = FALSE)

pheatmap(fullmat[sort(rownames(fullmat)),],
         cluster_rows = FALSE,cellheight=12,
         clustering_distance_cols = 'correlation',
         cluster_cols = TRUE, file='combined_heatmap.pdf',
         annotation_col = as.data.frame(fannotes) ,height=15)

pheatmap(as.matrix(prot.mat[,rownames(fannotes)]),
         cluster_rows = TRUE,cellheight=12,
        # clustering_distance_cols = 'correlation',
         cluster_cols = FALSE, file='prot_heatmap.pdf',
         annotation_col = as.data.frame(fannotes))
pheatmap(ph.mat[,rownames(fannotes)],
         cluster_rows = TRUE,cellheight=12,
         #clustering_distance_cols = 'correlation',
         cluster_cols = FALSE, file='phos_heatmap.pdf',
         annotation_col = as.data.frame(fannotes))
pheatmap(t.mat[,rownames(fannotes)],
         cluster_rows = TRUE,cellheight=12,
         #clustering_distance_cols = 'correlation',
         cluster_cols = FALSE, file='rna_heatmap.pdf',
         annotation_col = as.data.frame(fannotes))


dotplots<-lapply(subset(genecors,!is.na(pcor))$Gene,function(x)
  subset(flt3.pats,Gene==x)%>%
 #   mutate(mRNALevels=log10(0.01+mRNALevels))%>%
    mutate(mutated=as.factor(binaryMutations))%>%
    ggplot(aes(x=proteinLevels,y=Phosphosite,color=AUC,shape=site))+geom_point()+ggtitle(x)
)

p<-cowplot::plot_grid(plotlist=dotplots, ncol=2)
p
ggsave('flt3_phosphoDotPlots.pdf',p,width=12,height=12)

```

This shows that proteomic data is important to understanding the impact of targeted therapies. 


## ERK pathway visualization

Let's see how each data type is represented across the ERK1/2 pathway

```{r erkpathway, message=F,warning=F}
library(ggplot2)
erk1prots<-c("CDK1","MAP2K1","IL6","IL6R","IL6ST",'JAK1','JAK2','MAPK3','TYK2','PTPN11')


quiz.auc<-auc.dat%>%
#  subset(`AML sample`%in%erk1.pats$`AML sample`)%>%
  dplyr::select(Condition,AUC,'AML sample')%>%
  subset(Condition=='Trametinib (GSK1120212)')

erk1.pats<-combined%>%
  subset(Gene%in%erk1prots)%>%
  inner_join(quiz.auc)
  #subset(`AML sample`%in%auc.dat$`AML sample`)

genecors<-erk1.pats%>%
  group_by(Gene)%>%
  summarize(tcor=cor(mRNALevels,proteinLevels,use='pairwise.complete.obs',method='spearman'))


phoscors<-erk1.pats%>%
  group_by(Gene)%>%
  summarize(pcor=cor(Phosphosite,proteinLevels,use='pairwise.complete.obs',method='spearman'))

genecors<-genecors%>%left_join(phoscors)

DT::datatable(genecors)
write.csv(genecors,file='erk1CorVals.csv')

cor.genes<-genecors$Gene#subset(genecors,abs(tcor)>0.1)$Gene

dotplots<-lapply(cor.genes,function(x)
  subset(erk1.pats,Gene==x)%>%
    mutate(mRNALevels=log10(0.01+mRNALevels))%>%
    mutate(mutated=as.factor(binaryMutations))%>%
    ggplot(aes(x=proteinLevels,y=mRNALevels,shape=mutated,color=AUC))+geom_point()+ggtitle(x)
)

p<-cowplot::plot_grid(plotlist=dotplots)
p

ggsave('erk1_correlatedDotPlots.pdf',p,width=12,height=12)

annotes<-quiz.auc%>%
  pivot_wider(values_from='AUC',names_from='Condition',values_fn=list(AUC=mean))%>%
  tibble::column_to_rownames('AML sample')

pat.order<-arrange(annotes, `Trametinib (GSK1120212)`)%>%rownames()
##get prot
prot.mat<-erk1.pats%>%
  dplyr::select(Gene,`AML sample`,proteinLevels)%>%
  pivot_wider(values_from = proteinLevels,names_from=`AML sample`,values_fill=0.0,
              values_fn=list(proteinLevels=mean))%>%
  mutate(Gene=paste(Gene,'prot'))%>%
  tibble::column_to_rownames('Gene')

##get phospho
ph.mat<-pat.phos%>%
    subset(Sample%in%auc.dat$`AML sample`)%>%
  subset(Gene%in%erk1prots)%>%
  dplyr::select(site,Sample,LogFoldChange)%>%
  pivot_wider(values_from='LogFoldChange',names_from='Sample',values_fill=0.0,
              values_fn=list(LogFoldChange=mean))%>%
  tibble::column_to_rownames('site')
## get mrna
t.mat<-erk1.pats%>%
  dplyr::select(Gene,`AML sample`,mRNALevels)%>%
  pivot_wider(values_from = mRNALevels,names_from=`AML sample`,values_fill=0.0,
              values_fn=list(mRNALevels=mean))%>%
  mutate(Gene=paste(Gene,'mRNA'))%>%
  tibble::column_to_rownames('Gene')
t.mat<-log10(0.01+t.mat)
missing=which(apply(t.mat,2,var)==0)

t.mat<-t.mat[,-missing]

##get VAF
v.mat<-erk1.pats%>%
  dplyr::select(Gene,`AML sample`,binaryMutations)%>%
  pivot_wider(values_from = binaryMutations,names_from=`AML sample`,values_fill=0.0,
              values_fn=list(binaryMutations=max))%>%
  mutate(Gene=paste(Gene,'Gene'))%>%
  tibble::column_to_rownames('Gene')


pat.order<-intersect(pat.order,colnames(t.mat))
fullmat<-rbind(prot.mat[,pat.order],t.mat[,pat.order],ph.mat[,pat.order])

fannotes<-data.frame(Trametinib=annotes[colnames(fullmat),1],flt3.vars[colnames(fullmat),])


#rownames(tannotes)<-colnames(fullmat)
#,t(v.mat)[colnames(fullmat),])
#allz<-which(apply(fannotes,2,var)==0)
#fannotes<-fannotes[,-allz]

pheatmap(fullmat,cluster_rows = FALSE)

pheatmap(fullmat[sort(rownames(fullmat)),],
         cluster_rows = FALSE,cellheight=12,
         clustering_distance_cols = 'correlation',
         cluster_cols = TRUE, file='erk1_combined_heatmap.pdf',
         annotation_col = as.data.frame(fannotes) ,height=15)

pheatmap(as.matrix(prot.mat[,rownames(fannotes)]),
         cluster_rows = TRUE,cellheight=12,
        # clustering_distance_cols = 'correlation',
         cluster_cols = FALSE, file='erk1_prot_heatmap.pdf',
         annotation_col = as.data.frame(fannotes))

pheatmap(ph.mat[,rownames(fannotes)],
         cluster_rows = TRUE,cellheight=12,
         #clustering_distance_cols = 'correlation',
         cluster_cols = FALSE, file='erk1_phos_heatmap.pdf',
         annotation_col = as.data.frame(fannotes))
pheatmap(t.mat[,rownames(fannotes)],
         cluster_rows = TRUE,cellheight=12,
         #clustering_distance_cols = 'correlation',
         cluster_cols = FALSE, file='erk1_rna_heatmap.pdf',
         annotation_col = as.data.frame(fannotes))


dotplots<-lapply(subset(genecors,!is.na(pcor))$Gene,function(x)
  subset(erk1.pats,Gene==x)%>%
 #   mutate(mRNALevels=log10(0.01+mRNALevels))%>%
    mutate(mutated=as.factor(binaryMutations))%>%
    ggplot(aes(x=proteinLevels,y=Phosphosite,color=AUC,shape=site))+geom_point()+ggtitle(x)
)

p<-cowplot::plot_grid(plotlist=dotplots, ncol=2)
p
ggsave('erk1_phosphoDotPlots.pdf',p,width=12,height=12)

```


These figures suggest that mRNA and protein data are not highly correlated with drug response, as we'd expect, nor is phosphosite and protein data.

```{r phospho plots}






```


