---
title: "Cluster 5 binomial signature enrichment"
author: "Camilo Posso"
date: "04/04/2022"
output: 
  html_document:
    code_folding: hide
    toc: true
editor_options: 
  chunk_output_type: inline
---

## Goal




```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(glmnet)
library(clusterProfiler)
library(msigdbr)
library(RColorBrewer)

source("../../../util/synapseUtil.R")
source("../../../util/loading_data.R")
source("../../../util/mutational_analysis_helper.R")


load("../../../Misc/load.combined.data 3-09-2022.RData")
# load.combined.data()
global_data <- global.data
RNA_data <- RNA.data %>%
  dplyr::rename(LogRatio = `RNA counts`)


```



```{r}

cluster_assignments <- read.table("../Data/sample_to_cluster_assignments.txt", sep = "\t")
diff_markers_k5 <- read.table("../../../Misc/Extra NMF/whoknows/differential_markers_k5.txt", sep = "\t")

summary.syn <- "syn26642974"
summary.table <- read.table("../Data/PNNL_clinical_summary_12_08_2021_updated_OS_4patients_02_28_2022.txt") %>%
  as.data.frame()

rownames(summary.table) <- summary.table$labId
summary.table <- summary.table[rownames(meta), ]
meta$vitalStatus <- summary.table$vitalStatus
meta <- meta %>%
  dplyr::mutate(vitalStatus = case_when(vitalStatus == "Dead" ~ 2,
                                 TRUE ~ 1))

cluster_samples <- cluster_assignments$Barcode.ID
global_data <- global.data %>%
  dplyr::filter(Barcode.ID %in% cluster_samples)
global_data_test <- global.data %>%
  dplyr::filter(!(Barcode.ID %in% cluster_samples))


mat_df <- pivot_wider(global_data, names_from = Barcode.ID, 
                   values_from = LogRatio) %>%
  as.data.frame() 
rownames(mat_df) <- mat_df$Gene

data_mat_train_all <- mat_df[, -1] %>% as.matrix()
mat_test_df <- pivot_wider(global_data_test, names_from = Barcode.ID, 
                           values_from = LogRatio) %>%
  as.data.frame() 
rownames(mat_test_df) <- mat_df$Gene
data_mat_test_all <- mat_test_df[, -1] %>% as.matrix()

complete_cases <- complete.cases(data_mat_train_all) & complete.cases(data_mat_test_all)
data_mat_train_all <- data_mat_train_all[complete_cases, ]
data_mat_test_all <- data_mat_test_all[complete_cases, ]

signatures <- read.table("./elastic_net_binomial_models/enet_binomial_all_signatures.txt", sep = "\t")
prot_universe <- rownames(data_mat_train_all)

```



Setting up helper function for finding signature enrichment.


```{r}

get_signature_helper <- function(chosen_cluster, chosen_alpha, diff_markers, t2g){
  sigs <- signatures %>% 
    dplyr::filter(diff_marker == diff_markers) %>%
    dplyr::filter(alpha == chosen_alpha) %>%
    dplyr::filter(Cluster == chosen_cluster) %>%
    pull(Gene)
    
  set.seed(117)
  enrichment <- enricher(sigs, 
                         TERM2GENE = t2g, 
                         universe = prot_universe,
                         pvalueCutoff = 1,
                         minGSSize = 10)@result %>%
    dplyr::select(Description, GeneRatio, BgRatio, pvalue, p.adjust, geneID) %>%
    dplyr::rename(core_enrichment = geneID) %>%
    compress_enrichment(colname = "p.adjust") %>%
    dplyr::select(Description, GeneRatio, BgRatio, pvalue, p.adjust) %>%
    dplyr::mutate(alpha = chosen_alpha, cluster = chosen_cluster, diff_markers = diff_markers)
  
  return(enrichment)
}


```


Running for all alpha and all clusters.


## GO BP

```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C5", subcategory = "BP") %>% 
  dplyr::select(gs_name, gene_symbol) %>%
  mutate(gs_name = gsub("_", " ", gs_name))

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste(chosen_alpha, chosen_cluster))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "GO BP")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_gobp.txt", 
            sep = "\t", quote = F)


```





## GO cellular component


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C5") %>%
  dplyr::filter(gs_subcat == "GO:CC") %>%
  dplyr::select(gs_name, gene_symbol) %>%
  mutate(gs_name = gsub("_", " ", gs_name))

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  print(paste("cluster =", chosen_cluster))
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste("alpha =", chosen_alpha))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "GO CC")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_gocc.txt", 
            sep = "\t", quote = F)


```




## GO Molecular function


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C5") %>%
  dplyr::filter(gs_subcat == "GO:MF") %>%
  dplyr::select(gs_name, gene_symbol) %>%
  mutate(gs_name = gsub("_", " ", gs_name))

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  print(paste("cluster =", chosen_cluster))
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste("alpha =", chosen_alpha))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "GO MF")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_gomf.txt", 
            sep = "\t", quote = F)


```




## GO Human phenotype


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C5") %>%
  dplyr::filter(gs_subcat == "HPO") %>%
  dplyr::select(gs_name, gene_symbol) %>%
  mutate(gs_name = gsub("_", " ", gs_name))

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  print(paste("cluster =", chosen_cluster))
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste("alpha =", chosen_alpha))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "GO HP")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_gohp.txt", 
            sep = "\t", quote = F)


```






## Hallmarks


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "H") %>% 
  dplyr::select(gs_name, gene_symbol)

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste(chosen_alpha, chosen_cluster))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "Hallmark")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_hallmark.txt", 
            sep = "\t", quote = F)


```


## Oncogenic


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C6") %>% 
  dplyr::select(gs_name, gene_symbol)

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste(chosen_alpha, chosen_cluster))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "Oncogenic")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_oncogenic.txt", 
            sep = "\t", quote = F)


```



## Curated: Chemical and genetic perturbations


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C2") %>% 
  dplyr::filter(gs_subcat == "CGP") %>%
  dplyr::select(gs_name, gene_symbol) 

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste(chosen_alpha, chosen_cluster))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "curated cgp")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_curated_cgp.txt", 
            sep = "\t", quote = F)


```




## Curated: REACTOME


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C2") %>% 
  dplyr::filter(gs_subcat == "CP:REACTOME") %>%
  dplyr::select(gs_name, gene_symbol) 

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste(chosen_alpha, chosen_cluster))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "curated reactome")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_curated_reactome.txt", 
            sep = "\t", quote = F)


```




## Regulatory


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C3") %>% 
  dplyr::select(gs_name, gene_symbol) 

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste(chosen_alpha, chosen_cluster))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "regulatory")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_regulatory.txt", 
            sep = "\t", quote = F)


```



## Computational Cancer Gene Neighborhoods


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C4") %>% 
  dplyr::filter(gs_subcat == "CGN") %>%
  dplyr::select(gs_description, gene_symbol, gs_name) %>%
  dplyr::mutate(gs_name = paste0(gs_name, "_", gs_description)) %>%
  dplyr::select(gs_name, gene_symbol)

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste(chosen_alpha, chosen_cluster))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "computational CGN")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_computational_cgn.txt", 
            sep = "\t", quote = F)


```



## Computational Cancer Modules


```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C4") %>% 
  dplyr::filter(gs_subcat == "CM") %>%
  dplyr::select(gs_description, gene_symbol, gs_name) %>%
  dplyr::mutate(gs_name = paste0(gs_name, "_", gs_description)) %>%
  dplyr::select(gs_name, gene_symbol)

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste(chosen_alpha, chosen_cluster))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "computational CM")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_computational_cm.txt", 
            sep = "\t", quote = F)


```



## Immunologic



```{r include=FALSE}
t2g <- msigdbr(species = "Homo sapiens", category = "C7") %>% 
  dplyr::select(gs_name, gene_symbol) 

all_signature_enrichment <- lapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
  lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), function(chosen_alpha){
    print(paste(chosen_alpha, chosen_cluster))
    sig_markers <- get_signature_helper(chosen_cluster, chosen_alpha, 
                                        diff_markers = TRUE, t2g = t2g)
    sig_all <- get_signature_helper(chosen_cluster, chosen_alpha,
                                    diff_markers = FALSE, t2g = t2g)
    
    out <- rbind(sig_markers, sig_all)
    return(out)
  }) %>% do.call("rbind", .)
}) %>% do.call("rbind", .) %>%
  dplyr::mutate(DB = "immunologic")

write.table(all_signature_enrichment, "elastic_net_binomial_models/protein_signatures_immunologic.txt", 
            sep = "\t", quote = F)


```





## Quick summary to decide on alpha



```{r}
sig_gobp <- read.table("elastic_net_binomial_models/protein_signatures_gobp.txt", sep = "\t")
sig_gocc <- read.table("elastic_net_binomial_models/protein_signatures_gocc.txt", sep = "\t")
sig_gomf <- read.table("elastic_net_binomial_models/protein_signatures_gomf.txt", sep = "\t")
sig_gohp <- read.table("elastic_net_binomial_models/protein_signatures_gohp.txt", sep = "\t")
sig_hallmarks <- read.table("elastic_net_binomial_models/protein_signatures_hallmark.txt", sep = "\t")
sig_oncogenic <- read.table("elastic_net_binomial_models/protein_signatures_oncogenic.txt", sep = "\t")
sig_curated_cgp <- read.table("elastic_net_binomial_models/protein_signatures_curated_cgp.txt", sep = "\t")
sig_curated_reactome <- read.table("elastic_net_binomial_models/protein_signatures_curated_reactome.txt", sep = "\t")
sig_regulatory <- read.table("elastic_net_binomial_models/protein_signatures_regulatory.txt", sep = "\t")
sig_computational_cgn <- read.table("elastic_net_binomial_models/protein_signatures_computational_cgn.txt", sep = "\t")
sig_computational_cm <- read.table("elastic_net_binomial_models/protein_signatures_computational_cm.txt", sep = "\t")
sig_immunologic <- read.table("elastic_net_binomial_models/protein_signatures_immunologic.txt", sep = "\t")

all_enrichment <- list("gobp" = sig_gobp,
                       "gocc" = sig_gocc,
                       "gomf" = sig_gomf,
                       "gohp" = sig_gohp,
                       "hallmarks" = sig_hallmarks,
                       "oncogenic" = sig_oncogenic,
                       "curated cgp" = sig_curated_cgp,
                       "curated reactome" = sig_curated_reactome,
                       "regulatory" = sig_regulatory,
                       "computational cgn" = sig_computational_cgn,
                       "computational cm" = sig_computational_cm,
                       "immunologic" = sig_immunologic)



```

Plotting the number of significant pathways for every alpha + cluster combination.

```{r include=FALSE}

lapply(names(all_enrichment), function(result_name){
  print(result_name)
  result <- all_enrichment[[result_name]]
  plot_title <- paste0("Total significant, database = ",
                       result_name, " diff markers FALSE")
  plot_path <- paste0("./elastic_net_binomial_models/total_significant_database_", 
                      result_name,
                      ".png")
  
  counts <- result %>%
    dplyr::filter(p.adjust < 0.05) %>%
    dplyr::filter(diff_markers == FALSE) %>%
    dplyr::group_by(cluster, alpha) %>%
    summarize(total_significant = n()) %>%
    dplyr::ungroup(cluster, alpha)
  
  if(length(counts$cluster %>% unique()) > 1 & length(counts$alpha %>% unique()) > 1){
    counts <- pivot_wider(counts, names_from = alpha, values_from = total_significant, values_fill = 0)
    counts_mat <- counts[, -1] %>% as.matrix()
    rownames(counts_mat) <- paste("Cluser", counts$cluster)
    counts_mat <- counts_mat[, sort(colnames(counts_mat))]
    
    png(plot_path, width = 800, height = 800)
    pheatmap(counts_mat, cluster_rows = F, cluster_cols = F, main = plot_title, display_numbers = TRUE)
    dev.off()
  }
  
})


lapply(names(all_enrichment), function(result_name){
  print(result_name)
  result <- all_enrichment[[result_name]]
  plot_title <- paste0("Total significant, database = ",
                       result_name, " diff markers TRUE")
  plot_path <- paste0("./elastic_net_binomial_models/total_significant_database_", 
                      result_name, "_diff_markers",
                      ".png")
  
  counts <- result %>%
    dplyr::filter(p.adjust < 0.05) %>%
    dplyr::filter(diff_markers == TRUE) %>%
    dplyr::group_by(cluster, alpha) %>%
    summarize(total_significant = n()) %>%
    dplyr::ungroup(cluster, alpha)
  
  if(length(counts$cluster %>% unique()) > 1 & length(counts$alpha %>% unique()) > 1){
    counts <- pivot_wider(counts, names_from = alpha, values_from = total_significant, values_fill = 0)
    counts_mat <- counts[, -1] %>% as.matrix()
    rownames(counts_mat) <- paste("Cluser", counts$cluster)
    counts_mat <- counts_mat[, sort(colnames(counts_mat))]
    
    png(plot_path, width = 800, height = 800)
    pheatmap(counts_mat, cluster_rows = F, cluster_cols = F, main = plot_title, display_numbers = TRUE)
    dev.off()
  }

})


```


In the end, enrichment analysis of the protein signatures (Elastic net $\alpha = 0.1$) using
the Curated CGP (Chemical and Genetic perturbations) and the Gene Ontology BP produced the most
significant terms for each cluster. We summarize these two below as a heatmaps. The Cluster 2 signature
was the most difficult to get enrichment for.



```{r}
library(pheatmap)

plot_results <- all_enrichment[c("gobp", "curated cgp")]
names(plot_results) <- c("Gene Ontology - Biological Process", 
                         "Curated Chemical and Genetic perturbations")

lapply(names(plot_results), function(result_name){
  
  plot_df <- plot_results[[result_name]] %>%
    dplyr::filter(diff_markers == FALSE) %>%
    dplyr::filter(p.adjust < 0.05) %>%
    dplyr::filter(alpha == 0.1) %>%
    dplyr::select(Description, cluster) %>%
    dplyr::mutate(sig = 1) %>%
    pivot_wider(names_from = cluster, values_from = sig)
  
  plot_df <- plot_df %>%
    dplyr::mutate(Description = str_wrap(Description, width = 55))
  
  plot_mat <- as.matrix(plot_df[, -1])
  rownames(plot_mat) <- plot_df$Description
  plot_mat[is.na(plot_mat)] <- 0 
  colnames(plot_mat) <- paste("Cluster", colnames(plot_mat))
  
  plot_title <- result_name
  result_name <- gsub(" ", "_", result_name)
  plot_path <- paste0("./elastic_net_binomial_models/protein_signatures_", result_name, ".png")
  
  png(plot_path, width = 2000, height = 2300, res = 200)
  pheatmap(plot_mat, cluster_cols = F, cluster_rows = F, 
           color = c("#F5F1E9", "#161729"), legend = F)
  dev.off()
})


```


