---
title: "Cluster 5 binoamial models"
author: "Camilo Posso"
date: "04/01/2022"
output: 
  html_document:
    code_folding: hide
    toc: true
editor_options: 
  chunk_output_type: inline
---

## Goal




```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(kableExtra)
library(glmnet)
library(groupdata2)
library(gridExtra)
library(grid)

source("../cluster_model_helper.R")

```



Finding the appropriate lambda range.



```{r eval=FALSE, include=FALSE}
bound_mult <- 1.0
bounds <- lapply(list(global_mat_train, phospho_mat_train, RNA_mat_train), function(data_mat){
  print(nrow(data_mat))
  lambda_min <- lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), function(chosen_alpha){
    cluster_lab <- enet_meta %>%
      mutate(cluster_lab = paste("Cluster", k.5)) %>%
      pull(cluster_lab) %>%
      as.factor()
    names(cluster_lab) <- enet_meta$Barcode.ID
    cluster_lab <- cluster_lab[colnames(data_mat)]
    
    model <- glmnet(x = t(data_mat),
                    y = cluster_lab,
                    family = 'multinomial',
                    type.measure = "class",
                    type.multinomial = "ungrouped",
                    alpha = chosen_alpha)
    min(model$lambda)
  }) %>% unlist()
  min_bounds <- as.matrix(lambda_min)*bound_mult 
  
  lambda_max <- lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), function(chosen_alpha){
    cluster_lab <- enet_meta %>%
      mutate(cluster_lab = paste("Cluster", k.5)) %>%
      pull(cluster_lab) %>%
      as.factor()
    names(cluster_lab) <- enet_meta$Barcode.ID
    cluster_lab <- cluster_lab[colnames(data_mat)]
    
    model <- glmnet(x = t(data_mat),
                    y = cluster_lab,
                    family = 'multinomial',
                    type.measure = "class",
                    type.multinomial = "ungrouped",
                    alpha = chosen_alpha)
    max(model$lambda)
  }) %>% unlist()
  max_bounds <- as.matrix(lambda_max)*bound_mult
  
  rownames(max_bounds) <- as.character(1:10/10)
  rownames(min_bounds) <- as.character(1:10/10)
  
  list("min_bounds" = min_bounds,
       "max_bounds" = max_bounds)
})

all_bounds <- list("min_bounds_Global" = bounds[[1]]$min_bounds,
                   "max_bounds_Global" = bounds[[1]]$max_bounds,
                   "min_bounds_Phospho" = bounds[[2]]$min_bounds,
                   "max_bounds_Phospho" = bounds[[2]]$max_bounds,
                   "min_bounds_RNA" = bounds[[3]]$min_bounds,
                   "max_bounds_RNA" = bounds[[3]]$max_bounds,
                   "bound_mult" = bound_mult)
saveRDS(all_bounds, "./enet_multinomial_data/all_bounds.RDS")

```




Setting up helper function for finding a statistically sound combination of $\alpha$
and $\lambda$ to build the model for cluster prediction on. 



```{r}
train_model <- function(data_mat, chosen_alpha, random_state){
  print(nrow(data_mat))
  print(random_state)
  cluster_lab <- enet_meta %>%
    mutate(cluster_lab = paste("Cluster", k.5)) %>%
    pull(cluster_lab) %>%
    as.factor()
  names(cluster_lab) <- enet_meta$Barcode.ID
  cluster_lab <- cluster_lab[colnames(data_mat)]
  
  meta_c_train <- enet_meta %>%
    mutate(Cluster = paste("Cluster", k.5)) %>%
    as.data.frame()
  rownames(meta_c_train) <- meta_c_train$Barcode.ID
  
  lambda_max <- 3.0
  ## Using an fixed lambda path. This is so that we can compare
  ## alpha values more appropriately. Makes a single lambda_path
  ## with containing an appropriate range for any alpha between 0 and 1. 
  lambda_path <- round(exp(seq(log(lambda_max), log(lambda_max*0.0001), 
                               length.out = 250)), digits = 10)
  # hist(log(lambda_path))
  
  set.seed(random_state)
  
  folds <- groupdata2::fold(meta_c_train, k = 5, cat_col = "Cluster") %>%
    select(Barcode.ID, Cluster, .folds) %>%
    as.data.frame()
  rownames(folds) <- folds$Barcode.ID
  folds <- folds[colnames(data_mat), ]
  
  fold_ids <- folds$.folds %>% as.numeric()
  
  model <- cv.glmnet(x = t(data_mat),
                     y = cluster_lab,
                     family = 'multinomial',
                     type.measure = "class",
                     alpha = chosen_alpha,
                     foldid = fold_ids,
                     lambda = lambda_path,
                     type.multinomial = "ungrouped")
  
  cross_validation_error <- data.frame(error = model$cvm, row.names = model$lambda)
  colnames(cross_validation_error) <- paste("alpha =", as.character(chosen_alpha))
  
  return(cross_validation_error)
}

```


Reproducible method to tune alpha and lambda. We use cross validation to find
the best performing alpha and lambda parameter. The folds used are balanced
for the class we are predicting, determined by the 'chosen_cluster' parameter.


```{r}
run_cv_grid <- function(data_mat,
                        n_trials = 20,
                        master_random_state = 117,
                        alpha_values = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)){
    
  set.seed(master_random_state)
  random_states <- sample(1:1000000, size = n_trials)
  
  all_cvm <- lapply(random_states, function(random_state){
    ## Running cv over the grid using the folds determined by random_state
    grid_results <- lapply(alpha_values, function(chosen_alpha){
      cvm_results <- train_model(data_mat, chosen_alpha, random_state)
    }) %>% do.call("cbind", .)
  })
  
  cvm_avg <- Reduce('+', all_cvm)/length(random_states)
  return(cvm_avg)
}

```



## Running over all clusters in K = 5.

Now we can run the above tool over all the clusters in $k = 5$, the question of
interest being how well can protein data recapture the clusters, which are
derived from a combined dataset which includes phospho data as well as RNA counts.


```{r eval=FALSE, include=FALSE}
results_global <- run_cv_grid(global_mat_train)
results_phospho <- run_cv_grid(phospho_mat_train)
results_RNA <- run_cv_grid(RNA_mat_train)

all_results <- list("Global" = results_global,
                    "Phospho" = results_phospho,
                    "RNA" = results_RNA)

saveRDS(all_results, "./enet_multinomial_data/elastic_net_multinomial_results_k_5.RDS")

```



```{r}
all_results <- readRDS("enet_multinomial_data/elastic_net_multinomial_results_k_5.RDS")
all_bounds <- readRDS("enet_multinomial_data/all_bounds.RDS")
bound_mult <- all_bounds$bound_mult

```



```{r}
get_model_metrics <- function(chosen_alpha, chosen_type){
  print(paste("alpha =", chosen_alpha, "type =", chosen_type))
  master_random_state <- 117
  set.seed(master_random_state)
  random_states <- sample(1:1000000, size = 20)
  
  if (chosen_type == "Global"){
      data_mat <- global_mat_train
    } else if (chosen_type == "Phospho"){
      data_mat <- phospho_mat_train
    } else if (chosen_type == "RNA"){
      data_mat <- RNA_mat_train
    }
    
  cluster_lab <- enet_meta %>%
    mutate(cluster_lab = paste("Cluster", k.5)) %>%
    pull(cluster_lab) %>%
    as.factor()
  names(cluster_lab) <- enet_meta$Barcode.ID
  cluster_lab <- cluster_lab[colnames(data_mat)]
  
  ## Getting best performing lambda from model results
  xx <- all_results[[chosen_type]]
  min_bounds <- all_bounds[[paste0("min_bounds_", chosen_type)]]
  max_bounds <- all_bounds[[paste0("max_bounds_", chosen_type)]]
  min_lambda <- min_bounds[as.character(chosen_alpha), ] %>% as.numeric()
  max_lambda <- max_bounds[as.character(chosen_alpha), ] %>% as.numeric()
  allowed_lambda <- (as.numeric(rownames(xx)) < max_lambda) & 
                    (as.numeric(rownames(xx)) > min_lambda)
  xx <- xx[allowed_lambda, ]
  col_name <- paste("alpha =", chosen_alpha)
  chosen_lambda <- rownames(xx)[[which.min(xx[[col_name]])]] %>%
    as.numeric()
  
  standardize_helper <- function(xx){
    xx <- sweep(xx, 1, apply(xx, 1, mean), FUN = '-')
    xx <- sweep(xx, 1, apply(xx, 1, sd), FUN = '/')
    return(xx)
  }
  
  class_error_helper <- function(prediction_df){
    prediction_df <- prediction_df %>%
      mutate(pred = sub("Predicted Cluster ", "", Cluster_pred),
             actual = sub("Actual Cluster ", "", Cluster))
    sapply(c("1", "2", "3", "4", "5"), function(chosen_cluster){
      prediction_df %>%
        mutate(pred = case_when(pred == chosen_cluster ~ pred,
                                TRUE ~ "0"),
               actual = case_when(actual == chosen_cluster ~ actual,
                                  TRUE ~ "0"),
               error = !(pred == actual)) %>%
        pull(error) %>%
        sum()/nrow(prediction_df)
    })
  }
  
  prediction_metrics <- lapply(random_states, function(random_state){
    set.seed(random_state)
    folds <- enet_meta %>%
      mutate(Cluster = paste("Cluster", k.5)) %>% 
      as.data.frame() %>%
      groupdata2::fold(k = 5, cat_col = "Cluster") %>%
      select(Barcode.ID, Cluster, .folds) %>%
      as.data.frame()
    rownames(folds) <- folds$Barcode.ID
    folds <- folds[colnames(data_mat), ]
    
    cluster_metrics <- lapply(c("1", "2", "3", "4", "5"), function(chosen_fold){
      samples_foldtrain <- folds %>%
        filter(.folds != chosen_fold) %>%
        pull(Barcode.ID)
      samples_foldtest <- folds %>%
        filter(.folds == chosen_fold) %>%
        pull(Barcode.ID)
      
      model <- glmnet(x = t(standardize_helper(data_mat[, samples_foldtrain])),
                      y = cluster_lab[samples_foldtrain],
                      alpha = chosen_alpha,
                      lambda = chosen_lambda,
                      family = "multinomial",
                      type.multinomial = "ungrouped")
      
      cluster_predictions <- predict(model, s = chosen_lambda, 
                                     newx = t(standardize_helper(data_mat[, samples_foldtest])), 
                                     type = "response") %>%
        matrix(ncol = 5) %>%
        as.data.frame()
      cluster_predictions$max <- apply(cluster_predictions, 1, max)
      cluster_predictions$Barcode.ID <- samples_foldtest
      cluster_predictions$Cluster_pred <- predict(model, s = chosen_lambda, 
                                          newx = t(standardize_helper(data_mat[, samples_foldtest])), 
                                          type = "class") %>%
        paste("Predicted", .) %>%
        factor(levels = paste("Predicted Cluster", 1:5))
      cluster_predictions$Cluster <- paste("Cluster", enet_meta[samples_foldtest, "k.5"]) %>%
        paste("Actual", .)
      
      dat <- table(cluster_predictions$Cluster_pred,cluster_predictions$Cluster)
      diagonal <- sapply(1:5, function(i){dat[i,i]})
      recall <- diagonal/apply(dat, 2, sum)
      precision <- diagonal/apply(dat, 1, sum)
      missclass_error <- class_error_helper(cluster_predictions)
      
      data.frame(Cluster = paste("Cluster", 1:5), recall = recall, 
                 precision = precision, missclass_error = missclass_error, 
                 fold = chosen_fold)
    }) %>% do.call("rbind", .) %>%
      ## If precision is 0/0 we get NaN, so we turn to 0 here.
      mutate(precision = case_when(!is.na(precision) ~ precision,
                                   TRUE ~ 0)) %>%
      group_by(Cluster) %>%
      summarize(recall = mean(recall),
                precision = mean(precision),
                missclass_error = mean(missclass_error)) %>%
      mutate(random_state = random_state)
    
  }) %>% do.call("rbind", .) %>%
    group_by(Cluster) %>%
    summarize(Recall = mean(recall), 
              Precision = mean(precision),
              missclass_error = mean(missclass_error))  %>% 
    mutate(chosen_alpha = chosen_alpha,
           data_type = chosen_type)
  
  return(prediction_metrics)
}

```


Putting all metrics into a table.


```{r eval=FALSE, include=FALSE}
model_metrics <- lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), function(chosen_alpha){
  metrics <- get_model_metrics(chosen_alpha, "Global") %>%
    rbind(get_model_metrics(chosen_alpha, "Phospho")) %>%
    rbind(get_model_metrics(chosen_alpha, "RNA"))
  return(metrics)
}) %>% do.call("rbind", .)

write.table(model_metrics, "enet_multinomial_data/all_model_metrics.txt", 
            sep = "\t", quote = F)

```



Plotting errors by cluster



```{r}
model_metrics <- read.table("enet_multinomial_data/all_model_metrics.txt", sep = "\t") %>%
  dplyr::rename(alpha = chosen_alpha)

plot_metrics <- function(chosen_alpha){
  global_model_metrics <- model_metrics %>%
    filter(data_type == "Global") %>%
    filter(alpha == chosen_alpha)
  phospho_model_metrics <- model_metrics %>%
    filter(data_type == "Phospho") %>%
    filter(alpha == chosen_alpha)
  rna_model_metrics <- model_metrics %>%
    filter(data_type == "RNA") %>%
    filter(alpha == chosen_alpha)
  
  summary <- data.frame(Cluster = c("1", "2", "3", "4", "5"), 
                        Recall = global_model_metrics$Recall,
                        Precision = global_model_metrics$Precision,
                        `Classification error` = global_model_metrics$missclass_error,
                        data_type = "Global") %>%
    rbind(data.frame(Cluster = c("1", "2", "3", "4", "5"), 
                     Recall = phospho_model_metrics$Recall,
                     Precision = phospho_model_metrics$Precision,
                     `Classification error` = phospho_model_metrics$missclass_error,
                     data_type = "Phospho")) %>%
    rbind(data.frame(Cluster = c("1", "2", "3", "4", "5"), 
                     Recall = rna_model_metrics$Recall,
                     Precision = rna_model_metrics$Precision,
                     `Classification error` = rna_model_metrics$missclass_error,
                     data_type = "RNA"))
  
  p_recall <- ggplot(summary, aes(x = Cluster, y = Recall, fill = data_type)) + 
    geom_bar(stat = 'identity', width = 0.33, position = position_dodge()) + 
    scale_fill_manual(values = data_type_colors) +
    ggtitle(paste0("Recall by cluster, ",
                   "alpha = ", chosen_alpha)) +
    xlab("Cluster") + ylim(0, 1) +
    theme_minimal()
  
  p_precision <- ggplot(summary, aes(x = Cluster, y = Precision, fill = data_type)) + 
    geom_bar(stat = 'identity', width = 0.33, position = position_dodge()) + 
    scale_fill_manual(values = data_type_colors) +
    ggtitle(paste0("Precision by cluster, ",
                   "alpha = ", chosen_alpha)) +
    xlab("Cluster") + ylim(0, 1) +
    theme_minimal()
  
  p_classerror <- ggplot(summary, aes(x = Cluster, y = Classification.error, fill = data_type)) + 
    geom_bar(stat = 'identity', width = 0.33, position = position_dodge()) + 
    scale_fill_manual(values = data_type_colors) +
    ggtitle(paste0("Classification error by cluster, ",
                   "alpha = ", chosen_alpha)) +
    xlab("Cluster") + ylim(0, 0.2) +
    theme_minimal()
  
  top_grob <- textGrob(paste0("Model metrics alpha = ", chosen_alpha), 
                       gp = gpar(fontsize = 15))
  p_all <- arrangeGrob(p_recall, p_precision, p_classerror, 
                       ncol = 3, top = top_grob)
  ggsave(paste0("enet_multinomial_data/model_metrics_alpha_", chosen_alpha, ".pdf"), p_all, 
         width = 15, height = 6)
  
  return(list(p_recall, p_precision, p_classerror))
}

```


```{r eval=FALSE, include=FALSE}
lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), plot_metrics)

```

