---
title: "Differentially expressed features in mutated samples: Network Analysis"
author: "Sara Gosline"
date: "3/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Network analysis 

The goal of this document is to study the larger networks that comprise the differentially expressed features. We want to map all the features to one single network.

```{r load data,message=F,echo=T,warning=F}

library(PCSF)
library(dplyr)
library(ggplot2)

source("../../util/synapseUtil.R")
syn<-synapseLogin()


combined_diffexp_results <- read.table(syn$get("syn27828350")$path, sep = "\t")

network_input <- combined_diffexp_results %>%
  filter(adj.P.Val < 0.05) %>%
  dplyr::rename(data_type = datatype) %>%
  mutate(weight = abs(logFC), nrand = 100)

```


## Create function to take proteins by type and map them to network

We need to keep track of two things - one is the type of feature we are handling, and the other is the weight for the network.

```{r network mapping,message=F,echo=T,warning=F}

build_run_network <- function(gene_name, 
                              gene_score, 
                              gene_type,
                              mutation,
                              rands=100,
                              store=FALSE){
  
  library(amlresistancenetworks)
  print(unique(mutation))
  
  wes_inds <- which(gene_type=='WES')
  rna_inds <- which(gene_type=='RNA')
  prot_inds <- which(gene_type=='Global')
  phos_inds <- which(gene_type=='Phospho')
  
  netname=paste0("differentially_expressed_in_", mutation, "_mutations")
  
    ##rename WES
  wes_terms <- gene_score[wes_inds]
  names(wes_terms) <- gene_name[wes_inds]
  
  #lets separate out the substrates from the phosphosite data
  phgenes<-data.frame(genes=gene_name[phos_inds],val=gene_score[phos_inds])%>%
    tidyr::separate(genes, into=c('gene','mod'),sep='-')%>%
    mutate(modlist=strsplit(mod,split='s|t|y'))%>%
    tidyr::unnest()%>%
    rowwise()%>%
    mutate(newgene=paste(gene,modlist,sep='-'))
  
  kin_terms <- phgenes$val
  names(kin_terms) <- phgenes$newgene
  
 # print(kin_terms)
  prot_terms <- gene_score[prot_inds]
  names(prot_terms) <- gene_name[prot_inds]
  

  rna_terms <- gene_score[rna_inds]
  names(rna_terms) <- gene_name[rna_inds]
  
  start_time <- Sys.time()
  print(start_time)
  netlist = amlresistancenetworks::computePhosphoNetwork(phos.vals=kin_terms,
                                                     prot.vals=prot_terms,
                                                     gene.vals=wes_terms,
                                                     rna.vals=rna_terms,
                                                     beta=1.5,
                                                     nrand=rands,
                                                     fname=netname)
  print(Sys.time() - start_time)
  graph =netlist$graph
  fname = netlist$fname[1]
  print(fname)
  if(store){
    synapseStore(fname,'syn27831953')
  }  
  return(graph)
}

```

Now we can run the analysis


```{r k5 cluster analysis}

nets <- network_input %>%
  group_by(mutation) %>%
  dplyr::summarize(net=list(build_run_network(feature,
                                           weight,
                                           data_type,
                                           mutation,
                                           nrand,store=FALSE)))


## KRAS+NRAS genes (12 in total) are making the PCSF_rand function crash:
    ## Error in PCSF_rand(ppi, abs(terms), n = 10, r = 0.2, w = 4, b = beta,  : 
    ## Subnetwork can not be identified for a given parameter set.
    ## Provide a compatible b or mu value with your terminal prize list...
# To recreate, get to the try(PCSF_rand(ppi, abs(terms), ...) line
# debug(build_run_network)
# k5_nets <- network_input %>%
#   filter(mutation == "KRAS+NRAS") %>%
#   mutate(nrand=10,
#          weight = -log10(adj.P.Val))%>%
#   group_by(mutation)%>%
#   dplyr::summarize(net=list(build_run_network(feature,
#                                            weight,
#                                            data_type,
#                                            mutation,
#                                            nrand,store=FALSE)))

getNetStats <- function(net_list,orig_tab){
  require(igraph)
  ##create table of metagene features based on wht was included/excluded from the network
  node_tab<-purrr::map2_dfr(.x=net_list$Cluster,.y=net_list$net,.f=function(Cluster,net){
    g<-net
    data.frame(node=names(V(g)),prizes=vertex.attributes(g)$prize,
               types=vertex.attributes(g)$type,lfc=vertex.attributes(g)$logFoldChange,
               geneTypes=vertex.attributes(g)$nodeType,Cluster=rep(Cluster,length(V(g))))
  } )
  
  graph_counts <- node_tab%>%group_by(geneTypes,types,Cluster)%>%summarize(count=n())%>%
    dplyr::rename(data_type='geneTypes',source='types')
  orig_counts <- orig_tab%>%group_by(data_type,Cluster)%>%summarize(count=n())%>%
    mutate(source='Metagene')

  ##counts of how man proteins/genes/phosphosites in each
  return(rbind(graph_counts,orig_counts))  
}

# DT::datatable(getNetStats(nets,network_input))

```


```{r}
source("../../util/make_plots_util.R")
net_paths <- list.files("./networks/", "network_differentially_expressed_in_")
for (net_path in net_paths){
  upload.plot(net_path, parentId = "syn27831953")
}


```


```{r}
## FLT3 already done. IDH1 fails with given mu and b values.
## IDH11+IDH2, IDH2, and KRAS all done, KRAS+NRAS fails.
## NPM1_clinical and TP53 also done
network_input <- network_input %>%
  filter(mutation != "FLT3.ITD") %>%
  filter(mutation != "IDH1+IDH2") %>%
  filter(mutation != "IDH2") %>%
  filter(mutation != "KRAS") %>%
  filter(mutation != "TP53") %>%
  filter(mutation != "NPM1_clinical")

table(network_input$mutation)

```


Now we can run these two smaller inputs with a greater beta value.


```{r}
## Tyring another beta value with networks that gave empty graphs before.
build_run_network <- function(gene_name, 
                              gene_score, 
                              gene_type,
                              mutation,
                              rands=100,
                              store=FALSE){
  
  library(amlresistancenetworks)
  print(unique(mutation))
  
  wes_inds <- which(gene_type=='WES')
  rna_inds <- which(gene_type=='RNA')
  prot_inds <- which(gene_type=='Global')
  phos_inds <- which(gene_type=='Phospho')
  
  netname=paste0("differentially_expressed_in_", mutation, "_mutations")
  
    ##rename WES
  wes_terms <- gene_score[wes_inds]
  names(wes_terms) <- gene_name[wes_inds]
  
  #lets separate out the substrates from the phosphosite data
  phgenes<-data.frame(genes=gene_name[phos_inds],val=gene_score[phos_inds])%>%
    tidyr::separate(genes, into=c('gene','mod'),sep='-')%>%
    mutate(modlist=strsplit(mod,split='s|t|y'))%>%
    tidyr::unnest()%>%
    rowwise()%>%
    mutate(newgene=paste(gene,modlist,sep='-'))
  
  kin_terms <- phgenes$val
  names(kin_terms) <- phgenes$newgene
  
 # print(kin_terms)
  prot_terms <- gene_score[prot_inds]
  names(prot_terms) <- gene_name[prot_inds]
  

  rna_terms <- gene_score[rna_inds]
  names(rna_terms) <- gene_name[rna_inds]
  
  start_time <- Sys.time()
  print(start_time)
  netlist = amlresistancenetworks::computePhosphoNetwork(phos.vals=kin_terms,
                                                     prot.vals=prot_terms,
                                                     gene.vals=wes_terms,
                                                     rna.vals=rna_terms,
                                                     beta=2.0,
                                                     nrand=rands,
                                                     fname=netname)
  print(Sys.time() - start_time)
  graph =netlist$graph
  fname = netlist$fname[1]
  print(fname)
  if(store){
    synapseStore(fname,'syn27831953')
  }  
  return(graph)
}


nets <- network_input %>%
  group_by(mutation) %>%
  dplyr::summarize(net=list(build_run_network(feature,
                                           weight,
                                           data_type,
                                           mutation,
                                           nrand,store=FALSE)))

```




```{r}
## IDH1 worked, but KRAS+NRAS is still failing
network_input <- network_input %>%
  filter(mutation != "IDH1")

table(network_input$mutation)

```


Now we can run these two smaller inputs with a greater beta value.


```{r}
## Tyring another beta value with networks that gave empty graphs before.
build_run_network <- function(gene_name, 
                              gene_score, 
                              gene_type,
                              mutation,
                              rands=100,
                              store=FALSE){
  
  library(amlresistancenetworks)
  print(unique(mutation))
  
  wes_inds <- which(gene_type=='WES')
  rna_inds <- which(gene_type=='RNA')
  prot_inds <- which(gene_type=='Global')
  phos_inds <- which(gene_type=='Phospho')
  
  netname=paste0("differentially_expressed_in_", mutation, "_mutations")
  
    ##rename WES
  wes_terms <- gene_score[wes_inds]
  names(wes_terms) <- gene_name[wes_inds]
  
  #lets separate out the substrates from the phosphosite data
  phgenes<-data.frame(genes=gene_name[phos_inds],val=gene_score[phos_inds])%>%
    tidyr::separate(genes, into=c('gene','mod'),sep='-')%>%
    mutate(modlist=strsplit(mod,split='s|t|y'))%>%
    tidyr::unnest()%>%
    rowwise()%>%
    mutate(newgene=paste(gene,modlist,sep='-'))
  
  kin_terms <- phgenes$val
  names(kin_terms) <- phgenes$newgene
  
 # print(kin_terms)
  prot_terms <- gene_score[prot_inds]
  names(prot_terms) <- gene_name[prot_inds]
  

  rna_terms <- gene_score[rna_inds]
  names(rna_terms) <- gene_name[rna_inds]
  
  start_time <- Sys.time()
  print(start_time)
  netlist = amlresistancenetworks::computePhosphoNetwork(phos.vals=kin_terms,
                                                     prot.vals=prot_terms,
                                                     gene.vals=wes_terms,
                                                     rna.vals=rna_terms,
                                                     beta=2.5,
                                                     nrand=rands,
                                                     fname=netname)
  print(Sys.time() - start_time)
  graph =netlist$graph
  fname = netlist$fname[1]
  print(fname)
  if(store){
    synapseStore(fname,'syn27831953')
  }  
  return(graph)
}


nets <- network_input %>%
  group_by(mutation) %>%
  dplyr::summarize(net=list(build_run_network(feature,
                                           weight,
                                           data_type,
                                           mutation,
                                           nrand,store=FALSE)))

```



```{r}
source("../../util/make_plots_util.R")
net_paths <- list.files("./networks/", "network_differentially_expressed_in_") %>%
  paste0("./networks/", .)
for (net_path in net_paths){
  upload.plot(net_path, parentId = "syn27831953")
}

```