---
title: "BeatAML Pathway Enrichment Analysis"
author: "Michael Nestor"
date: "6/8/2021"
output:
  html_document
---

```{r setup, message=F, warning=F}
knitr::opts_chunk$set(echo = F, message=F, warning=F)

library(dplyr)
library(tidyr)
library(MSnbase)
library(kableExtra)
library(ggplot2)

library(amlresistancenetworks)
library(leapr)

data(msigdb)
data(ncipid)
```

```{r}
syn = reticulate::import("synapseclient")
syn$login()

globalData <- querySynapseTable('syn25808020') %>%
  mutate(Sample = sprintf("%02d", Sample))

phenoData <- globalData %>%
  select(-Gene, -LogRatio) %>%
  distinct()
rownames(phenoData) <- phenoData$Sample

exprsData <- globalData %>%
  select(Gene, LogRatio, Sample) %>%
  pivot_wider(id_cols="Gene",
              names_from="Sample",
              values_from="LogRatio",
              values_fill=NA_real_) %>%
  as.data.frame()
rownames(exprsData) <- exprsData$Gene
exprsData <- exprsData %>% select(-Gene)

msnset <- MSnSet(as.matrix(exprsData))
pData(msnset) <- phenoData
```

# Summary statistics 

```{r}
pData(msnset) %>%
  group_by(specimen.type, PostChemotherapy) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = specimen.type, y = n, fill = PostChemotherapy)) + 
    geom_bar(position="dodge", stat="identity") +
  xlab("Specimen type") +
  ylab("Count") +
  ggtitle("BeatAML Patient Cohort Summary Statistics") +
  theme_bw()
```

# Correlation enrichment

This section applies the correlation enrichment algorithm from the `leapR` package. The pathways are from the NCI’s Pathway Interaction Database (PID). We report the
pathways with a Benjamini-Hochberg-adjusted p-value less than 5%.

```{r}
x <- exprs(msnset)
x <- sweep(x, 1, apply(x, 1, mean), "-")
x <- sweep(x, 1, apply(x, 1, sd), "/")
leapR_out <- leapR(geneset=ncipid,
             enrichment_method = "correlation_enrichment",
             datamatrix = x)
```

```{r}
res <- leapR_out
res <- res[order(res[,"pvalue"]),]

res$pathway <- rownames(res)
rownames(res) <- NULL
res <- select(res, pathway, everything())

res <- res[res$BH_pvalue<0.05,]
res <- filter(res_sig, !is.na(BH_pvalue))

res <- res %>% select(pathway, ingroup_n, ingroup_mean, outgroup_n, 
                          outgroup_mean, pvalue, BH_pvalue)

kbl(res) %>%
  kable_paper("hover", full_width = F)
```

```{r, fig.width=6, fig.height=20}
ggplot(res, aes(x=-log10(BH_pvalue), y=reorder(pathway, -BH_pvalue))) + geom_bar(stat="identity") +
  ggtitle("Top significantly correlated pathways") +
  xlab("Negative log p-value") +
  ylab("Pathway")
```

# Comparison of pre- and post- chemotherapy samples

This section applies the enrichment comparison algorithm from the `leapR` package. We compare chemotherapy-treated samples in bone marrow aspirate and peripheral
blood samples independently. (The leukapheresis samples don't appear to have a
large enough sample size for this analysis). Once again we report the
pathways with a Benjamini-Hochberg-adjusted p-value less than 5%.

## Bone marrow aspirate

```{r}
m <- msnset[,msnset$specimen.type=="Bone Marrow Aspirate"]

group1 = which(grepl("Post-Chemotherapy", m$Specimen.access.group.concatenated))
group2 = which(!grepl("Post-Chemotherapy", m$Specimen.access.group.concatenated))

x <- exprs(m)
x <- sweep(x, 1, apply(x, 1, mean), "-")
x <- sweep(x, 1, apply(x, 1, sd), "/")

leapR_out = leapR(geneset=ncipid,
                  enrichment_method='enrichment_comparison',
                  datamatrix=x,
                  primary_columns=group1,
                  secondary_columns=group2)
```

```{r}
res <- leapR_out
res <- res[order(res[,"pvalue"]),]

res <- res[res$BH_pvalue<0.05,]
res <- filter(res, !is.na(BH_pvalue))

res$pathway <- rownames(res)
rownames(res) <- NULL
res <- select(res, pathway, everything())

res_tab <- res %>% select(pathway, ingroup_n, ingroup_mean, outgroup_n, 
                          outgroup_mean, zscore, oddsratio,
                          pvalue, BH_pvalue)
library(kableExtra)
kbl(res_tab) %>%
  kable_paper("hover", full_width = F)
```

```{r, fig.width=6.5, fig.height=20}
ggplot(res, aes(x=-log10(BH_pvalue), y=reorder(pathway, -BH_pvalue))) + geom_bar(stat="identity") +
  ggtitle("Top significant pathways before/after chemo\n (bone marrow aspirate)") +
  xlab("Negative log p-value") +
  ylab("Pathway")
#ggsave("correlation_enrichment_ex10.png", height=10)
```


## Peripheral blood

```{r}
m <- msnset[,msnset$specimen.type=="Peripheral Blood"]

group1 = which(grepl("Post-Chemotherapy", m$Specimen.access.group.concatenated))
group2 = which(!grepl("Post-Chemotherapy", m$Specimen.access.group.concatenated))

x <- exprs(m)
x <- sweep(x, 1, apply(x, 1, mean), "-")
x <- sweep(x, 1, apply(x, 1, sd), "/")

leapR_out = leapR(geneset=ncipid,
                  enrichment_method='enrichment_comparison',
                  datamatrix=x,
                  primary_columns=group1,
                  secondary_columns=group2)
```

```{r}
res <- leapR_out
res <- res[order(res[,"pvalue"]),]

res <- res[res$BH_pvalue<0.05,]
res <- filter(res, !is.na(BH_pvalue))

res$pathway <- rownames(res)
rownames(res) <- NULL
res <- select(res, pathway, everything())

res_tab <- res %>% select(pathway, ingroup_n, ingroup_mean, outgroup_n, 
                          outgroup_mean, zscore, oddsratio,
                          pvalue, BH_pvalue)
kbl(res_tab) %>%
  kable_paper("hover", full_width = F)
```

```{r, fig.width=6.5, fig.height=20}
ggplot(res, aes(x=-log10(BH_pvalue), y=reorder(pathway, -BH_pvalue))) + geom_bar(stat="identity") +
  ggtitle("Top significant pathways before/after chemo\n (peripheral blood)") +
  xlab("Negative log p-value") +
  ylab("Pathway")
```


