---
title: "K = 4 multinomial predictions"
author: "Camilo Posso"
date: "04/27/2022"
output: 
  html_document:
    code_folding: hide
    toc: true
editor_options: 
  chunk_output_type: inline
---

## Goal

Predict subtypes on the 51 held out samples, as well as entirely new datasets. CHeck
survival, mutation, etc.


```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(glmnet)
library(survminer)
library(survival)
library(gridExtra)
library(grid)
library(ggpattern)

source("../NMF_helper.R")
source("../cluster_model_helper.R")

```


Loading model results.


```{r}
all_results <- readRDS("enet_multinomial_data/elastic_net_multinomial_results_k_4.RDS")
all_bounds <- readRDS("enet_multinomial_data/all_bounds.RDS")
bound_mult <- all_bounds$bound_mult

```



# Predicting subtypes on held-out samples from 210 cohort.

Here we train a prediction model using the full dataset and the best performing hyperparameters
to extend cluster 4 (good prognosis) to the 51 unclustered samples.


```{r}
## Function to make predictions and plots. 
make_prediction <- function(chosen_alpha, data_type, robust){
  if (data_type == "Global"){
    data_mat <- global_mat_train
    data_mat_test <- global_mat_test
    data_mat_full <- global.data %>%
      dplyr::rename(feature = Gene) %>%
      filter(feature %in% global_features) %>%
      pivot_wider(names_from = Barcode.ID, 
                  values_from = LogRatio) %>% 
      as.data.frame()
    rownames(data_mat_full) <- data_mat_full$feature
    data_mat_full <- data_mat_full[, -1] %>% as.matrix()
  } else if (data_type == "Phospho"){
    data_mat <- phospho_mat_train
    data_mat_test <- phospho_mat_test
    data_mat_full <- phospho.data %>%
      select(Barcode.ID, SiteID, LogRatio) %>%
      dplyr::rename(feature = SiteID) %>%
      filter(feature %in% phospho_features) %>%
      pivot_wider(names_from = Barcode.ID, 
                  values_from = LogRatio) %>% 
      as.data.frame()
    rownames(data_mat_full) <- data_mat_full$feature
    data_mat_full <- data_mat_full[, -1] %>% as.matrix()
  }
  
  xx <- all_results[[data_type]]
  data_mat_full <- sweep(data_mat_full, 1, apply(data_mat_full, 1, mean), FUN = '-')
  data_mat_full <- sweep(data_mat_full, 1, apply(data_mat_full, 1, sd), FUN = '/')
  
  ## Key line
  if (robust){
    data_mat_test <- data_mat_full[, colnames(data_mat_test)]
  }
  
  min_bounds <- all_bounds[[paste0("min_bounds_", data_type)]]
  max_bounds <- all_bounds[[paste0("max_bounds_", data_type)]]
  min_lambda <- min_bounds[as.character(chosen_alpha), ]
  max_lambda <- max_bounds[as.character(chosen_alpha), ]
  allowed_lambda <- (as.numeric(rownames(xx)) < max_lambda) & 
                    (as.numeric(rownames(xx)) > min_lambda)
  
  xx <- xx[allowed_lambda, ]
  col_name <- paste("alpha =", chosen_alpha)
  chosen_lambda <- rownames(xx)[[which.min(xx[[col_name]])]] %>%
    as.numeric()
  print(chosen_lambda)
  
  cluster_lab <- enet_meta %>%
    mutate(cluster_lab = paste("Cluster", k.4)) %>%
    pull(cluster_lab) %>%
    as.factor()
  names(cluster_lab) <- enet_meta$Barcode.ID
  cluster_lab <- cluster_lab[colnames(data_mat)]

  model <- glmnet(x = t(data_mat),
                  y = cluster_lab,
                  family = 'multinomial',
                  alpha = chosen_alpha,
                  lambda = chosen_lambda)
  
  cluster_predictions <- predict(model, s = chosen_lambda, 
                                 newx = t(data_mat_test), 
                                 type = "response")[, , 1] %>% as.data.frame()
  colnames(cluster_predictions) <- paste("New", colnames(cluster_predictions))
  cluster_predictions$max_new <- apply(cluster_predictions, 1, max) %>% unname()
  cluster_predictions$pred_cluster <- predict(model, s = chosen_lambda, 
                                              newx = t(data_mat_test), 
                                              type = "class")
  cluster_predictions$Barcode.ID <- rownames(cluster_predictions)
  
  cluster_predictions_full <- predict(model, s = chosen_lambda, 
                                      newx = t(data_mat_full), 
                                      type = "response")[, , 1] %>% as.data.frame()
  colnames(cluster_predictions_full) <- paste("Full", colnames(cluster_predictions_full))
  cluster_predictions_full$max_full <- apply(cluster_predictions_full, 1, max) %>% unname()
  cluster_predictions_full$pred_cluster_v2 <- predict(model, s = chosen_lambda, 
                                                      newx = t(data_mat_full), 
                                                      type = "class")
  cluster_predictions_full$Barcode.ID <- rownames(cluster_predictions_full)
  
  out <- full_join(cluster_predictions, cluster_predictions_full, by = "Barcode.ID")
  return(out)
}


make_prediction_plot <- function(chosen_alpha, data_type, confidence_cutoff = 0.40, robust = TRUE){
  cluster_predictions <- make_prediction(chosen_alpha, data_type, robust) %>%
    mutate(pred_cluster = sub("Cluster ", "", pred_cluster),
           pred_cluster_v2 = sub("Cluster ", "", pred_cluster_v2))
  if (robust){
    path_ending <- "_robust_normalization.pdf"
  } else {
    path_ending <- ".pdf"
  }
  
  meta_c <- left_join(meta, cluster_assignments, by = "Barcode.ID") %>%
    mutate(overallSurvival = as.numeric(overallSurvival),
           k.4 = as.character(k.4)) %>%
    left_join(cluster_predictions, by = "Barcode.ID")
  rownames(meta_c) <- meta_c$Barcode.ID
  
  meta_c_new <- meta_c %>%
      filter(is.na(k.4)) %>%
    mutate(conf = case_when(max_new <= confidence_cutoff ~ "Low confidence",
                            TRUE ~ "OK")) %>%
    dplyr::rename(Cluster = pred_cluster)
  meta_c_plot <- meta_c_new %>%
    filter(conf == "OK")
  meta_c_train <- meta_c %>%
    filter(!is.na(k.4)) %>%
    dplyr::rename(Cluster = k.4)
  
  plot_title <- paste0("Predicting on 51 new samples, conf > ", 
                       confidence_cutoff, "(", nrow(meta_c_plot), " samples)")
  color_palette <- c("1" = subtype_colors[[1]],
                     "2" = subtype_colors[[2]],
                     "3" = subtype_colors[[3]],
                     "4" = subtype_colors[[4]])
  
  sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ Cluster, data = meta_c_train)
  p1 <- ggsurvplot(sfit, color = "Cluster", data = meta_c_train,
                   palette = color_palette, pval = TRUE)$plot +
          theme(legend.text = element_text(size = 12, color = "black"),
                legend.title = element_text(size = 12, color = "black")) + 
          ggtitle("159 Clustered samples")
  
  meta_c$Cluster <- meta_c$pred_cluster_v2
  sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ Cluster, data = meta_c)
  p2 <- ggsurvplot(sfit, color = "Cluster", data = meta_c,
                   palette = color_palette, pval = TRUE)$plot +
          theme(legend.text = element_text(size = 12, color = "black"),
                legend.title = element_text(size = 12, color = "black")) + 
          ggtitle("Predicting on all 210 samples")
  
  sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ Cluster, data = meta_c_plot)
  p3 <- ggsurvplot(sfit, color = "Cluster", data = meta_c_plot,
                   palette = color_palette, pval = TRUE)$plot +
          theme(legend.text = element_text(size = 12, color = "black"),
                legend.title = element_text(size = 12, color = "black")) +
          ggtitle(plot_title)
  
  sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ Cluster, data = meta_c_new)
  p4 <- ggsurvplot(sfit, color = "Cluster", data = meta_c_new,
                   palette = color_palette,  pval = TRUE)$plot +
          theme(legend.text = element_text(size = 12, color = "black"),
                legend.title = element_text(size = 12, color = "black")) + 
          ggtitle("Predicting on 51 new samples")
  
  p_all <- arrangeGrob(p1, p2, p4, ncol = 2, nrow = 2,
                       top = textGrob(paste0("Survival plots, ", data_type, ", alpha = ", chosen_alpha), 
                                      gp = gpar(fontsize = 20)))
  ggsave(paste0("enet_multinomial_data/predicted_survival_", 
                data_type, "_", chosen_alpha, path_ending), 
         width = 16, height = 12, p_all)
  return(list(p1, p2, p4))
}

```



```{r eval=FALSE, include=FALSE}
## Making survival plots by predicted subtype
lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), function(chosen_alpha){
  make_prediction_plot(chosen_alpha, "Global", robust = TRUE)
  make_prediction_plot(chosen_alpha, "Phospho", robust = TRUE)
  make_prediction_plot(chosen_alpha, "Global", robust = FALSE)
  make_prediction_plot(chosen_alpha, "Phospho", robust = FALSE)
})

prediction_survival_plots <- make_prediction_plot(0.7, "Global", robust = TRUE)
ggsave("enet_multinomial_data/prediction_survival_alpha_0.7_all_samples.pdf", 
       width = 7, height = 7, plot = prediction_survival_plots[[2]])
ggsave("enet_multinomial_data/prediction_survival_alpha_0.7_heldout_samples.pdf", 
       width = 7, height = 7, plot = prediction_survival_plots[[3]])

```



### Mutation status by subtype



```{r}
## Computing enrichment status of each mutation within the subtypes.
all_mutation_data <- load_mutational_sample_data()
mutation_data <- all_mutation_data %>%
  mutate(value = TRUE) %>%
  pivot_wider(names_from = Gene, values_from = value, values_fn = any)
mutation_data <- mutation_data[, c("Barcode.ID", "KRAS+NRAS", "NPM1_clinical")] %>%
  full_join(meta[, c("Barcode.ID", "FLT3.ITD")], by = "Barcode.ID") %>%
  mutate(RAS = case_when(!is.na(`KRAS+NRAS`) ~ `KRAS+NRAS`,
                          TRUE ~ FALSE),
         NPM1_clinical = case_when(!is.na(NPM1_clinical) ~ NPM1_clinical,
                                   TRUE ~ FALSE),
         FLT3 = case_when(FLT3.ITD == "TRUE" ~ TRUE,
                          TRUE ~ FALSE)) %>%
  select(Barcode.ID, FLT3, NPM1_clinical, RAS)

mut_enrichment <- function(sample_categories, assignments, log.scale = FALSE) {
  k = length(unique(assignments$Cluster))
  
  sample_categories <- sample_categories %>%
    filter(rownames(.) %in% rownames(assignments)) %>%
    left_join(assignments, by = "Barcode.ID") %>%
    column_to_rownames("Barcode.ID") %>% 
    as.data.frame()
  
  categories <- sample_categories %>%
    select(-Cluster) %>%
    colnames()
  
  out <- lapply(categories, function(category){
    
    p.values <- lapply(1:k, function(i){
      cluster.df <- sample_categories %>%
        mutate(Cluster = case_when(Cluster == i ~ TRUE,
                                   TRUE ~ FALSE)) %>%
        mutate(Cluster = as.factor(Cluster))
      dat <- table(cluster.df[[category]], cluster.df$Cluster)
      background_dist <- apply(dat, 1, sum)
      background_prop <- background_dist[[2]]/(background_dist[[2]] + background_dist[[1]])
      cluster_prop <- dat[2, 2]/(dat[1, 2] + dat[2, 2])
      p_value <- fisher.test(dat, workspace = 2e8, alternative = "greater")$p.value
      if (log.scale) {
        p_value <- -log10(p_value)
      } else {
        p_value <- p_value
      }
      xx <- data.frame(background_prop, cluster_prop, p_value)
      colnames(xx) <- c(paste("Background", category, "positive fraction"),
                        paste(category, "positive fraction"),
                        paste("Enrichment", category, "p-value"))
      rownames(xx) <- paste("Cluster", i)
      return(xx)
    }) %>% do.call("rbind", .)
    
  }) %>% do.call(cbind, .) %>% 
    as.data.frame()
  return(out)
}


mut_enrichment_plot <- function(chosen_alpha, data_type, robust = TRUE){
  if (robust){
    path_ending <- "_robust_normalization.pdf"
  } else {
    path_ending <- ".pdf"
  }
  
  meta_pred <- make_prediction(chosen_alpha, data_type, robust) %>%
    select(Barcode.ID, pred_cluster, pred_cluster_v2) %>%
    full_join(mutation_data, by = "Barcode.ID") %>%
    mutate(pred_cluster = sub("Cluster ", "", pred_cluster),
           pred_cluster_v2 = sub("Cluster ", "", pred_cluster_v2))
  rownames(meta_pred) <- meta_pred$Barcode.ID
  
  sample_assn <- meta_pred[, c("Barcode.ID", "pred_cluster")] %>%
    dplyr::rename(Cluster = pred_cluster) %>%
    filter(!is.na(Cluster))
  sample_cats <- meta_pred[rownames(sample_assn), c("Barcode.ID", "FLT3", "RAS", "NPM1_clinical")]
  mutation_enrichment <- mut_enrichment(sample_cats, sample_assn) %>%
    mutate(Cluster = 1:4) %>%
    mutate(pred_type = "51 Samples")
  
  sample_assn <- meta_pred[, c("Barcode.ID", "pred_cluster_v2")] %>%
    dplyr::rename(Cluster = pred_cluster_v2) %>%
    filter(!is.na(Cluster))
  sample_cats <- meta_pred[rownames(sample_assn), c("Barcode.ID", "FLT3", "RAS", "NPM1_clinical")]
  mutation_enrichment <- mut_enrichment(sample_cats, sample_assn) %>%
    mutate(Cluster = 1:4) %>%
    mutate(pred_type = "210 Samples") %>%
    rbind(mutation_enrichment)
  
  sample_assn <- enet_meta[, c("Barcode.ID", "k.4")] %>%
    dplyr::rename(Cluster = k.4) %>%
    filter(!is.na(Cluster)) 
  sample_cats <- meta_pred[rownames(sample_assn), c("Barcode.ID", "FLT3", "RAS", "NPM1_clinical")]
  mutation_enrichment <- mut_enrichment(sample_cats, sample_assn) %>%
    mutate(Cluster = 1:4) %>%
    mutate(pred_type = "159 Samples") %>%
    rbind(mutation_enrichment) %>%
    mutate(pred_type = factor(pred_type, levels = c("159 Samples", 
                                                    "210 Samples",
                                                    "51 Samples")))
  
  plot_df1 <- mutation_enrichment %>%
    select(Cluster, `FLT3 positive fraction`,
           `Enrichment FLT3 p-value`, pred_type) %>%
    dplyr::rename(`P-value` = `Enrichment FLT3 p-value`,
                  `Fraction mutated` = `FLT3 positive fraction`,
                  `Type` = pred_type) %>%
    mutate(`Significant enrichment` = case_when(`P-value` < 0.05 ~ 1,
                                   TRUE ~ 0))
  p1 <- ggplot(plot_df1, aes(x = Cluster, y = `Fraction mutated`, 
                             group = `Type`, 
                             fill = `Type`)) + 
    geom_bar(stat = "identity", pos = position_dodge(), width = 0.77, alpha = 0.7) + 
    scale_fill_manual(values = prediction_colors) +
    geom_col_pattern(aes(x = Cluster, y = `Fraction mutated`,
                         pattern_alpha = `Significant enrichment`),
                     width = 0.77,  pos = position_dodge(),
                     pattern = 'stripe',
                     color = 'black') +
    scale_pattern_alpha(guide = "none") + 
    ggtitle("FLT3 Status") + ylim(0, 1)
  
  plot_df2 <- mutation_enrichment %>%
    select(Cluster, `RAS positive fraction`,
           `Enrichment RAS p-value`, pred_type) %>%
    dplyr::rename(`P-value` = `Enrichment RAS p-value`,
                  `Fraction mutated` = `RAS positive fraction`,
                  `Type` = pred_type) %>%
    mutate(`Significant enrichment` = case_when(`P-value` < 0.05 ~ 1,
                                   TRUE ~ 0))
  p2 <- ggplot(plot_df2, aes(x = Cluster, y = `Fraction mutated`, 
                             group = `Type`, 
                             fill = `Type`)) + 
    geom_bar(stat = "identity", pos = position_dodge(), width = 0.77, alpha = 0.7) + 
    scale_fill_manual(values = prediction_colors) +
    geom_col_pattern(aes(x = Cluster, y = `Fraction mutated`,
                         pattern_alpha = `Significant enrichment`),
                     width = 0.77,  pos = position_dodge(),
                     pattern = 'stripe',
                     color = 'black') +
    scale_pattern_alpha(guide = "none") + 
    ggtitle("RAS Status") + ylim(0, 1)
  
  plot_df3 <- mutation_enrichment %>%
    select(Cluster, `NPM1_clinical positive fraction`,
           `Enrichment NPM1_clinical p-value`, pred_type) %>%
    dplyr::rename(`P-value` = `Enrichment NPM1_clinical p-value`,
                  `Fraction mutated` = `NPM1_clinical positive fraction`,
                  `Type` = pred_type) %>%
    mutate(`Significant enrichment` = case_when(`P-value` < 0.05 ~ 1,
                                   TRUE ~ 0))
  p3 <- ggplot(plot_df3, aes(x = Cluster, y = `Fraction mutated`, 
                             group = `Type`, 
                             fill = `Type`)) + 
    geom_bar(stat = "identity", pos = position_dodge(), width = 0.77, alpha = 0.7) + 
    scale_fill_manual(values = prediction_colors) +
    geom_col_pattern(aes(x = Cluster, y = `Fraction mutated`,
                         pattern_alpha = `Significant enrichment`),
                     width = 0.77,  pos = position_dodge(),
                     pattern = 'stripe',
                     color = 'black') +
    scale_pattern_alpha(guide = "none") + 
    ggtitle("NPM1_clinical Status") + ylim(0, 1)
  
  p_all <- arrangeGrob(p1, p2, p3, ncol = 3,
                       top = textGrob(paste0("Mutation Status - ", data_type, " predicions - alpha = ", chosen_alpha), 
                                      gp = gpar(fontsize = 20)))
  
  ggsave(paste0("enet_multinomial_data/predicted_mutation_status_", 
                  data_type, "_", chosen_alpha, path_ending), 
           width = 15, height = 5, p_all)
  
  return(mutation_enrichment)
}

```



```{r eval=FALSE, include=FALSE}
sanity_check <- mut_enrichment_plot(0.7, "Global", TRUE)

```


# Checking consistency of subtypes: Predicting on 210 vs NMF assignment.


```{r eval=FALSE, include=FALSE}
## Uploading 210 subtype prediction to synapse
predictions <- make_prediction(0.7, "Global", TRUE) %>%
  select(Barcode.ID, pred_cluster_v2) %>%
  dplyr::rename(Cluster = pred_cluster_v2) %>%
  mutate(Cluster = as.character(Cluster))

write.table(predictions, "enet_multinomial_data/full_210_prediction_cluster_assignments_k4.txt", sep = "\t")
upload.plot("enet_multinomial_data/full_210_prediction_cluster_assignments_k4.txt", "syn29612385")

```

Checking how the predicted clusters on the 210 patients compare to the NMF assignments.


```{r}
prediction_210 <- read.table(syn$get("syn30030154")$path, sep = "\t") %>%
  mutate(Cluster = paste("Predicted", Cluster))
NMF_159 <- read.table(syn$get("syn26642544")$path, sep = "\t") %>%
  mutate(k.4 = paste("NMF Cluster", k.4))

combined <- left_join(prediction_210, NMF_159, by = "Barcode.ID") %>% 
  select(Barcode.ID, Cluster, k.4)

table(combined$k.4, combined$Cluster)

```


# Predicting on new data.



```{r}
signatures <- read.table("enet_multinomial_data/enet_multinomial_all_signatures.txt", sep = "\t")

## Functions to predict subtypes on new data.
## mat_new should be normalized and include all the features of global_mat_train.
subtype_prediction <- function(mat_new, metadata){
  chosen_alpha = 0.7
  chosen_lambda = 0.0114996
  data_mat <- global_mat_train
  
  cluster_lab <- enet_meta %>%
    mutate(cluster_lab = paste("Cluster", k.4)) %>%
    pull(cluster_lab) %>%
    as.factor()
  names(cluster_lab) <- enet_meta$Barcode.ID
  cluster_lab <- cluster_lab[colnames(data_mat)]
  
  model <- glmnet(x = t(data_mat),
                  y = cluster_lab,
                  family = 'multinomial',
                  alpha = chosen_alpha,
                  lambda = chosen_lambda)
  
  cluster_predictions <- predict(model, s = chosen_lambda, 
                                 newx = t(mat_new), 
                                 type = "response")[, , 1] %>% as.data.frame()
  colnames(cluster_predictions) <- paste("New", colnames(cluster_predictions))
  cluster_predictions$max_new <- apply(cluster_predictions, 1, max) %>% unname()
  cluster_predictions$pred_cluster <- predict(model, s = chosen_lambda, 
                                              newx = t(mat_new), 
                                              type = "class")
  cluster_predictions$SampleName <- colnames(mat_new)
  cluster_predictions <- cluster_predictions %>%
    merge(metadata, by = "SampleName")
  return(cluster_predictions)
}

sig_heatmap <- function(data_mat, col_df, file_path, plot_title){
  signatures <- signatures %>%
    group_by(chosen_type, alpha, Gene) %>%
    dplyr::mutate(total = n()) %>%
    ungroup()
  
  data_mat <- data_mat[which(apply(data_mat, 1, sd) != 0), ]
  
  row_df <- signatures %>%
    dplyr::filter(alpha == 0.7,
                  Gene %in% rownames(data_mat)) %>%
    dplyr::mutate(Cluster = paste("Cluster", Cluster)) %>%
    dplyr::filter(chosen_type == "Global") %>%
    dplyr::select(Gene, Cluster) %>%
    group_by(Gene) %>%
    dplyr::mutate(total = n()) %>%
    dplyr::mutate(Cluster = case_when(total == 1 ~ as.character(Cluster),
                                      TRUE ~ "Mixed"),
                  Cluster = factor(Cluster, levels = c("Cluster 1", "Cluster 2", 
                                                       "Cluster 3", "Cluster 4", "Mixed"))) %>%
    unique() %>%
    as.data.frame()
  rownames(row_df) <- row_df$Gene
  
  col_df <- col_df %>%
    dplyr::select(pred_cluster, SampleName, Time, Treatment) %>%
    dplyr::rename(Cluster = pred_cluster) %>%
    dplyr::mutate(Cluster = as.factor(Cluster)) 
  rownames(col_df) <- col_df$SampleName
  col_df <- col_df %>%
    select(Cluster, Time, Treatment)
  
  ann_colors <- list("Cluster" = c("Cluster 1" = subtype_colors[[1]], 
                                   "Cluster 2" = subtype_colors[[2]],
                                   "Cluster 3" = subtype_colors[[3]],
                                   "Cluster 4" = subtype_colors[[4]],
                                   "Mixed" = "grey"))
  
  heatmap_breaks <- seq(-3, 3, by = 0.2)
  heatmap_colors <- colorRampPalette(rev(brewer.pal(n = 10, name = "RdBu")))(length(heatmap_breaks))
  cluster_row = pheatmap:::cluster_mat(data_mat[rownames(row_df), ], distance = "euclidean", 
                                       method = "ward.D")
  row_df <- row_df[cluster_row$order, ] %>%
    mutate(order = 1:n()) %>%
    arrange(Cluster) %>%
    select(Cluster)
  sigs <- rownames(row_df)
  
  make.pheatmap(data_mat[sigs, ], filename = file_path, format = "pdf",
                scale = "none", cluster_rows = FALSE, clustering_method = "ward.D", 
                annotation_col = col_df, annotation_row = row_df, annotation_colors = ann_colors,
                breaks = heatmap_breaks, color = heatmap_colors, show_rownames = F, show_colnames = F,
                main = plot_title, height = 7, width = 7)
}


```


### Predicting on MOLM14 and HL60 cell line data.


```{r}

## Preparing the cell line data.
HL60 <- querySynapseTable("syn25813233")
MOLM14 <- querySynapseTable("syn25813133")

signature_proteins <- read.table(syn$get("syn30029402")$path, sep = "\t") %>%
  filter(data_type == "Global") %>%
  pull(Gene) %>% unique()

metadata_HL60 <- HL60 %>%
  select(SampleName, Treatment, Time, LoadingMass, Pair, Plex) %>%
  unique()
rownames(metadata_HL60) <- metadata_HL60$SampleName
m_HL60 <- make.msnset(HL60, feature.col = "Gene", 
                      sample.col = "SampleName", 
                      value.col = "LogRatio", metadata = metadata_HL60)

metadata_MOLM14 <- MOLM14 %>%
  select(SampleName, Treatment, Time, LoadingMass, Pair, Plex) %>%
  unique()
rownames(metadata_MOLM14) <- metadata_MOLM14$SampleName
m_MOLM14 <- make.msnset(MOLM14, feature.col = "Gene", 
                        sample.col = "SampleName", 
                        value.col = "LogRatio", metadata = metadata_MOLM14)

mat_HL60 <- exprs(m_HL60)
mat_MOLM14 <- exprs(m_MOLM14)

mat_HL60 <- sweep(mat_HL60, 1, apply(mat_HL60, 1, mean, na.rm = T), FUN = '-')
mat_HL60 <- sweep(mat_HL60, 1, apply(mat_HL60, 1, sd, na.rm = T), FUN = '/')
mat_MOLM14 <- sweep(mat_MOLM14, 1, apply(mat_MOLM14, 1, mean, na.rm = T), FUN = '-')
mat_MOLM14 <- sweep(mat_MOLM14, 1, apply(mat_MOLM14, 1, sd, na.rm = T), FUN = '/')

mat_HL60 <- mat_HL60[rownames(mat_HL60) %in% rownames(global_mat_train), ]
dummy_rows <- setdiff(rownames(global_mat_train), rownames(mat_HL60))
dummy_mat <- matrix(0, ncol = ncol(mat_HL60), nrow = length(dummy_rows))
rownames(dummy_mat) <- dummy_rows
colnames(dummy_mat) <- colnames(mat_HL60)
mat_HL60 <- rbind(mat_HL60, dummy_mat)

mat_MOLM14 <- mat_MOLM14[rownames(mat_MOLM14) %in% rownames(global_mat_train), ]
dummy_rows <- setdiff(rownames(global_mat_train), rownames(mat_MOLM14))
dummy_mat <- matrix(0, ncol = ncol(mat_MOLM14), nrow = length(dummy_rows))
rownames(dummy_mat) <- dummy_rows
colnames(dummy_mat) <- colnames(mat_MOLM14)
mat_MOLM14 <- rbind(mat_MOLM14, dummy_mat)


```




```{r}

cluster_MOLM14 <- subtype_prediction(mat_MOLM14, metadata_MOLM14)
cluster_HL60 <- subtype_prediction(mat_HL60, metadata_HL60)

sig_heatmap(mat_MOLM14, cluster_MOLM14, "enet_multinomial_data/predicted_subtypes_MOLM14", "MOLM14")
sig_heatmap(mat_HL60, cluster_HL60, "enet_multinomial_data/predicted_subtypes_HL60", "HL60")


```




### Predicting on the Pilot Data.



```{r}
## Loading pilot data using scripts from the BEATAMLPilot github
orig.data<-querySynapseTable('syn22172602')

orig.data <- orig.data %>% dplyr::rename(proteinLevels ='LogFoldChange') %>%
  dplyr::rename(mRNALevels = 'transcriptCounts') %>%
  dplyr::rename(geneMutations = 'Tumor VAF') %>%
  mutate(Gene = unlist(Gene)) %>% rowwise() %>%
  mutate(binaryMutations = ifelse(geneMutations == 0, 0, 1))

pat.data <- querySynapseTable("syn22314121") %>%
  mutate(Gene = unlist(Gene)) %>%
  subset(Treatment == 'Vehicle') %>%
  subset(`Cell number` >= 10000000) %>%
  dplyr::select(Gene, LogFoldChange, `AML sample`)%>%
  distinct() %>%
  full_join(orig.data, by = c('AML sample','Gene')) %>%
  rowwise() %>%
  mutate(proteinLevels = max(proteinLevels, LogFoldChange, na.rm = T)) %>%
  dplyr::select(-LogFoldChange) %>%
  mutate(mRNALevels = tidyr::replace_na(mRNALevels, 0)) %>%
  mutate(geneMutations = tidyr::replace_na(geneMutations, 0)) %>%
  mutate(binaryMutations = tidyr::replace_na(binaryMutations, 0)) %>%
  dplyr::select(-countMetric) %>%
  distinct()

pats.with.prot <- pat.data %>%
  group_by(`AML sample`) %>%
  summarize(hasProt = all(proteinLevels == 0)) %>%
  subset(hasProt == FALSE) %>%
  dplyr::select('AML sample')

pilot_proteomics <- pat.data %>% subset(`AML sample` %in% pats.with.prot$`AML sample`) %>%
  filter(proteinLevels != 0) %>%
  dplyr::rename(Barcode.ID = `AML sample`) %>%
  dplyr::group_by(Barcode.ID, Gene) %>%
  mutate(LogFC = mean(proteinLevels)) %>%
  select(Gene, Barcode.ID, LogFC) %>%
  unique()
pilot_samples <- pilot_proteomics %>%
  pull(Barcode.ID) %>% unique() %>% sort()


# pilot_samples[which(!(pilot_samples %in% metadata_pilot$Barcode.ID))]
# "16-01253" "16-01259" "17-00638" "CB-1" "CB-2"
metadata_pilot <- openxlsx::read.xlsx(syn$get("syn23538805")$path, "wv1to4") %>%
  dplyr::rename(Barcode.ID = labId) %>%
  select(Barcode.ID, overallSurvival, vitalStatus, causeOfDeath) %>%
  filter(Barcode.ID %in% pilot_samples) %>%
  right_join(data.frame(Barcode.ID = pilot_samples, dummy = "TRUE"), by = "Barcode.ID") %>%
  mutate(SampleName = Barcode.ID)
rownames(metadata_pilot) <- metadata_pilot$Barcode.ID

m_pilot <- make.msnset(pilot_proteomics, feature.col = "Gene", value.col = "LogFC",
                       sample.col = "Barcode.ID", metadata = metadata_pilot)

mat_pilot <- exprs(m_pilot)
## Centering and imputing NA with mean, which is zero.
mat_pilot <- sweep(mat_pilot, 1, apply(mat_pilot, 1, mean, na.rm = T), FUN = '-')
mat_pilot[is.na(mat_pilot)] <- 0
## Dividing by sd. Note a few NA's are introduced by this from 0/0. Happens to a given
## Gene if that gene has only one sample in pilot_proteomics.
mat_pilot <- sweep(mat_pilot, 1, apply(mat_pilot, 1, sd, na.rm = T), FUN = '/')
mat_pilot[is.na(mat_pilot)] <- 0

mat_pilot <- mat_pilot[rownames(mat_pilot) %in% rownames(global_mat_train), ]
dummy_rows <- setdiff(rownames(global_mat_train), rownames(mat_pilot))
dummy_mat <- matrix(0, ncol = ncol(mat_pilot), nrow = length(dummy_rows))
rownames(dummy_mat) <- dummy_rows
colnames(dummy_mat) <- colnames(mat_pilot)

mat_pilot <- rbind(mat_pilot, dummy_mat)
mat_pilot <- mat_pilot[rownames(global_mat_train), ]

```


```{r}
pilot_subtypes <- subtype_prediction(mat_pilot, metadata_pilot) %>%
  dplyr::mutate(overallSurvival = as.numeric(overallSurvival),
                vitalStatus = case_when(vitalStatus == "Dead" ~ 1,
                                        TRUE ~ 0))

sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ pred_cluster, data = pilot_subtypes)
p2 <- ggsurvplot(sfit, color = "pred_cluster", data = pilot_subtypes,
                 palette = subtype_colors)$plot +
        theme(legend.text = element_text(size = 12, color = "black"),
              legend.title = element_text(size = 12, color = "black"))

sig_heatmap(mat_pilot, pilot_subtypes, "enet_multinomial_data/pilot_subtypes", "Pilot")
```










