---
title: "K = 4 multinomial predictions"
author: "Camilo Posso"
date: "04/27/2022"
output: 
  html_document:
    code_folding: hide
    toc: true
editor_options: 
  chunk_output_type: inline
---

## Goal

Predict subtypes on the 51 held out samples, as well as entirely new datasets. CHeck
survival, mutation, etc.


```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(glmnet)
library(survminer)
library(survival)
library(gridExtra)
library(grid)
library(ggpattern)

source("../NMF_helper.R")
source("../cluster_model_helper.R")

```


Loading model results.


```{r}
all_results <- readRDS("enet_multinomial_data/elastic_net_multinomial_results_k_4.RDS")
all_bounds <- readRDS("enet_multinomial_data/all_bounds.RDS")
bound_mult <- all_bounds$bound_mult

```



# Predicting subtypes on held-out samples from 210 cohort.

Here we train a prediction model using the full dataset and the best performing hyperparameters
to extend cluster 4 (good prognosis) to the 51 unclustered samples.


```{r}
## Function to make predictions and plots. 
make_prediction <- function(chosen_alpha, data_type, robust){
  if (data_type == "Global"){
    data_mat <- global_mat_train
    data_mat_test <- global_mat_test
    data_mat_full <- global.data %>%
      dplyr::rename(feature = Gene) %>%
      filter(feature %in% global_features) %>%
      pivot_wider(names_from = Barcode.ID, 
                  values_from = LogRatio) %>% 
      as.data.frame()
    rownames(data_mat_full) <- data_mat_full$feature
    data_mat_full <- data_mat_full[, -1] %>% as.matrix()
    
  } else if (data_type == "Phospho"){
    data_mat <- phospho_mat_train
    data_mat_test <- phospho_mat_test
    data_mat_full <- phospho.data %>%
      select(Barcode.ID, SiteID, LogRatio) %>%
      dplyr::rename(feature = SiteID) %>%
      filter(feature %in% phospho_features) %>%
      pivot_wider(names_from = Barcode.ID, 
                  values_from = LogRatio) %>% 
      as.data.frame()
    rownames(data_mat_full) <- data_mat_full$feature
    data_mat_full <- data_mat_full[, -1] %>% as.matrix()
    
  } else if (data_type == "Combined"){
    data_mat <- combined_mat_train
    data_mat_test <- combined_mat_test
    data_mat_full <- global.data %>%
      dplyr::rename(feature = Gene) %>%
      filter(feature %in% global_features) %>%
      pivot_wider(names_from = Barcode.ID, 
                  values_from = LogRatio) %>% 
      as.data.frame()
    rownames(data_mat_full) <- data_mat_full$feature
    data_mat_full <- data_mat_full[, -1] %>% as.matrix()
    
    data_mat_full <- rbind(data_mat_full[, colnames(phospho_imp)], phospho_imp)
  } else if (data_type == "Phospho_imp"){
    data_mat <- phospho_mat_train_imp
    data_mat_test <- phospho_mat_test_imp
    data_mat_full <- phospho_imp
  } else if (data_type == "Combined_v2"){
    data_mat <- combined_v2_mat_train
    data_mat_test <- combined_v2_mat_test
    data_mat_full_g <- global.data %>%
      dplyr::rename(feature = Gene) %>%
      filter(feature %in% global_features) %>%
      pivot_wider(names_from = Barcode.ID, 
                  values_from = LogRatio) %>% 
      as.data.frame()
    rownames(data_mat_full_g) <- data_mat_full_g$feature
    data_mat_full_g <- data_mat_full_g[, -1] %>% as.matrix()
    
    data_mat_full_p <- phospho.data %>%
      select(Barcode.ID, SiteID, LogRatio) %>%
      dplyr::rename(feature = SiteID) %>%
      filter(feature %in% phospho_features) %>%
      pivot_wider(names_from = Barcode.ID, 
                  values_from = LogRatio) %>% 
      as.data.frame()
    rownames(data_mat_full_p) <- data_mat_full_p$feature
    data_mat_full_p <- data_mat_full_p[, -1] %>% as.matrix()
    
    data_mat_full <- rbind(data_mat_full_g,
                           data_mat_full_p[, colnames(data_mat_full_g)])
  }
  
  xx <- all_results[[data_type]]
  data_mat_full <- sweep(data_mat_full, 1, apply(data_mat_full, 1, mean), FUN = '-')
  data_mat_full <- sweep(data_mat_full, 1, apply(data_mat_full, 1, sd), FUN = '/')
  
  ## Key line
  if (robust){
    data_mat_test <- data_mat_full[, colnames(data_mat_test)]
  }
  
  min_bounds <- all_bounds[[paste0("min_bounds_", data_type)]]
  max_bounds <- all_bounds[[paste0("max_bounds_", data_type)]]
  min_lambda <- min_bounds[as.character(chosen_alpha), ]
  max_lambda <- max_bounds[as.character(chosen_alpha), ]
  allowed_lambda <- (as.numeric(rownames(xx)) < max_lambda) & 
                    (as.numeric(rownames(xx)) > min_lambda)
  
  xx <- xx[allowed_lambda, ]
  col_name <- paste("alpha =", chosen_alpha)
  chosen_lambda <- rownames(xx)[[which.min(xx[[col_name]])]] %>%
    as.numeric()
  print(chosen_lambda)
  
  cluster_lab <- enet_meta %>%
    mutate(cluster_lab = paste("Cluster", k.4)) %>%
    pull(cluster_lab) %>%
    as.factor()
  names(cluster_lab) <- enet_meta$Barcode.ID
  cluster_lab <- cluster_lab[colnames(data_mat)]

  model <- glmnet(x = t(data_mat),
                  y = cluster_lab,
                  family = 'multinomial',
                  alpha = chosen_alpha,
                  lambda = chosen_lambda)
  
  cluster_predictions <- predict(model, s = chosen_lambda, 
                                 newx = t(data_mat_test), 
                                 type = "response")[, , 1] %>% as.data.frame()
  colnames(cluster_predictions) <- paste("New", colnames(cluster_predictions))
  cluster_predictions$max_new <- apply(cluster_predictions, 1, max) %>% unname()
  cluster_predictions$pred_cluster <- predict(model, s = chosen_lambda, 
                                              newx = t(data_mat_test), 
                                              type = "class")
  cluster_predictions$Barcode.ID <- rownames(cluster_predictions)
  
  cluster_predictions_full <- predict(model, s = chosen_lambda, 
                                      newx = t(data_mat_full), 
                                      type = "response")[, , 1] %>% as.data.frame()
  colnames(cluster_predictions_full) <- paste("Full", colnames(cluster_predictions_full))
  cluster_predictions_full$max_full <- apply(cluster_predictions_full, 1, max) %>% unname()
  cluster_predictions_full$pred_cluster_v2 <- predict(model, s = chosen_lambda, 
                                                      newx = t(data_mat_full), 
                                                      type = "class")
  cluster_predictions_full$Barcode.ID <- rownames(cluster_predictions_full)
  
  out <- full_join(cluster_predictions, cluster_predictions_full, by = "Barcode.ID")
  return(out)
}


make_prediction_plot <- function(chosen_alpha, data_type, confidence_cutoff = 0.40, robust = TRUE){
  cluster_predictions <- make_prediction(chosen_alpha, data_type, robust) %>%
    mutate(pred_cluster = sub("Cluster ", "", pred_cluster),
           pred_cluster_v2 = sub("Cluster ", "", pred_cluster_v2))
  if (robust){
    path_ending <- "_robust_normalization.pdf"
  } else {
    path_ending <- ".pdf"
  }
  
  meta_c <- left_join(meta, cluster_assignments, by = "Barcode.ID") %>%
    mutate(overallSurvival = as.numeric(overallSurvival),
           k.4 = as.character(k.4)) %>%
    left_join(cluster_predictions, by = "Barcode.ID")
  rownames(meta_c) <- meta_c$Barcode.ID
  
  meta_c_new <- meta_c %>%
      filter(is.na(k.4)) %>%
    mutate(conf = case_when(max_new <= confidence_cutoff ~ "Low confidence",
                            TRUE ~ "OK")) %>%
    dplyr::rename(Cluster = pred_cluster)
  meta_c_plot <- meta_c_new %>%
    filter(conf == "OK")
  meta_c_train <- meta_c %>%
    filter(!is.na(k.4)) %>%
    dplyr::rename(Cluster = k.4)
  
  plot_title <- paste0("Predicting on 51 new samples, conf > ", 
                       confidence_cutoff, "(", nrow(meta_c_plot), " samples)")
  color_palette <- c("1" = subtype_colors[[1]],
                     "2" = subtype_colors[[2]],
                     "3" = subtype_colors[[3]],
                     "4" = subtype_colors[[4]])
  
  sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ Cluster, data = meta_c_train)
  p1 <- ggsurvplot(sfit, color = "Cluster", data = meta_c_train,
                   palette = color_palette, pval = TRUE)$plot +
          theme(legend.text = element_text(size = 12, color = "black"),
                legend.title = element_text(size = 12, color = "black")) + 
          ggtitle("159 Clustered samples", data_type)
  
  meta_c$Cluster <- meta_c$pred_cluster_v2
  sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ Cluster, data = meta_c)
  p2 <- ggsurvplot(sfit, color = "Cluster", data = meta_c,
                   palette = color_palette, pval = TRUE)$plot +
          theme(legend.text = element_text(size = 12, color = "black"),
                legend.title = element_text(size = 12, color = "black")) + 
          ggtitle(paste("Predicting on all 210 samples", data_type))
  
  sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ Cluster, data = meta_c_plot)
  p3 <- ggsurvplot(sfit, color = "Cluster", data = meta_c_plot,
                   palette = color_palette, pval = TRUE)$plot +
          theme(legend.text = element_text(size = 12, color = "black"),
                legend.title = element_text(size = 12, color = "black")) +
          ggtitle(plot_title)
  
  sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ Cluster, data = meta_c_new)
  p4 <- ggsurvplot(sfit, color = "Cluster", data = meta_c_new,
                   palette = color_palette,  pval = TRUE)$plot +
          theme(legend.text = element_text(size = 12, color = "black"),
                legend.title = element_text(size = 12, color = "black")) + 
          ggtitle(paste("Predicting on 51 new samples", data_type))
  
  p_all <- arrangeGrob(p1, p2, p4, ncol = 2, nrow = 2,
                       top = textGrob(paste0("Survival plots, ", data_type, ", alpha = ", chosen_alpha), 
                                      gp = gpar(fontsize = 20)))
  
  save_path <- paste0("enet_multinomial_data/predicted_survival_", 
                data_type, "_", chosen_alpha, path_ending)
  ggsave(save_path, width = 16, height = 12, p_all)
  return(list(p1, p2, p4))
}

```



```{r eval=FALSE, include=FALSE}
## Making survival plots by predicted subtype
lapply(c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), function(chosen_alpha){
  make_prediction_plot(chosen_alpha, "Global", robust = TRUE)
  make_prediction_plot(chosen_alpha, "Phospho_imp", robust = TRUE)
  make_prediction_plot(chosen_alpha, "Combined", robust = TRUE)
  make_prediction_plot(chosen_alpha, "Combined_v2", robust = TRUE)
  make_prediction_plot(chosen_alpha, "Global", robust = FALSE)
  make_prediction_plot(chosen_alpha, "Phospho_imp", robust = FALSE)
  make_prediction_plot(chosen_alpha, "Combined", robust = FALSE)
  make_prediction_plot(chosen_alpha, "Combined_v2", robust = FALSE)
})

prediction_survival_plots <- make_prediction_plot(0.9, "Global", robust = TRUE)
ggsave("enet_multinomial_data/global_prediction_survival_alpha_0.9_all_samples.png",
       width = 7, height = 7, plot = prediction_survival_plots[[2]])
ggsave("enet_multinomial_data/global_prediction_survival_alpha_0.9_heldout_samples.png",
       width = 7, height = 7, plot = prediction_survival_plots[[3]])

prediction_survival_plots <- make_prediction_plot(0.9, "Combined_v2", robust = TRUE)
ggsave("enet_multinomial_data/combined_v2_prediction_survival_alpha_0.9_all_samples.png",
       width = 7, height = 7, plot = prediction_survival_plots[[2]])
ggsave("enet_multinomial_data/combined_v2_prediction_survival_alpha_0.9_heldout_samples.png",
       width = 7, height = 7, plot = prediction_survival_plots[[3]])


prediction_survival_plots <- make_prediction_plot(0.9, "Global", robust = TRUE)
ggsave("enet_multinomial_data/global_prediction_survival_alpha_0.9_all_samples.pdf",
       width = 7, height = 7, plot = prediction_survival_plots[[2]])
ggsave("enet_multinomial_data/global_prediction_survival_alpha_0.9_heldout_samples.pdf",
       width = 7, height = 7, plot = prediction_survival_plots[[3]])

prediction_survival_plots <- make_prediction_plot(0.9, "Combined_v2", robust = TRUE)
ggsave("enet_multinomial_data/combined_v2_prediction_survival_alpha_0.9_all_samples.pdf",
       width = 7, height = 7, plot = prediction_survival_plots[[2]])
ggsave("enet_multinomial_data/combined_v2_prediction_survival_alpha_0.9_heldout_samples.pdf",
       width = 7, height = 7, plot = prediction_survival_plots[[3]])


```



### Mutation status by subtype



```{r}
## Computing enrichment status of each mutation within the subtypes.
all_mutation_data <- load_mutational_sample_data()
mutation_data <- all_mutation_data %>%
  mutate(value = TRUE) %>%
  pivot_wider(names_from = Gene, values_from = value, values_fn = any)
mutation_data <- mutation_data[, c("Barcode.ID", "KRAS+NRAS", "NPM1_clinical")] %>%
  full_join(meta[, c("Barcode.ID", "FLT3.ITD")], by = "Barcode.ID") %>%
  mutate(RAS = case_when(!is.na(`KRAS+NRAS`) ~ `KRAS+NRAS`,
                          TRUE ~ FALSE),
         NPM1_clinical = case_when(!is.na(NPM1_clinical) ~ NPM1_clinical,
                                   TRUE ~ FALSE),
         FLT3 = case_when(FLT3.ITD == "TRUE" ~ TRUE,
                          TRUE ~ FALSE)) %>%
  select(Barcode.ID, FLT3, NPM1_clinical, RAS)

rownames(mutation_data) <- mutation_data$Barcode.ID

mut_enrichment <- function(sample_categories, assignments, log.scale = FALSE) {
  k = length(unique(assignments$Cluster))
  
  sample_categories <- sample_categories %>%
    filter(rownames(.) %in% rownames(assignments)) %>%
    left_join(assignments, by = "Barcode.ID") %>%
    column_to_rownames("Barcode.ID") %>% 
    as.data.frame()
  
  categories <- sample_categories %>%
    select(-Cluster) %>%
    colnames()
  
  out <- lapply(categories, function(category){
    
    p.values <- lapply(1:k, function(i){
      cluster.df <- sample_categories %>%
        mutate(Cluster = case_when(Cluster == i ~ TRUE,
                                   TRUE ~ FALSE)) %>%
        mutate(Cluster = as.factor(Cluster))
      dat <- table(cluster.df[[category]], cluster.df$Cluster)
      background_dist <- apply(dat, 1, sum)
      background_prop <- background_dist[[2]]/(background_dist[[2]] + background_dist[[1]])
      cluster_prop <- dat[2, 2]/(dat[1, 2] + dat[2, 2])
      p_value <- fisher.test(dat, workspace = 2e8, alternative = "greater")$p.value
      if (log.scale) {
        p_value <- -log10(p_value)
      } else {
        p_value <- p_value
      }
      xx <- data.frame(background_prop, cluster_prop, p_value)
      colnames(xx) <- c(paste("Background", category, "positive fraction"),
                        paste(category, "positive fraction"),
                        paste("Enrichment", category, "p-value"))
      rownames(xx) <- paste("Cluster", i)
      return(xx)
    }) %>% do.call("rbind", .)
    
  }) %>% do.call(cbind, .) %>% 
    as.data.frame()
  return(out)
}


mut_enrichment_plot <- function(chosen_alpha, data_type, robust = TRUE){
  if (robust){
    path_ending <- "_robust_normalization.pdf"
  } else {
    path_ending <- ".pdf"
  }
  
  meta_pred <- make_prediction(chosen_alpha, data_type, robust) %>%
    select(Barcode.ID, pred_cluster, pred_cluster_v2) %>%
    full_join(mutation_data, by = "Barcode.ID") %>%
    mutate(pred_cluster = sub("Cluster ", "", pred_cluster),
           pred_cluster_v2 = sub("Cluster ", "", pred_cluster_v2))
  rownames(meta_pred) <- meta_pred$Barcode.ID
  
  sample_assn <- meta_pred[, c("Barcode.ID", "pred_cluster")] %>%
    dplyr::rename(Cluster = pred_cluster) %>%
    filter(!is.na(Cluster))
  sample_cats <- meta_pred[rownames(sample_assn), c("Barcode.ID", "FLT3", "RAS", "NPM1_clinical")]
  mutation_enrichment <- mut_enrichment(sample_cats, sample_assn) %>%
    mutate(Cluster = 1:4) %>%
    mutate(pred_type = "51 Samples")
  
  sample_assn <- meta_pred[, c("Barcode.ID", "pred_cluster_v2")] %>%
    dplyr::rename(Cluster = pred_cluster_v2) %>%
    filter(!is.na(Cluster))
  sample_cats <- meta_pred[rownames(sample_assn), c("Barcode.ID", "FLT3", "RAS", "NPM1_clinical")]
  mutation_enrichment <- mut_enrichment(sample_cats, sample_assn) %>%
    mutate(Cluster = 1:4) %>%
    mutate(pred_type = "210 Samples") %>%
    rbind(mutation_enrichment)
  
  sample_assn <- enet_meta[, c("Barcode.ID", "k.4")] %>%
    dplyr::rename(Cluster = k.4) %>%
    filter(!is.na(Cluster)) 
  sample_cats <- meta_pred[rownames(sample_assn), c("Barcode.ID", "FLT3", "RAS", "NPM1_clinical")]
  mutation_enrichment <- mut_enrichment(sample_cats, sample_assn) %>%
    mutate(Cluster = 1:4) %>%
    mutate(pred_type = "159 Samples") %>%
    rbind(mutation_enrichment) %>%
    mutate(pred_type = factor(pred_type, levels = c("159 Samples", 
                                                    "210 Samples",
                                                    "51 Samples")))
  
  plot_df1 <- mutation_enrichment %>%
    select(Cluster, `FLT3 positive fraction`,
           `Enrichment FLT3 p-value`, pred_type) %>%
    dplyr::rename(`P-value` = `Enrichment FLT3 p-value`,
                  `Fraction mutated` = `FLT3 positive fraction`,
                  `Type` = pred_type) %>%
    mutate(`Significant enrichment` = case_when(`P-value` < 0.05 ~ 1,
                                   TRUE ~ 0))
  p1 <- ggplot(plot_df1, aes(x = Cluster, y = `Fraction mutated`, 
                             group = `Type`, 
                             fill = `Type`)) + 
    geom_bar(stat = "identity", pos = position_dodge(), width = 0.77, alpha = 0.7) + 
    scale_fill_manual(values = prediction_colors) +
    geom_col_pattern(aes(x = Cluster, y = `Fraction mutated`,
                         pattern_alpha = `Significant enrichment`),
                     width = 0.77,  pos = position_dodge(),
                     pattern = 'stripe',
                     color = 'black') +
    scale_pattern_alpha(guide = "none") + 
    ggtitle("FLT3 Status") + ylim(0, 1)
  
  plot_df2 <- mutation_enrichment %>%
    select(Cluster, `RAS positive fraction`,
           `Enrichment RAS p-value`, pred_type) %>%
    dplyr::rename(`P-value` = `Enrichment RAS p-value`,
                  `Fraction mutated` = `RAS positive fraction`,
                  `Type` = pred_type) %>%
    mutate(`Significant enrichment` = case_when(`P-value` < 0.05 ~ 1,
                                   TRUE ~ 0))
  p2 <- ggplot(plot_df2, aes(x = Cluster, y = `Fraction mutated`, 
                             group = `Type`, 
                             fill = `Type`)) + 
    geom_bar(stat = "identity", pos = position_dodge(), width = 0.77, alpha = 0.7) + 
    scale_fill_manual(values = prediction_colors) +
    geom_col_pattern(aes(x = Cluster, y = `Fraction mutated`,
                         pattern_alpha = `Significant enrichment`),
                     width = 0.77,  pos = position_dodge(),
                     pattern = 'stripe',
                     color = 'black') +
    scale_pattern_alpha(guide = "none") + 
    ggtitle("RAS Status") + ylim(0, 1)
  
  plot_df3 <- mutation_enrichment %>%
    select(Cluster, `NPM1_clinical positive fraction`,
           `Enrichment NPM1_clinical p-value`, pred_type) %>%
    dplyr::rename(`P-value` = `Enrichment NPM1_clinical p-value`,
                  `Fraction mutated` = `NPM1_clinical positive fraction`,
                  `Type` = pred_type) %>%
    mutate(`Significant enrichment` = case_when(`P-value` < 0.05 ~ 1,
                                   TRUE ~ 0))
  p3 <- ggplot(plot_df3, aes(x = Cluster, y = `Fraction mutated`, 
                             group = `Type`, 
                             fill = `Type`)) + 
    geom_bar(stat = "identity", pos = position_dodge(), width = 0.77, alpha = 0.7) + 
    scale_fill_manual(values = prediction_colors) +
    geom_col_pattern(aes(x = Cluster, y = `Fraction mutated`,
                         pattern_alpha = `Significant enrichment`),
                     width = 0.77,  pos = position_dodge(),
                     pattern = 'stripe',
                     color = 'black') +
    scale_pattern_alpha(guide = "none") + 
    ggtitle("NPM1_clinical Status") + ylim(0, 1)
  
  p_all <- arrangeGrob(p1, p2, p3, ncol = 3,
                       top = textGrob(paste0("Mutation Status - ", data_type, " predicions - alpha = ", chosen_alpha), 
                                      gp = gpar(fontsize = 20)))
  
  ggsave(paste0("enet_multinomial_data/predicted_mutation_status_", 
                  data_type, "_", chosen_alpha, path_ending), 
           width = 15, height = 5, p_all)
  
  return(mutation_enrichment)
}

```


## Uploading table with subtype probabilities

```{r}
predictions_df <- make_prediction(0.9, "Combined_v2", TRUE) %>%
  select(Barcode.ID, 
         `Cluster 1` = `Full Cluster 1`, 
         `Cluster 2` = `Full Cluster 2`, 
         `Cluster 3` = `Full Cluster 3`, 
         `Cluster 4` = `Full Cluster 4`, 
         max = max_full, pred_cluster_v2) %>%
  dplyr::rename(Cluster_global_phospho = pred_cluster_v2) %>%
  mutate(data_type = "Global + Phospho")
write.table(predictions_df, "enet_multinomial_data/full_210_subtyping_probabilities_global_phospho.txt", 
            sep = "\t", row.names = F, quote = F)

synapseStore("enet_multinomial_data/full_210_subtyping_probabilities_global_phospho.txt", "syn29612385")

```


```{r}

prediction_210 <- read.table(syn$get("syn30030154")$path, sep = "\t") %>%
  mutate(Cluster = sub("Cluster ", "", Cluster))
rownames(prediction_210) <- prediction_210$Barcode.ID
NMF_samples <- enet_meta %>% pull(Barcode.ID)
NMF_assn <- prediction_210 %>%
  filter(Barcode.ID %in% NMF_samples)


mutation_enrichment <- mut_enrichment(mutation_data, prediction_210, log.scale = F)
write.table(mutation_enrichment, "mutation_enrichment_210.txt", sep = "\t", quote = F)
mutation_enrichment_NMF <- mut_enrichment(mutation_data, NMF_assn, log.scale = F)
write.table(mutation_enrichment_NMF, "mutation_enrichment_159.txt", sep = "\t", quote = F)

```


```{r eval=FALSE, include=FALSE}
sanity_check <- mut_enrichment_plot(0.7, "Global", TRUE)

```

# Extensions by data type.

```{r eval=FALSE, include=FALSE}
predictions_1 <- make_prediction(0.7, "Global", TRUE) %>%
  select(Barcode.ID, pred_cluster_v2) %>%
  dplyr::rename(Cluster_global = pred_cluster_v2) %>%
  mutate(data_type = "Global")
write.table(predictions_1, "enet_multinomial_data/full_210_prediction_cluster_assignments_global.txt", sep = "\t")

predictions_2 <- make_prediction(0.9, "Combined_v2", TRUE) %>%
  select(Barcode.ID, pred_cluster_v2) %>%
  dplyr::rename(Cluster_global_phospho = pred_cluster_v2) %>%
  mutate(data_type = "Global + Phospho")
write.table(predictions_2, "enet_multinomial_data/full_210_prediction_cluster_assignments_global_phospho.txt", sep = "\t")

predictions_3 <- make_prediction(0.9, "Combined", TRUE) %>%
  select(Barcode.ID, pred_cluster_v2) %>%
  dplyr::rename(Cluster_global_phospho_imp = pred_cluster_v2) %>%
  mutate(data_type = "Global + Phospho imputed")
write.table(predictions_3, "enet_multinomial_data/full_210_prediction_cluster_assignments_global_phospho_imp.txt", sep = "\t")

predictions_combined <- merge(predictions_1, predictions_2, by = "Barcode.ID") %>%
  merge(predictions_3, by = "Barcode.ID") %>%
  mutate(NMF = case_when(Barcode.ID %in% cluster_assignments$Barcode.ID ~ TRUE,
                         TRUE ~ FALSE))

predictions_combined_new <- predictions_combined %>%
  filter(NMF == TRUE)
table(predictions_combined_new$Cluster_global_phospho_imp, predictions_combined_new$Cluster_global_phospho)
table(predictions_combined$Cluster_global_phospho_imp, predictions_combined$Cluster_global_phospho)
table(predictions_combined$Cluster_global_phospho == predictions_combined$Cluster_global_phospho_imp)


```



```{r eval=FALSE, include=FALSE}
## Combines all the best models for each alpha.
data_type <- "Global"
predictions <- make_prediction(0.1, data_type, TRUE) %>%
  select(Barcode.ID, pred_cluster_v2) %>%
  dplyr::rename(!!paste("Cluster", 0.1) := pred_cluster_v2)
predictions_prob <- make_prediction(0.1, data_type, TRUE) 
for(alpha in c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)){
    predictions <- make_prediction(alpha, data_type, TRUE) %>%
      select(Barcode.ID, pred_cluster_v2) %>%
      dplyr::rename(!!paste("Cluster", alpha) := pred_cluster_v2) %>%
      left_join(predictions, by = "Barcode.ID")
    predictions_prob <-  make_prediction(alpha, data_type, TRUE) %>%
      rbind(predictions_prob)
}

predictions_prob_summary <- predictions_prob %>%
  group_by(Barcode.ID) %>%
  mutate(Cluster1 = mean(`Full Cluster 1`),
         Cluster2 = mean(`Full Cluster 2`),
         Cluster3 = mean(`Full Cluster 3`),
         Cluster4 = mean(`Full Cluster 4`)) %>%
  select(Barcode.ID, Cluster1, Cluster2, Cluster3, Cluster4) %>% unique()

predictions_prob_summary$assn <- apply(predictions_prob_summary[, -1], 1, which.max)

predictions_prob_summary <- merge(predictions_prob_summary, meta, by = "Barcode.ID") %>%
  mutate(overallSurvival = as.numeric(overallSurvival)) %>%
  filter(!(Barcode.ID %in% cluster_assignments$Barcode.ID))
sfit <- survfit(Surv(overallSurvival, vitalStatus) ~ assn, data = predictions_prob_summary)
p1 <- ggsurvplot(sfit, color = "assn", data = predictions_prob_summary)$plot

```



# Checking consistency of subtypes: Predicting on 210 vs NMF assignment.



```{r eval=FALSE, include=FALSE}
## Uploading 210 subtype prediction to synapse
predictions <- make_prediction(0.9, "Combined_v2", TRUE) %>%
  select(Barcode.ID, pred_cluster_v2) %>%
  dplyr::rename(Cluster = pred_cluster_v2) %>%
  mutate(Cluster = as.character(Cluster))

write.table(predictions, "enet_multinomial_data/full_210_prediction_cluster_assignments_k4.txt", sep = "\t")
upload.plot("enet_multinomial_data/full_210_prediction_cluster_assignments_k4.txt", "syn29612385")

```


Checking how the predicted clusters on the 210 patients compare to the NMF assignments.


```{r}
prediction_210 <- read.table(syn$get("syn30030154")$path, sep = "\t") %>%
  mutate(Cluster = paste("Predicted", Cluster))
NMF_159 <- read.table(syn$get("syn26642544")$path, sep = "\t") %>%
  mutate(k.4 = paste("NMF Cluster", k.4))

combined <- left_join(prediction_210, NMF_159, by = "Barcode.ID") %>% 
  select(Barcode.ID, Cluster, k.4)

table(combined$k.4, combined$Cluster)

```


