---
title: "210 Patient Summary Figures"
author: "Camilo Posso"
date: "10/04/2021"
output: 
  html_document:
    code_folding: hide
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

```{r libraries, include=FALSE}
library(kableExtra)
library(gridExtra)
library(dplyr)
library(MSnSet.utils)
library(ggplot2)
library(amlresistancenetworks)
library(tidyr)
library(tibble)
library(RColorBrewer)
library(pheatmap)
source("../util/synapseUtil.R")
source("../util/loading_data.R")

global.data <- load.global.data()
global.meta <- global.data$Metadata
global.data <- global.data$`Long-form global`

phospho.data <- load.phospho.data()
phospho.meta <- phospho.data$Metadata
phospho.data <- phospho.data$`Long-form phospho`

functional.data <- load.functional.data()
functional.meta <- functional.data$Metadata %>%
  merge(global.meta[, c("Barcode.ID", "SpecimenType", 
                        "InitialAMLDiagnosis", "PostChemotherapy", 
                        "FLT3.ITD")], by = "Barcode.ID")
functional.data <- functional.data$`Long-form functional` %>%
  merge(functional.meta, by = "Barcode.ID")
rownames(functional.meta) <- functional.meta$Barcode.ID

```

In this short report we collect a few figures exploring the functional,
global, and phospho datasets we have for the large 210 patient experiment.

# Functional Data


```{r missing data, include=FALSE}
functional.mat <- functional.data %>%
  filter(!is.na(auc)) %>%
  group_by(Barcode.ID, Inhibitor) %>%
  slice(1) %>%
  select(Barcode.ID, AUC, Inhibitor) %>%
  pivot_wider(names_from = "Barcode.ID", values_from = "AUC") %>%
  column_to_rownames("Inhibitor")

### Some rows (inhibitors) pairings have no overlap to compute
### distance or correlation for the heatmap. So we omit those here.
### Filtering so that at least ~20 patients overlap when computing the correlation.
idx <- rowSums(is.na(functional.mat)) < 0.4*ncol(functional.mat)
N <- sum(idx)

missing.row <- rowSums(is.na(functional.mat))
missing.col <- colSums(is.na(functional.mat))

dff.row <- data.frame(row = missing.row)
dff.col <- data.frame(col = missing.col)

p1 <- ggplot(dff.row, aes(x = row)) + geom_histogram(binwidth = 7) + xlab("Number Missing") +
  ylab("Count") + ggtitle("Missing by Inhibitor") +
  theme(plot.title = element_text(hjust = 0.5))

p2 <- ggplot(dff.col, aes(x = col)) + geom_histogram(binwidth = 5) + xlab("Number Missing") +
  ylab("Count") + ggtitle("Missing by Sample") +
  theme(plot.title = element_text(hjust = 0.5))
  
```


We have 210 patients and a total of `r nrow(functional.mat)` inhibitors.
As we can see, there's a number of Inhibitors with little missing data, and quite a few
with only a small number of samples present. Likewise, we see that most of the 210 patients
have around ~130 Inhibitors with AUC data present.


```{r functional missing plots, echo=FALSE}
grid.arrange(grobs = list(p1,p2), ncol = 2)

```

Next we plot a histogram of the AUC values themselves (Area under curve), coloring by 
FLT3 status and Specimen Type.

```{r AUC histograms, include=FALSE}
functional.meta <- merge(functional.meta, global.meta)
xx <- functional.data %>%
  filter(!is.na(AUC))

p1 <- ggplot(xx, aes(x = AUC, fill = SpecimenType)) + geom_histogram(bins = 30) +
  ylab("Total") + ggtitle("AUC Histogram by specimen type") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13),
        legend.key.size = unit(0.5,"cm"),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 9))
  
p2 <- ggplot(xx, aes(x = AUC, fill = FLT3.ITD)) + geom_histogram(bins = 30) +
  ylab("Total") + ggtitle("AUC Histogram by FLT3 status") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13),
        legend.key.size = unit(0.5,"cm"),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 9))

```


```{r AUC show histograms}
layout <- as.matrix(c(1,2))
grid.arrange(grobs = list(p1,p2), layout_matrix = layout)

```


Below we have a heatmap of the AUC values. Note that we cluster the columns (samples)
and rows (inhibitors) using correlation. Furthermore, in order to compute the distances
required for clustering, we filter the Inhibitors in such a way that at least
~20 observations are used when computing these distances. We are thus left with
`r N` inhibitors and all 210 patients when making the heatmap.


```{r include=FALSE}
ann.labels <- functional.meta %>%
  select("SpecimenType", "PostChemotherapy", "FLT3.ITD") %>%
  mutate(PostChemotherapy = as.character(PostChemotherapy),
         FLT3.ITD = as.character(FLT3.ITD))
rownames(ann.labels) <- functional.meta$Barcode.ID

ann.colors <- list(SpecimenType = c("Bone Marrow Aspirate" = "dodgerblue2", 
                                    "Leukapheresis" = "darkorange", 
                                    "Peripheral Blood" = "firebrick3"),
                   FLT3.ITD = c("TRUE" = "darkorchid1", "FALSE" = "forestgreen"),
                   PostChemotherapy = c("TRUE" = "cadetblue1", "FALSE" = "darkgoldenrod1"))

functional.mat <- functional.mat[idx, ]

tree_col = as.dist(1-cor(functional.mat, use = "pairwise.complete.obs")) %>%
  hclust(method = "ward.D")
                   
tree_row = as.dist(1-cor(t(functional.mat), use = "pairwise.complete.obs")) %>%
  hclust(method = "ward.D")

```



```{r show AUC heatmap}
pheatmap(functional.mat, scale = "none", annotation_col = ann.labels, annotation_colors = ann.colors,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(32), fontsize = 8,
         clustering_distance_cols = "correlation", clustering_distance_rows = "correlation",
         clustering_method = "ward.D", cluster_rows = tree_row, cluster_cols = tree_col,
         show_rownames = F, show_colnames = F, na_col = "gray69")

```

<!-- Not working, as no complete cases are present for the AUC data.
<!-- Finally, below we show PCA plots of the functional data, coloring by Specimen Type, FLT3.ITD -->
<!-- and Post Chemotherapy status. -->

<!-- ```{r pca plots phospho, include=FALSE} -->
<!-- m <- MSnSet(exprs = as.matrix(functional.mat), pData = functional.meta[colnames(functional.mat), ]) -->
<!-- p1 <- plot_pca_v4(m, "FLT3.ITD", show.ellipse = F) -->
<!-- p2 <- plot_pca_v4(m, "SpecimenType", show.ellipse = F) -->
<!-- p3 <- plot_pca_v4(m, "PostChemotherapy", show.ellipse = F) -->

<!-- ``` -->


<!-- ```{r, fig.width=12} -->
<!-- grid.arrange(grobs = list(p1,p3,p2), ncol = 2) -->

<!-- ``` -->



# Global Data


```{r global missing, include=FALSE}
global.mat <- global.data %>%
  select(Gene, Barcode.ID, LogRatio) %>%
  pivot_wider(names_from = "Barcode.ID", values_from = "LogRatio") %>%
  column_to_rownames("Gene")

missing.row <- rowSums(is.na(global.mat))
missing.col <- colSums(is.na(global.mat))

dff.row <- data.frame(row = missing.row)
dff.col <- data.frame(col = missing.col)

p1 <- ggplot(dff.row, aes(x = row)) + geom_histogram(binwidth = 1) + xlab("Number Missing") +
  ylab("Count") + ggtitle("Missing by Gene") +
  theme(plot.title = element_text(hjust = 0.5))

p2 <- ggplot(dff.col, aes(x = col)) + geom_histogram(binwidth = 1) + xlab("Number Missing") +
  ylab("Count") + ggtitle("Missing by Sample") +
  theme(plot.title = element_text(hjust = 0.5))
  
m <- MSnSet(exprs = as.matrix(global.mat), pData = global.meta[colnames(global.mat), ])
```


Below we see a summary of the distribution of missing values in the global data.


```{r}
grid.arrange(grobs = list(p1,p2), ncol = 2)

```


Finally, we show PCA plots by FLT3.ITD status, Specimen Type and Post Chemotherapy status.

```{r pca plots global, include=FALSE}
p1 <- plot_pca_v4(m, "FLT3.ITD", show.ellipse = F)
p2 <- plot_pca_v4(m, "SpecimenType", show.ellipse = F)
p3 <- plot_pca_v4(m, "PostChemotherapy", show.ellipse = F)

```


```{r, fig.width=12}
grid.arrange(grobs = list(p1,p3,p2), ncol = 2)

```




# Phospho Data


```{r phospho missing, include=FALSE}
phospho.mat <- phospho.data %>%
  select(SiteID, Barcode.ID, LogRatio) %>%
  pivot_wider(names_from = "Barcode.ID", values_from = "LogRatio") %>%
  column_to_rownames("SiteID")

missing.row <- rowSums(is.na(phospho.mat))
missing.col <- colSums(is.na(phospho.mat))

dff.row <- data.frame(row = missing.row)
dff.col <- data.frame(col = missing.col)

p1 <- ggplot(dff.row, aes(x = row)) + geom_histogram(binwidth = 1) + xlab("Number Missing") +
  ylab("Count") + ggtitle("Missing by Gene") +
  theme(plot.title = element_text(hjust = 0.5))

p2 <- ggplot(dff.col, aes(x = col)) + geom_histogram(binwidth = 3) + xlab("Number Missing") +
  ylab("Count") + ggtitle("Missing by Sample") +
  theme(plot.title = element_text(hjust = 0.5))
  
m <- MSnSet(exprs = as.matrix(phospho.mat), pData = phospho.meta[colnames(phospho.mat), ])
```


Below we see a summary of the distribution of missing values in the phospho data.


```{r}
grid.arrange(grobs = list(p1,p2), ncol = 2)

```


Finally, we show PCA plots by FLT3.ITD status, Specimen Type and Post Chemotherapy status.

```{r pca plots phospho, include=FALSE}
p1 <- plot_pca_v4(m, "FLT3.ITD", show.ellipse = F)
p2 <- plot_pca_v4(m, "SpecimenType", show.ellipse = F)
p3 <- plot_pca_v4(m, "PostChemotherapy", show.ellipse = F)

```


```{r, fig.width=12}
grid.arrange(grobs = list(p1,p3,p2), ncol = 2)

```




