---
title: "Build K-fold models for use"
author: "Sara Gosline"
date: "10/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(amlresistancenetworks)
library(dplyr)
library(Matrix)

# If you are loading data from Synapse, source -
#source("../util/synapseUtil.R")
#source("../util/loading_data.R")

# If you are unable to connect to Synapse, source -
source("../util/loading_data_local.R")
```

## Load data
We have created a standard file to load all the data needed for the analysis. We also merge the gene based data and the phosphosite data by Gene and Sample

```{r load data, error=FALSE, message=FALSE, warning=FALSE}
global.data <- load.global.data("//Users//dann850//Desktop//CPTAC_PTRC//BeatAML_210//BeatAMLProteomics.csv")
pat.data <- as.data.frame(global.data$`Long-form global`)
pat.data$Sample <- NULL

phospho.data <- load.phospho.data("//Users//dann850//Desktop//CPTAC_PTRC//BeatAML_210//BeatAMLPhosphoproteomics.csv")
pat.phos <- as.data.frame(phospho.data$`Long-form phospho`)
pat.phos$Sample <- NULL

functional.data <- load.functional.data("//Users//dann850//Desktop//CPTAC_PTRC//BeatAML_210//FullPTCohortFunctionalData.csv")
auc.dat <- as.data.frame(functional.data$`Long-form functional`)

##For testing purposes -
auc.dat <- auc.dat[(auc.dat$Condition == "AST-487") | (auc.dat$Condition == "Alisertib (MLN8237)") | (auc.dat$Condition == "CYT387"), ]

##create combined data frame of gene and phosphosite
pat.data <- pat.data %>%
  left_join(rename(pat.phos, site = 'SiteID', Phosphosite='LogRatio'))

```

## Now we can build the models

We want to build both the LASSO and logistic regression models using the following code. This code will divide the patients into 5 batches and train/test the model holding out 1/5 of the data each time. Then we will have predictions on each of the samples. 

```{r build models, warning=FALSE, message=FALSE, error=FALSE}

all.pats <- intersect(pat.data$`AML sample`,auc.dat$`AML sample`)
library(caret)
folds<-createFolds(all.pats,k=5)
##separate out data?

res<-lapply(folds,function(testpats){
    test.pats<-all.pats[testpats]
    train.pats<-setdiff(all.pats,test.pats)

# print("Phospho preds first")
#  substrate.dat<-pat.phos%>%
#    dplyr::select(`AML sample`='Sample',Gene='site',Phosphosite='LogFoldChange')
  
#          tr.dat<-subset(substrate.dat,`AML sample`%in%train.pats)
#      te.dat<-subset(substrate.dat,`AML sample`%in%test.pats)
      
    # phospho.reg.results<-drugMolRegressionEval(auc.dat,
    #                                              tr.dat,
    #                                            'Phosphosite',
    #                                            dplyr::rename(auc.dat,Sample='AML sample'),
    #                                            dplyr::rename(te.dat,Sample='AML sample'))%>%
    #     do.call(rbind,.)
    #   
    # 
    # phospho.lr.results<-drugMolLogRegEval(auc.dat,tr.dat,'Phosphosite',
    #                                       dplyr::rename(auc.dat,Sample='AML sample'),
    #                                        dplyr::rename(te.dat,Sample='AML sample'))%>%
    #   do.call(rbind,.)
    # 
    # phospho.enet.results<-drugMolRegressionEval(auc.dat,
    #                                            tr.dat,
    #                                            'Phosphosite',
    #                                            dplyr::rename(auc.dat,Sample='AML sample'),
    #                                            dplyr::rename(te.dat,Sample='AML sample'),
    #                                            category='Condition',
    #                                            doEnet=TRUE)%>%
    #   do.call(rbind,.)
    # 
  print('Now getting gene-based preds')
  
  tr.dat<-subset(pat.data,`AML sample`%in%train.pats)
  te.dat<-subset(pat.data,`AML sample`%in%test.pats)
  
  ##This list describes the features to be used  
  eval.list<-list(
    #combo=list(mf=c('Gene','Gene','Gene','site'), fn=c('mRNALevels','proteinLevels','binaryMutations','Phosphosite')),
    #mRNA=c('Gene','mRNALevels'),
    prot=c("Gene","proteinLevels"),
    #mut=c('Gene','binaryMutations'),
    phosph=c('site','Phosphosite'))

    message("Comparing logistic preds")
  log.reg.preds<-purrr::map_df(eval.list,
                             ~ drugMolLogRegEval(auc.dat,
                                                tr.dat,
                                               mol.feature=.x[1],
                                                           mol.feature.name=.x[2],
                                         dplyr::rename(auc.dat,Sample='AML sample'),
                                               dplyr::rename(te.dat,Sample='AML sample'),
                                              category='Condition'))

      
  message('Running lasso')
    ##now train model on AML and eval on depmap data
  reg.preds<-purrr::map_df(eval.list,
                         ~ amlresistancenetworks::drugMolRegressionEval(auc.dat,
                            tr.dat,
                            mol.feature=.x[1],
                            mol.feature.name=.x[2],                                                   
                            dplyr::rename(auc.dat,Sample='AML sample'),
                                             dplyr::rename(te.dat,Sample='AML sample'),
                                                 category='Condition'))
  message("Running elastic net")
  enet.reg.preds<-purrr::map_df(eval.list,
                         ~ drugMolRegressionEval(auc.dat,
                            tr.dat,
                             mol.feature=.x[1],
                                mol.feature.name=.x[2],
                                          dplyr::rename(auc.dat,Sample='AML sample'),
                                               dplyr::rename(te.dat,Sample='AML sample'),
                                                 category='Condition',doEnet=TRUE))
  
  
  

  e.results<-enet.reg.preds%>%
    mutate(method='ElasticNet')
 
  full.results<-reg.preds%>%
    mutate(method='LASSO')

  lr.results<-log.reg.preds%>%
    mutate(method='LogisticReg')%>%
    mutate(MSE=MSE*10000)
  
  #full.results<-rbind(full.results,lr.results)
  saveRDS(full.results,'lassoRegPreds.rds')
  saveRDS(lr.results,'logRegPreds.rds')
  saveRDS(e.results,'enetRegPreds.rds')
#  saveRDS(full.results,'mostlyCompletePredictions.rds')
  
  all.res<-rbind(full.results,lr.results,e.results)
  print(dim(all.res))
  return(all.res)
  
})

res<-do.call(rbind,res)
saveRDS(res,'combinedKfoldRes.rds')
```

Now we have all the data files and ready to go!

Let's plot the k-fold results by data type. 
```{r plot results}
library(ggplot2)

library(wesanderson)
pal<-wes_palette('Darjeeling1')

p3<-res%>%
  subset(numFeatures>0)%>%
    ggplot(aes(y=corVal,x=Molecular,fill=Molecular))+
          geom_boxplot(aes(alpha=0.5))+
          geom_jitter(aes(color=Molecular))+
  scale_color_manual(values=pal)+
scale_fill_manual(values=pal)+facet_grid(~method)

p3
write.csv(res%>%group_by(Molecular,method)%>%summarize(meanCor=mean(corVal,na.rm=T))%>%arrange(desc(meanCor)),file='table1_overallPerf.csv')
ggsave('5foldCVperformance.pdf',p3,width=12)

```

It seems that the protein data performs well. We can break down by drug as well to see how individual drugs perform.

```{r plot drugs, message=F}

  p4<-res%>%subset(numFeatures>0)%>%
    ggplot(aes(y=corVal,x=compound,fill=Molecular))+
          geom_boxplot(aes(alpha=0.5))+
       #   geom_jitter(aes(color=Molecular))+
  scale_color_manual(values=pal)+
scale_fill_manual(values=pal)+coord_flip()+facet_grid(~method)

p4

write.csv(res%>%group_by(Molecular,method,compound)%>%summarize(meanCor=mean(corVal,na.rm=T),samples=mean(numSamples,na.rm=T))%>%arrange(desc(meanCor)),'table2.csv')
ggsave('5foldCVperformanceByDrug.pdf',p4,height=12,width=10)
```

We need a new plot to evaluate size of predictors. 

```{r warning=FALSE, message=FALSE}

#get the multiomic preds
multi<-grep(';',res$Molecular)


```