---
title: "Build K-fold models for use"
author: "Sara Gosline"
date: "10/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#source("../util/synapseUtil.R")
#source("../util/loading_data.R")
#library(amlresistancenetworks)

#If unable to connect to synapse, source -
source("../util/loading_data_local.R")
```

## Load data
We have created a standard file to load all the data needed for the analysis.

```{r load data,warning=FALSE, message=FALSE}
global.data <- load.global.data("//Users//dann850//Desktop//CPTAC_PTRC//BeatAML_210//BeatAMLProteomics.csv")
phospho.data <- load.phospho.data("//Users//dann850//Desktop//CPTAC_PTRC//BeatAML_210//BeatAMLPhosphoproteomics.csv")
functional.data <- load.functional.data("//Users//dann850//Desktop//CPTAC_PTRC//BeatAML_210//FullPTCohortFunctionalData.csv")
```

## Now we can build the models

We want to build both the LASSO and logistic regression models using the following code. This code will divide the patients into 3 batches and train/test the model holding out 1/3 of the data each time. Then we will have predictions on each of the samples. 

```{r build models, warning=FALSE, message=FALSE}

all.pats <- intersect(pat.data$`AML sample`,auc.dat$`AML sample`)
library(caret)
folds<-createFolds(all.pats,k=5)
##separate out data?

res<-lapply(folds,function(testpats){
    test.pats<-all.pats[testpats]
    train.pats<-setdiff(all.pats,test.pats)

 print("Phospho preds first")
  substrate.dat<-pat.phos%>%
    dplyr::select(`AML sample`='Sample',Gene='site',Phosphosite='LogFoldChange')
  
          tr.dat<-subset(substrate.dat,`AML sample`%in%train.pats)
      te.dat<-subset(substrate.dat,`AML sample`%in%test.pats)
      
    phospho.reg.results<-drugMolRegressionEval(auc.dat,
                                                 tr.dat,
                                               'Phosphosite',
                                               dplyr::rename(auc.dat,Sample='AML sample'),
                                               dplyr::rename(te.dat,Sample='AML sample'))%>%
        do.call(rbind,.)
      
    
    phospho.lr.results<-drugMolLogRegEval(auc.dat,tr.dat,'Phosphosite',
                                          dplyr::rename(auc.dat,Sample='AML sample'),
                                           dplyr::rename(te.dat,Sample='AML sample'))%>%
      do.call(rbind,.)
  
    
  print('Now getting gene-based preds')
  
      tr.dat<-subset(pat.data,`AML sample`%in%train.pats)
      te.dat<-subset(pat.data,`AML sample`%in%test.pats)
    ##now train model on AML and eval on depmap data
  reg.preds<-purrr::map_df(list(mRNA='mRNALevels',
                              protein='proteinLevels',
                            #  gene='geneMutations',
                               binGene='binaryMutations'),
                         ~ drugMolRegressionEval(auc.dat,
                            tr.dat,
    .x,
                                                    dplyr::rename(auc.dat,Sample='AML sample'),
                                               dplyr::rename(te.dat,Sample='AML sample'),
                                                 category='Condition'))
    print("Comparing logistic preds")
  log.reg.preds<-purrr::map_df(list(protein='proteinLevels',
                                  mRNA='mRNALevels',
                           #   gene='geneMutations',
                              binGene='binaryMutations'),
                             ~ drugMolLogRegEval(auc.dat,
                                                tr.dat,
                                                .x,
                                               dplyr::rename(auc.dat,Sample='AML sample'),
                                               dplyr::rename(te.dat,Sample='AML sample'),
                                              category='Condition'))


 
  full.results<-rbind(reg.preds,phospho.reg.results)%>%
    mutate(method='LASSO')

  lr.results<-rbind(log.reg.preds,phospho.lr.results)%>%
    mutate(method='LogisticReg')%>%
    mutate(MSE=MSE*10000)
  
  #full.results<-rbind(full.results,lr.results)
  saveRDS(full.results,'lassoRegPreds.rds')
  saveRDS(lr.results,'logRegPreds.rds')
#  saveRDS(full.results,'mostlyCompletePredictions.rds')
  return(rbind(full.results,lr.results))
})

res<-do.call(rbind,res)
saveRDS(res,'combinedKfoldRes.rds')
```

Now we have all the data files and ready to go!

Let's plot the k-fold results by data type. 
```{r plot results}
library(ggplot2)

library(wesanderson)
pal<-wes_palette('Darjeeling1')

p3<-res%>%
  subset(numFeatures>0)%>%
    ggplot(aes(y=corVal,x=Molecular,fill=Molecular))+
          geom_boxplot(aes(alpha=0.5))+
          geom_jitter(aes(color=Molecular,size=numFeatures,shape=method))+
  scale_color_manual(values=pal)+
scale_fill_manual(values=pal)

p3

ggsave('5foldCVperformance.pdf',p3)

```

It seems that the protein data performs well. We can break down by drug as well to see how individual drugs perform.

```{r plot drugs, message=F}

  p4<-res%>%subset(numFeatures>0)%>%
    ggplot(aes(y=corVal,x=compound,fill=Molecular))+
          geom_boxplot(aes(alpha=0.5))+
          geom_jitter(aes(color=Molecular,size=numFeatures,shape=method))+
  scale_color_manual(values=pal)+
scale_fill_manual(values=pal)+coord_flip()

p4

ggsave('5foldCVperformanceByDrug.pdf',p4,height=12)
```