---
title: "Figure 4 Predictor results"
author: "Sara Gosline"
date: "03/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(amlresistancenetworks)
library(wesanderson)
source("../beatAMLdata.R")
pal<-wes_palette('Darjeeling1')
if(!exists('dataLoaded')){
  loadBeatAMLData()
  #loadUnNormPhosData()
  dataLoaded=TRUE
}

```

## Building of predictors

Building the predictors - LASSO and Logistic regression. We build predictors using the best performing models from leave-one-out cross validation. 

```{r build predictors, echo=FALSE, warning=FALSE}
 
substrate.dat<-pat.phos%>%
    dplyr::select(`AML sample`='Sample',Gene='site',Phosphosite='LogFoldChange')

if(file.exists('lassoRegPreds.rds')){ 
  reg.preds <- readRDS('lassoRegPreds.rds')
}else{
  
  print("Getting LASSO preds")
  phospho.reg.results<-drugMolRegression(auc.dat,substrate.dat,'Phosphosite')
  reg.preds<-purrr::map_df(list(mRNA='mRNALevels',
                                protein='proteinLevels',
                               # gene='geneMutations',
                                binGene='binaryMutations'),
                           ~ drugMolRegression(auc.dat,
                                               pat.data,                                
                                               .x,
                                               category='Condition'))%>%
    rbind(phospho.reg.results)
  saveRDS(reg.preds,'lassoRegPreds.rds')

}

if(file.exists("logRegPreds.rds")){
  logr.preds <- readRDS('logRegPreds.rds')
}else{
  print("Getting Logistic Preds")
  
  phospho.lr.results<-drugMolLogReg(auc.dat,substrate.dat,'Phosphosite')

  logr.preds<-purrr::map_df(list(binGene='binaryMutations',
                                  mRNA='mRNALevels',
                                 protein='proteinLevels'),
                                 #    gene='geneMutations'),                                
                            ~ drugMolLogReg(auc.dat,pat.data,.x,category='Condition'))%>%
  rbind(phospho.lr.results)
  saveRDS(logr.preds,'logRegPreds.rds')

}


  ##now let's summarize how many predictors we have for each drug
  l.sum.tab<-logr.preds%>%subset(numFeatures>0)%>%
    group_by(Molecular)%>%summarize(NumberModels=n_distinct(var),Error=mean(MSE),
                                    Cor=mean(corVal),
                                   SamplesUsed=mean(numSamples,na.rm=T))%>%mutate(Type='Logistic')
  
  sum.tab<-reg.preds%>%subset(numFeatures>0)%>%
    group_by(Molecular)%>%
    summarize(NumberModels=n_distinct(var),Error=mean(MSE),                                    Cor=mean(corVal),
 SamplesUsed=mean(numSamples,na.rm=T))%>%mutate(Type='LASSO')%>%
    rbind(l.sum.tab)
  knitr::kable(sum.tab)
  write.csv(sum.tab,file='table2_perf.csv')
  
  ##save both predictions to file
    
```

Now we have predictive models for each of the four data types, though not all data types had enough for the LOO cross validation. 

## Plot model summary
```{r plot model performance}

reg.preds<-sum.tab%>%subset(Type=='LASSO')%>%
  dplyr::select(Molecular,NumberModels)%>%distinct()%>%
  right_join(reg.preds)

logr.preds<-sum.tab%>%subset(Type=='Logistic')%>%
  dplyr::select(Molecular,NumberModels)%>%distinct()%>%
  right_join(logr.preds)


```

## Plot of overall performers

We compare the plots in two ways - one using the misclassification error and another using the spearman rank correlation. 
```{r best performers, echo=FALSE, warning=FALSE}

library(ggplot2)

p1<-logr.preds%>%subset(numFeatures>0)%>%
 ggplot(aes(y=MSE,x=Molecular,fill=Molecular))+
          geom_boxplot(aes(alpha=0.5))+
          geom_jitter(aes(color=Molecular,size=numFeatures))+
  scale_color_manual(values=pal)+
  scale_fill_manual(values=pal)+
    #scale_fill_viridis_d()+
  #scale_color_viridis_d()+
  ggtitle("Logistic Regression Predictor Performance")
print(p1)

p2<-reg.preds%>%subset(numFeatures>0)%>%
ggplot(aes(y=MSE,x=Molecular,fill=Molecular))+
          geom_boxplot(aes(alpha=0.5))+
          geom_jitter(aes(color=Molecular,size=numFeatures))+
  scale_color_manual(values=pal)+
  scale_fill_manual(values=pal)+
  ggtitle("LASSO Regression Predictor Performance")
print(p2)

ggsave('suppfig2b_logistic.pdf',p1)
ggsave('suppfig2a_lasso.pdf',p2)

```

Now we can plot the correlation values to see how the models co-exist

```{r rank performance}
library(ggplot2)

p1<-logr.preds%>%subset(numFeatures>0)%>%
 ggplot(aes(y=corVal,x=Molecular,fill=Molecular))+
          geom_violin(aes(alpha=0.5))+
          geom_jitter(aes(color=Molecular,size=numFeatures))+
  scale_color_manual(values=pal)+
  scale_fill_manual(values=pal)+
    #scale_fill_viridis_d()+
  #scale_color_viridis_d()+
  ggtitle("Logistic Regression Predictor Performance")
print(p1)

p2<-reg.preds%>%subset(numFeatures>0)%>%
ggplot(aes(y=corVal,x=Molecular,fill=Molecular))+
          geom_violin(aes(alpha=0.5))+
          geom_jitter(aes(color=Molecular,size=numFeatures))+
  scale_color_manual(values=pal)+
  scale_fill_manual(values=pal)+
  ggtitle("LASSO Regression Predictor Performance")
print(p2)

p3<-reg.preds%>%
    mutate(Model='LASSO')%>%
    rbind(mutate(logr.preds,Model='Logistic'))%>%
  subset(numFeatures>0)%>%
ggplot(aes(y=corVal,x=Molecular,fill=Molecular))+
          geom_boxplot(aes(alpha=0.5))+
          geom_jitter(aes(color=Molecular,size=numFeatures,shape=Model))+
  scale_color_manual(values=pal)+
  scale_fill_manual(values=pal)+
  ggtitle("Total Predictor Correlation")
print(p3)
ggsave('fig4a_corlogistic.pdf',p1)
ggsave('fig4b_corlasso.pdf',p2)
ggsave('fig4_corFull.pdf',p3)



```
