---
title: "Proteomic data processing"
author: "Sara Gosline"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
```

## Data formats

Using the data available in the crosstabs and a second metadata file, we can create a Synapse table that is in 'long form' and has all the protein/phosphosite values together with their relevant metadata.

```{r synapselogin}
source("../util/synapseUtil.R")

syn <- synapseLogin()

##CUSTOMIZE THIS FOR YOUR PROJECT
global.syn<-'syn25572051' 
##CUSTOMIZE THIS FOR YOUR PROJECT
phos.syn<- 'syn25572049'

##CUSTOMIZE THIS FOR YOUR PROJECT
metadata.syn <- 'syn25572050'

```



## Merge metadata and proteomics data

Now we can merge 

```{r pressure, echo=FALSE}
##CUSTOMIZE THIS FOR YOUR PROJECT - what columsn do we want? 
  meta.tab <- readxl::read_xlsx(syn$get(metadata.syn)$path)%>%
    dplyr::select(cellLine='Sample description',Ligand,Treatment, 
                  'Time (minutes)', Sample='SampleName')%>%
    distinct()

  print(head(meta.tab))
  
  mat <- read.csv(syn$get(global.syn)$path,sep='\t',header=T)
  
  ##LogRatio and Sample might be updated to something else
  mat<-mat%>%tidyr::pivot_longer(cols=c(5:ncol(mat)),names_to='Sample',
                        values_to='LogRatio')%>%
    left_join(meta.tab)%>%
        subset(!is.na(LogRatio))
  print(head(mat))

```

Now we have the proteomics and metadata joined in a single table.

## Merge metadata and phosphoproteomics data

```{r phospho, echo=FALSE}

 pmat <- read.csv(syn$get(phos.syn)$path,sep='\t',header=T)

     ##LogRatio and Sample might be updated to something else
  phos.mat<-pmat%>%tidyr::pivot_longer(cols=c(6:ncol(pmat)),names_to='Sample',
                        values_to='LogRatio')%>%
     left_join(meta.tab)%>%
    subset(!is.na(LogRatio))
  print(head(phos.mat))
```

Now we have the phospho data as well.  The last thing that needs to be done is to store them both to a synapse table. This requires knowing the synapse project (in our case, the [P3 Project](http://synapse.org/p3)).

```{r store tables}

##UPDATE TABLE NAME! UPDATE SYNAPSE ID!
pt<-synTableStore(mat,"Test proteomics data",parentId='syn25005572')
pht<-synTableStore(phos.mat,"Test phosphoproteomics data",parentId='syn25005572')


```