---
title: "Network analysis of predictive genes"
author: "Sara Gosline"
date: "6/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Network analysis 

The goal of this document is to study the larger networks that comprise the genes that predict subtype.

```{r load data,message=F,echo=T,warning=F}

source("../util/synapseUtil.R")
library(dplyr)
syn<-synapseLogin()
##load in metagenes

tab <- syn$tableQuery('select model,drug_name,source_data,feature_names,feature_scores from syn31854061 where "model"=\'gbt2\' and "source_data"=\'phospho_proteomics\'')$asDataFrame()

reg_prots<-tab%>%
  select(-feature_scores)%>%
  ###what is the upward limit of model size? 
  tidyr::separate(feature_names,into=paste0('col',c(1:100)),sep='\\|')%>%
  tidyr::pivot_longer(names_to=NULL,cols=starts_with('col'))%>%
  subset(!is.na(value))%>%
  dplyr::select(-source_data)%>%
  tidyr::separate(value,into=c('gene','data'),sep='_')%>%
  tidyr::replace_na(list(data='Phospho'))

reg_scores<-tab%>%
  select(-feature_names)%>%
    tidyr::separate(feature_scores,into=c(paste0('col',c(1:100),sep='\\|')))%>%
 tidyr::pivot_longer(names_to=NULL,cols=starts_with('col'))%>%
  subset(!is.na(value))%>%
  dplyr::select(-source_data)
  #tidyr::separate(value,into=c('score','data'),sep='_')%>%
  #tidyr::replace_na(list(data='Phospho'))

reg_sigs<-cbind(reg_prots,score=reg_scores$value)

reg_sigs$data_type=sapply(reg_sigs$data,function(x) ifelse(x=='prot','Global',x))


cols=nationalparkcolors::park_palette("Badlands")
clus.cols=RColorBrewer::brewer.pal(8,'Dark2')[c(2,3,5,6)]


```

First we want to look at how many features are being selected by each metagene here.

```{r feature summaries}
library(ggplot2)


##summary of features
ggplot(reg_sigs)+
    geom_bar(aes(x=drug_name,fill=data_type),position='dodge')+
    scale_fill_manual(values=cols)+
  coord_flip()
#    theme(axis.text.x = element_text(angle = 90))+


ggsave('drug_gbt_genes.pdf')



```
There is generally a division between various data types 


## Create function to take proteins by type and map them to network

Then we can do network creation to build larger networks. This will enhance our ability to do functional enrichment. We need to keep track of two things - one is the type of feature we are handling, and the other is the weight for the network.

```{r network mapping,message=F,echo=T,warning=F}

build_run_network <- function(gene_name, 
                              gene_score, 
                              gene_type,
                              drug_name,
                              rands=100,
                              store=FALSE){
  
  library(amlresistancenetworks)
  
  wes_inds <- which(gene_type=='WES')
  rna_inds <- which(gene_type=='RNA')
  prot_inds <- which(gene_type=='Global')
  phos_inds <- which(gene_type=='Phospho')
  
  netname=paste0('predsOf',drug_name)
  
    ##rename WES
  wes_terms <- gene_score[wes_inds]
  names(wes_terms) <- gene_name[wes_inds]
  
  #lets separate out the substrates from the phosphosite data
  phgenes<-data.frame(genes=gene_name[phos_inds],val=gene_score[phos_inds])%>%
    tidyr::separate(genes, into=c('gene','mod'),sep='-')%>%
    mutate(modlist=strsplit(mod,split='s|t|y'))%>%
    tidyr::unnest()%>%
    rowwise()%>%
    mutate(newgene=paste(gene,modlist,sep='-'))
  
  kin_terms <- phgenes$val
  names(kin_terms) <- phgenes$newgene
  
 # print(kin_terms)
  prot_terms <- gene_score[prot_inds]
  names(prot_terms) <- gene_name[prot_inds]
  

  rna_terms <- gene_score[rna_inds]
  names(rna_terms) <- gene_name[rna_inds]
  if(length(gene_name)<10)
    x=100
  else if(length(gene_name)>400)
    x=.1
  else
    x=1
  
  netlist = amlresistancenetworks::computePhosphoNetwork(phos.vals=kin_terms,
                                                     prot.vals=prot_terms,
                                                     gene.vals=wes_terms,
                                                     rna.vals=rna_terms,
                                                     beta=x,
                                                     nrand=rands,
                                                     fname=netname)
  graph =netlist$graph
  fname = netlist$fname[1]
  print(fname)
  if(store){
    synapseStore(fname,'syn26997954')
  }  
  return(graph)
}

```

Now we can run the analysis


```{r k4 cluster analysis, message=FALSE, warning=FALSE}
library(dplyr)


##compute jaccard distance between sets of proteins

drugs <- unique(reg_sigs$drug_name)

#ggplot(reg_sigs,aes(x=drug,fill=data_type))+geom_bar(position='dodge')+
#    coord_flip()+ggtitle("number of nodes in each model")

 if(!file.exists('regressionDrugSigNets.Rds')){
 reg_nets <- reg_sigs %>%
   #subset(drug%in%drugas)%>%
    mutate(weight=as.numeric(score))%>%
     mutate(nrand=10)%>%
 #  subset(Cluster==4)%>%
     group_by(drug_name)%>%
   #mutate(weight=10)%>%
   dplyr::rename(feature='gene')%>%
     dplyr::summarize(net=list(build_run_network(feature,
                                              weight,
                                              data_type,
                                              drug_name,
                                              nrand,store=TRUE)))
  saveRDS(reg_nets,file='regressionDrugSigNets.Rds')
 }else
   reg_nets<-readRDS('regressionDrugSigNets.Rds')

 
  net_nodes<-reg_nets%>%
    rowwise()%>%
    mutate(proteins=list(vertex.attributes(net)$name),
                            dataType=list(vertex.attributes(net)$nodeType),
                            nodeType=list(vertex.attributes(net)$type))%>%
    tidyr::unnest(cols=c(proteins,dataType,nodeType))
  
  net_nodes$data_type=sapply(net_nodes$dataType,function(x) ifelse(x=='proteins','Global','Phospho'))
  
  write.table(select(net_nodes,-net),file='regressioNetworkNodes.tsv',sep='\t')
  
  ggplot(net_nodes,aes(x=drug_name,fill=data_type))+geom_bar(position='dodge')+
    facet_grid(~nodeType)+
    coord_flip()+scale_fill_manual(values=cols)+
    ggtitle("number of nodes in each network")
  ggsave('gbt_pcsf_net_nodes.pdf')
```


Compare networks to original sets, by looking at distance between the various signatures.
```{r compare drug signatures}

jaccard <- function(lista,listb){
  return(length(intersect(lista,listb))/length(union(lista,listb)))
}


##compute jaccard distance between networks
resmat<-sapply(drugs,function(x){
  sapply(drugs,function(y) { jaccard(subset(reg_sigs,drug==x)$gene,subset(reg_sigs,drug==y)$gene)})})


pred_dist<-resmat%>%as.data.frame()%>%
  tibble::rownames_to_column('drug2')%>%
  tidyr::pivot_longer(c(2:(ncol(netmat)+1)),names_to='drug1',values_to='Jaccard Regression')%>%
  subset(`Jaccard Regression`!=1)%>%
  arrange(desc(`Jaccard Regression`))

write.table(pred_dist,file='predictorJaccardDists.tsv',sep='\t')


##create table to plot
library(pheatmap)
pheatmap(resmat,filename='predictorJaccardHeatmap.pdf')
library(ggrepel)

##multidimensional scale
toplot<-cmdscale(as.dist(1-resmat))

colnames(toplot)=c('dim1','dim2')
toplot%>%as.data.frame()%>%
  tibble::rownames_to_column('Drug')%>%
  ggplot(aes(x=dim1,y=dim2,label=Drug,alpha=0.5))+geom_point()+geom_text_repel()+ggtitle('Jaccard-based distance between proteins')

ggsave('predictorJaccard_cmdScale.pdf')




```

We can see that panobinostat and venetoclax are very similar compared to the other drugs. what happens when we compare the nodes in the network?

```{r compare drug networks}

netmat<-sapply(drugs,function(x){
  sapply(drugs,function(y){
    jaccard(unlist(subset(net_nodes,drug==x)$proteins),
            unlist(subset(net_nodes,drug==y)$proteins))
  })
})

pheatmap(netmat,filename='predcitorNetworkHeatmap.pdf')

net_dist<-netmat%>%as.data.frame()%>%
  tibble::rownames_to_column('drug2')%>%
  tidyr::pivot_longer(c(2:(ncol(netmat)+1)),names_to='drug1',values_to='Jaccard Network')%>%
  subset(`Jaccard Network`!=1)%>%
  arrange(desc(`Jaccard Network`))

write.table(net_dist,file='networkJaccardDists.tsv',sep='\t')


full_dist <- net_dist%>%full_join(pred_dist,on=c('drug2','drug1'))

p<-full_dist%>%subset(drug2=='Venetoclax')%>%
    ggplot(aes(x=`Jaccard Network`,
             y=`Jaccard Regression`,col=drug1))+geom_point()
  
 # tidyr::unite(drug1,drug2,col='combo')

toplot<-cmdscale(as.dist(1-netmat))
colnames(toplot)=c('dim1','dim2')
toplot%>%as.data.frame()%>%
  tibble::rownames_to_column('Drug')%>%
  ggplot(aes(x=dim1,y=dim2,label=Drug,alpha=0.5))+
  geom_point()+geom_text_repel()+
  ggtitle('Jaccard-based distance between networks')

```