---
title: "Metagene Network Analysis"
author: "Sara Gosline"
date: "1/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Network analysis 

The goal of this document is to study the larger networks that comprise the metagenes. We want to map all the metagene features to one single network.

```{r load data,message=F,echo=T,warning=F}

library(PCSF)
source("../util/synapseUtil.R")
library(dplyr)
syn<-synapseLogin()
##load in metagenes
k5  <- read.table(syn$get('syn26718017')$path,sep='\t',header=T)
k8 <- read.table(syn$get('syn26718020')$path,sep='\t',header=T)

```

## Create function to take proteins by type and map them to network

We need to keep track of two things - one is the type of feature we are handling, and the other is the weight for the network.

```{r network mapping,message=F,echo=T,warning=F}

build_run_network <- function(gene_name, 
                              gene_score, 
                              gene_type,
                              cluster,
                              k,
                              rands=100){
  
  library(amlresistancenetworks)
  wes_inds <- which(gene_type=='WES')
  rna_inds <- which(gene_type=='RNA')
  prot_inds <- which(gene_type=='Global')
  phos_inds <- which(gene_type=='Phospho')
  
  netname=paste0('k',k,'_cluster',cluster)
  
    ##rename WES
  wes_terms <- gene_score[wes_inds]
  names(wes_terms) <- gene_name[wes_inds]
  
  #lets separate out the substrates from the phosphosite data
  phgenes<-data.frame(genes=gene_name[phos_inds],val=gene_score[phos_inds])%>%
    tidyr::separate(genes, into=c('gene','mod'),sep='-')%>%
    mutate(modlist=strsplit(mod,split='s|t|y'))%>%
    tidyr::unnest()%>%
    rowwise()%>%
    mutate(newgene=paste(gene,modlist,sep='-'))
  
  kin_terms <- phgenes$val
  names(kin_terms) <- phgenes$newgene
  
 # print(kin_terms)
  prot_terms <- gene_score[prot_inds]
  names(prot_terms) <- gene_name[prot_inds]
  
  net = amlresistancenetworks::computePhosphoNetwork(phos.vals=kin_terms,
                                                     prot.vals=prot_terms,
                                                     gene.vals=wes_terms,
                                                     nrand=rands,
                                                     fname=netname)
  return(net)
}

print("K5 Clusters")
k5%>%group_by(data_type)%>%summarize(numFeatures=n())


print('K8 Clusters')
k8%>%group_by(data_type)%>%summarize(numFeatures=n())
```



```{r k5 cluster analysis}
k5_nets <- k5 %>%
    mutate(k=5,nrand=10)%>%
    group_by(Cluster)%>%
    dplyr::mutate(net=list(build_run_network(feature,
                                             feature_score,
                                             data_type,
                                             Cluster,
                                             k,
                                             nrand)))

```

Now we can summarize the networks for the k5 clusters

```{r k8 cluster analysis}
k8_nets <- k8 %>%
    mutate(k=8,nrand=20)%>%
    group_by(Cluster)%>%
    dplyr::mutate(net=list(build_run_network(feature,feature_score,data_type,Cluster,k,nrand)))


```