---
title: "Metagene Network Analysis"
author: "Sara Gosline"
date: "1/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Network analysis 

The goal of this document is to study the larger networks that comprise the metagenes. We want to map all the metagene features to one single network.

```{r load data,message=F,echo=T,warning=F}

source("../util/synapseUtil.R")
library(dplyr)
syn<-synapseLogin()
##load in metagenes
k5  <- read.table(syn$get('syn26718017')$path,sep='\t',header=T)
k8 <- read.table(syn$get('syn26718020')$path,sep='\t',header=T)

##now load in alternate scores
k5_weights <- read.table(syn$get('syn26969682')$path,sep='\t',header=T)%>%
  tidyr::pivot_longer(starts_with('Cluster.'),names_to='clusterWeight',values_to = 'weight')%>%
  mutate(fromCluster=Cluster==gsub('Cluster.','',clusterWeight,fixed=T))%>%
  subset(fromCluster)


##now load in alternate scores
k8_weights <- read.table(syn$get('syn26969686')$path,sep='\t',header=T)%>%
  tidyr::pivot_longer(starts_with('Cluster.'),names_to='clusterWeight',values_to = 'weight')%>%
  mutate(fromCluster=Cluster==gsub('Cluster.','',clusterWeight,fixed=T))%>%
  subset(fromCluster)


```

First we want to look at how many features are being selected by each metagene here.

```{r feature summaries}
library(ggplot2)


##summary of features
ggplot(k5_weights)+geom_bar(aes(x=Cluster,fill=data_type),position='dodge')

ggplot(k8_weights)+geom_bar(aes(x=Cluster,fill=data_type),position='dodge')


```

There is clearly a disparity. However, that makes sense as the molecules correlate in various ways. 

Now we can evaluate how similar the weights are to the feature scores

```{r feature_scores}


ggplot(k5_weights)+
  geom_point(aes(x=feature_score,y=weight,alpha=0.5,shape=data_type,col=as.factor(Cluster)))


ggplot(k8_weights)+geom_point(aes(x=feature_score,y=weight,alpha=0.5,
                                  shape=data_type,col=as.factor(Cluster)))


```
## Create function to take proteins by type and map them to network

We need to keep track of two things - one is the type of feature we are handling, and the other is the weight for the network.

```{r network mapping,message=F,echo=T,warning=F}

build_run_network <- function(gene_name, 
                              gene_score, 
                              gene_type,
                              cluster,
                              k,
                              rands=100,
                              store=FALSE){
  
  library(amlresistancenetworks)
  
  wes_inds <- which(gene_type=='WES')
  rna_inds <- which(gene_type=='RNA')
  prot_inds <- which(gene_type=='Global')
  phos_inds <- which(gene_type=='Phospho')
  
  netname=paste0('k',k,'_cluster',cluster)
  
    ##rename WES
  wes_terms <- gene_score[wes_inds]
  names(wes_terms) <- gene_name[wes_inds]
  
  #lets separate out the substrates from the phosphosite data
  phgenes<-data.frame(genes=gene_name[phos_inds],val=gene_score[phos_inds])%>%
    tidyr::separate(genes, into=c('gene','mod'),sep='-')%>%
    mutate(modlist=strsplit(mod,split='s|t|y'))%>%
    tidyr::unnest()%>%
    rowwise()%>%
    mutate(newgene=paste(gene,modlist,sep='-'))
  
  kin_terms <- phgenes$val
  names(kin_terms) <- phgenes$newgene
  
 # print(kin_terms)
  prot_terms <- gene_score[prot_inds]
  names(prot_terms) <- gene_name[prot_inds]
  

  rna_terms <- gene_score[rna_inds]
  names(rna_terms) <- gene_name[rna_inds]
  if(length(gene_name)>200)
    x=1
  else
    x=100
  
  netlist = amlresistancenetworks::computePhosphoNetwork(phos.vals=kin_terms,
                                                     prot.vals=prot_terms,
                                                     gene.vals=wes_terms,
                                                     rna.vals=rna_terms,
                                                     beta=x*.1,
                                                     nrand=rands,
                                                     fname=netname)
  graph =netlist$graph
  fname = netlist$fname[1]
  print(fname)
  if(store){
    synapseStore(fname,'syn26997954')
  }  
  return(graph)
}

print("K5 Clusters")
k5_weights%>%group_by(data_type,Cluster)%>%summarize(numFeatures=n())%>%arrange(Cluster)%>%
  DT::datatable()


print('K8 Clusters')
k8_weights%>%group_by(data_type, Cluster)%>%summarize(numFeatures=n())%>%arrange(Cluster)%>%
  DT::datatable()
```

Now we can run the analysis


```{r k5 cluster analysis}
k5_nets <- k5_weights %>%
    mutate(k=5,nrand=100)%>%
    group_by(Cluster)%>%
    dplyr::summarize(net=list(build_run_network(feature,
                                             weight,
                                             data_type,
                                             Cluster,
                                             k,
                                             nrand,store=TRUE)))

getNetStats <- function(net_list,orig_tab){
  require(igraph)
  ##create table of metagene features based on wht was included/excluded from the network
  node_tab<-purrr::map2_dfr(.x=net_list$Cluster,.y=net_list$net,.f=function(Cluster,net){
    g<-net
    data.frame(node=names(V(g)),prizes=vertex.attributes(g)$prize,
               types=vertex.attributes(g)$type,lfc=vertex.attributes(g)$logFoldChange,
               geneTypes=vertex.attributes(g)$nodeType,Cluster=rep(Cluster,length(V(g))))
  } )
  
  graph_counts <- node_tab%>%group_by(geneTypes,types,Cluster)%>%summarize(count=n())%>%
    dplyr::rename(data_type='geneTypes',source='types')
  orig_counts <- orig_tab%>%group_by(data_type,Cluster)%>%summarize(count=n())%>%
    mutate(source='Metagene')

  ##counts of how man proteins/genes/phosphosites in each
  return(rbind(graph_counts,orig_counts))  
}

DT::datatable(getNetStats(k5_nets,k5_weights))

```

Now we can summarize the networks for the k5 clusters, and move on to k8 clusters

```{r k8 cluster analysis}

#k8_nets <- k8_weights %>%
#    mutate(k=8,nrand=100)%>%
#    group_by(Cluster)%>%
#    dplyr::summarize(net=list(build_run_network(feature,weight,data_type,Cluster,k,nrand,
                                            #    store=TRUE)))

#DT::datatable(getNetStats(k8_nets,k8_weights))


```
