---
title: "Metagene Network Analysis"
author: "Sara Gosline"
date: "1/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Network analysis 

The goal of this document is to study the larger networks that comprise the metagenes. We want to map all the metagene features to one single network.

```{r load data,message=F,echo=T,warning=F}

source("../util/synapseUtil.R")
library(dplyr)
syn<-synapseLogin()
##load in metagenes
k5  <- read.table(syn$get('syn26718017')$path,sep='\t',header=T)
k8 <- read.table(syn$get('syn26718020')$path,sep='\t',header=T)

##now load in alternate scores
k5_weights <- read.table(syn$get('syn26969682')$path,sep='\t',header=T)%>%
  tidyr::pivot_longer(starts_with('Cluster.'),names_to='clusterWeight',values_to = 'weight')%>%
  mutate(fromCluster=Cluster==gsub('Cluster.','',clusterWeight,fixed=T))%>%
  subset(fromCluster)


##now load in alternate scores
k8_weights <- read.table(syn$get('syn26969686')$path,sep='\t',header=T)%>%
  tidyr::pivot_longer(starts_with('Cluster.'),names_to='clusterWeight',values_to = 'weight')%>%
  mutate(fromCluster=Cluster==gsub('Cluster.','',clusterWeight,fixed=T))%>%
  subset(fromCluster)

k4_weights <- read.table(syn$get('syn26969681')$path,sep='\t',header=T)%>%
  tidyr::pivot_longer(starts_with('Cluster.'),names_to='clusterWeight',values_to = 'weight')%>%
  mutate(fromCluster=Cluster==gsub('Cluster.','',clusterWeight,fixed=T))%>%
  subset(fromCluster)


cols=nationalparkcolors::park_palette("Badlands")
clus.cols=RColorBrewer::brewer.pal(8,'Dark2')[c(2,3,5,6)]


```

First we want to look at how many features are being selected by each metagene here.

```{r feature summaries}
library(ggplot2)


##summary of features
ggplot(k4_weights)+
    geom_bar(aes(x=Cluster,fill=data_type),position='dodge')+
    scale_fill_manual(values=cols)

ggsave('k4_metagenes.pdf')

ggplot(k5_weights)+
    geom_bar(aes(x=Cluster,fill=data_type),position='dodge')+
  scale_fill_manual(values=cols)


```

There is clearly a disparity. However, that makes sense as the molecules correlate in various ways. 

## Metagene plotting

To plot the proteins and transcripts, we need to collect the original expression data.

```{r plot metagenes}
library(pheatmap)

source("../util/loading_data.R")
#combined <- load.combined.data()
```



## Metagene weights for gene enrichment.  

Now we can evaluate how similar the weights are to the feature scores

```{r feature_scores}


ggplot(k5_weights)+
  geom_point(aes(x=feature_score,y=weight,alpha=0.5,
                 shape=data_type,col=as.factor(Cluster)))


ggplot(k8_weights)+geom_point(aes(x=feature_score,y=weight,alpha=0.5,
                                  shape=data_type,col=as.factor(Cluster)))



ggplot(k4_weights)+geom_point(aes(x=feature_score,
                                  y=weight,alpha=0.5,                                                                  shape=data_type,
                                  col=as.factor(Cluster)))+
    scale_color_manual(values=clus.cols)
```
They seem to correlate so we can use them for pathway enrichment

## Pathway enrichment

I believe camilo did GSEA enrichment of the metagene pathways. what is the result? How does it compare to GO enrichmnet?



## Create function to take proteins by type and map them to network

Then we can do network creation to build larger networks. This will enhance our ability to do functional enrichment. We need to keep track of two things - one is the type of feature we are handling, and the other is the weight for the network.

```{r network mapping,message=F,echo=T,warning=F}

build_run_network <- function(gene_name, 
                              gene_score, 
                              gene_type,
                              cluster,
                              k,
                              rands=100,
                              store=FALSE){
  
  library(amlresistancenetworks)
  
  wes_inds <- which(gene_type=='WES')
  rna_inds <- which(gene_type=='RNA')
  prot_inds <- which(gene_type=='Global')
  phos_inds <- which(gene_type=='Phospho')
  
  netname=paste0('k',k,'_cluster',cluster)
  
    ##rename WES
  wes_terms <- gene_score[wes_inds]
  names(wes_terms) <- gene_name[wes_inds]
  
  #lets separate out the substrates from the phosphosite data
  phgenes<-data.frame(genes=gene_name[phos_inds],val=gene_score[phos_inds])%>%
    tidyr::separate(genes, into=c('gene','mod'),sep='-')%>%
    mutate(modlist=strsplit(mod,split='s|t|y'))%>%
    tidyr::unnest()%>%
    rowwise()%>%
    mutate(newgene=paste(gene,modlist,sep='-'))
  
  kin_terms <- phgenes$val
  names(kin_terms) <- phgenes$newgene
  
 # print(kin_terms)
  prot_terms <- gene_score[prot_inds]
  names(prot_terms) <- gene_name[prot_inds]
  

  rna_terms <- gene_score[rna_inds]
  names(rna_terms) <- gene_name[rna_inds]
  if(length(gene_name)<10)
    x=100
  else if(length(gene_name)>400)
    x=.01
  else
    x=1
  
  netlist = amlresistancenetworks::computePhosphoNetwork(phos.vals=kin_terms,
                                                     prot.vals=prot_terms,
                                                     gene.vals=wes_terms,
                                                     rna.vals=rna_terms,
                                                     beta=x,
                                                     nrand=rands,
                                                     fname=netname)
  graph =netlist$graph
  fname = netlist$fname[1]
  print(fname)
  if(store){
    synapseStore(fname,'syn26997954')
  }  
  return(graph)
}


print("K4 Clusters")
k4_weights%>%group_by(data_type,Cluster)%>%summarize(numFeatures=n())%>%arrange(Cluster)%>%
  DT::datatable()



print("K5 Clusters")
k5_weights%>%group_by(data_type,Cluster)%>%summarize(numFeatures=n())%>%arrange(Cluster)%>%
  DT::datatable()


print('K8 Clusters')
k8_weights%>%group_by(data_type, Cluster)%>%
  summarize(numFeatures=n())%>%
  arrange(Cluster)%>%
  DT::datatable()
```

Now we can run the analysis


```{r k4 cluster analysis}
library(dplyr)
 k4_nets <- k4_weights %>%
     mutate(k=4,nrand=100)%>%
     group_by(Cluster)%>%
     dplyr::summarize(net=list(build_run_network(feature,
                                              weight,
                                              data_type,
                                              Cluster,
                                              k,
                                              nrand,store=TRUE)))

# k5_nets <- k5_weights %>%
#     mutate(k=5,nrand=100)%>%
#     group_by(Cluster)%>%
#     dplyr::summarize(net=list(build_run_network(feature,
#                                              weight,
#                                              data_type,
#                                              Cluster,
#                                              k,
#                                              nrand,store=TRUE)))

getNetStats <- function(net_list,orig_tab){
  require(igraph)
  ##create table of metagene features based on wht was included/excluded from the network
  node_tab<-purrr::map2_dfr(.x=net_list$Cluster,.y=net_list$net,.f=function(Cluster,net){
    g<-net
    data.frame(node=names(V(g)),prizes=vertex.attributes(g)$prize,
               types=vertex.attributes(g)$type,lfc=vertex.attributes(g)$logFoldChange,
               geneTypes=vertex.attributes(g)$nodeType,Cluster=rep(Cluster,length(V(g))))
  } )
  
  graph_counts <- node_tab%>%group_by(geneTypes,types,Cluster)%>%summarize(count=n())%>%
    dplyr::rename(data_type='geneTypes',source='types')
  orig_counts <- orig_tab%>%group_by(data_type,Cluster)%>%summarize(count=n())%>%
    mutate(source='Metagene')

  ##counts of how man proteins/genes/phosphosites in each
  return(rbind(graph_counts,orig_counts))  
}

DT::datatable(getNetStats(k4_nets,k4_weights))

```

Now we can summarize the networks for the k5 clusters, and move on to k8 clusters

```{r k8 cluster analysis}

#k8_nets <- k8_weights %>%
#    mutate(k=8,nrand=100)%>%
#    group_by(Cluster)%>%
#    dplyr::summarize(net=list(build_run_network(feature,weight,data_type,Cluster,k,nrand,
                                            #    store=TRUE)))

#DT::datatable(getNetStats(k8_nets,k8_weights))


```
