---
title: "AML_Network_Pacentine"
author: "Itallia V. Pacentine"
date: 2022
output: html_document
---

**Pathway enrichment with and without network analysis for k=4**

```{r setup}
source("../util/synapseUtil.R")
syn<-synapseLogin()
#browseVignettes(package="synapser")
#install.packages("devtools")
#devtools::install_github("biodataganache/leapR", build_vignette=FALSE, force = TRUE)
library(leapR)
library(tidyverse)
library(igraph)
```

**Required functions:**
```{r}
## Function to download network files from synapse for Beat-AML project
  # Input: 
    # 1. "syn": synID of a .gml file in quotes (""). For example: "syn27042882"
    # 2. "n": cluster number
  # Output: dataframe with 3 columns labeled "feature", "Cluster", and "data_type"
extract_graph <- function(syn, n){
  entity <- synGet(syn)
  filepath <- entity$path
  df <- read.graph(filepath, format=c("gml"))
  df <- data.frame(V(df)$name, 
                          c(rep(n,length(V(df)$name))),
                          V(df)$nodeType)
  colnames(df) <- c("feature","Cluster","data_type")
  return(df)
}


## Functions to do LeapR enrichment by sets.
  #required input:
    # 1. "df": dataframe with columns:
        # "feature" - gene names
        # "Cluster" - assigned cluster number.
        # any number of other columns
    # 2. "gset": name of database to extract pathway information. Example: krbpaths.
          # for more information on pathways, check leapR documentation.
    # 3. "alpha": set the BH-adjusted pvalue of significance.
  #Output:
    # A list of dataframes representing pathway enrichment by sets in each 
    # cluster. Filters for pathways which are overrepresented in each cluster,
    # does not include when a pathway is under-represented relative to other clusters.
    # Each dataframe in the list will be named by cluster number.
leapR_enrich_sets <- function(df, gset, alpha) {
  modules <- as.numeric(df$Cluster)
  names(modules) <- df$feature
  clusters <- lapply(unique(modules),function(x) names(which(modules==x)))
  list_of_dfs <- list()
  for (c in 1:length(clusters)) {
    df_current = leapR(geneset=gset, 
                  enrichment_method="enrichment_in_sets",
                  background=names(modules), 
                  targets=names(modules[which(modules==c)]))
    df_current <- df_current %>% filter(BH_pvalue<alpha) %>% filter(oddsratio>1)
    list_of_dfs[[(length(list_of_dfs)+1)]] <- df_current
  }
  return(list_of_dfs)
}


## Function to count the number of rows per list in a list of lists. Used here
  # to count how many enriched paths are in each cluster.
get_dim <- function(Cluster) {
  #Obtain dim
  output <-as.data.frame(do.call(rbind,lapply(Cluster,function(x) dim(x)[1])))
  #Format
  output$Cluster <- rownames(output)
  output <- output[,c(2,1)]
  names(output)[2]<-'count'
  rownames(output)<-NULL
  return(output)
}

## Function to create a dataframe listing all the unique pathways per cluster
  #Input: "list_of_dfs" : a list of lists, with the enriched pathways as row names
  #Output: a dataframe with a column for each cluster, listing unique pathways
    #ordered by BH-adjusted p-value
unique_paths_info <- function(list_of_dfs) {
  for (c in 1:length(list_of_dfs)) {
    #extract pathway information from each cluster, save as a new dataframe
      #with a "pathways" column
    list_of_dfs[[c]]$pathways <- rownames(list_of_dfs[[c]])
    #df <- list_of_dfs[[c]]
    #df$pathways <- rownames(df)
    #assign(paste("paths", c, sep = ""), df)
    #print(list_of_dfs[[c]]$pathways) #WORKS
  }
  clust <- c(1:length(list_of_dfs))
  for (c in 1:length(list_of_dfs)) {
    #save the list of other clusters minus the current cluster
    non_c <- clust[-c]
    #print(non_c) #WORKS
    for (other_c in non_c) {
      #print(other_c) #WORKS
      # find paths unique to the current cluster
      if (other_c == non_c[1]) {
        # First comparison is between current cluster and the first other cluster
        unique_paths <- setdiff(list_of_dfs[[c]]$pathways, list_of_dfs[[other_c]]$pathways)
        #Remaining comparisons sequentially compare to the updating list of
          # paths unique to the current cluster
      } else {
        unique_paths <- setdiff(unique_paths, list_of_dfs[[other_c]]$pathways)
      }
    }
    #for each cluster, save the unique path information
    unique_paths_df <- data.frame("pathways" = unique_paths)
    #merge the rest of the enrichment info from the original info
      #save by overwriting old list of dfs for each cluster
    list_of_dfs[[c]] <- merge(unique_paths_df, list_of_dfs[[c]], 
                            by = "pathways", all.x = TRUE)
  }
  return(list_of_dfs)
}


#Function to get a multi-column list of unique pathways for multiple clusters
  #Input: a list of dataframes, each dataframe corresponding to a cluster
    # column 1 should be "pathways" (the unique pathways for that cluster)
  #Output: a dataframe with a column for each cluster's unique pathways
list_unique_paths <- function(list_of_dfs) {
  #save the largest number of pathways of any cluster
  largest <- max(sapply(list_of_dfs, nrow))
  for (c_df in 1:length(list_of_dfs)) {
    if(c_df == 1) {
      paths <- data.frame(list_of_dfs[[c_df]][order(list_of_dfs[[c_df]][,"BH_pvalue"]),][1:largest,1])
    } else {
      paths <- cbind(paths, list_of_dfs[[c_df]][order(list_of_dfs[[c_df]][,"BH_pvalue"]),][1:largest,1])
    }
  }
  colnames(paths) <- c(1:ncol(paths))
  return(paths)
}
```

**With Network: Download and process data from folder syn26997954 on Synapse**
Data is already clustered by ElasticNet.
##Still need to change to net by ElasticNet
```{r data prep net}
# With network created from protein, RNA, phosphosites.
  # Networks made with Sara's code using package amlresistancenetworks.
  # Note kinases are predicted from phosphosites, and transcription factors are
  # predicted from transcriptomics. The network also includes proteins imputed
  # based on the presence of others in the network, increasing the number of 
  # features per cluster relative to the no-network feature list.


# Download network files from synapse
graph_k4_enet_c1 <- extract_graph("syn27042867", 1)
graph_k4_enet_c2 <- extract_graph("syn27042881", 2)
graph_k4_enet_c3 <- extract_graph("syn27042882", 3)
graph_k4_enet_c4 <- extract_graph("syn27042883", 4)
graph_k4_enet_c5 <- extract_graph("syn27042884", 5)

#combine into single network
net <- rbind(graph_k4_enet_c1[,], graph_k4_enet_c2[,], graph_k4_enet_c3[,], graph_k4_enet_c4[,], graph_k4_enet_c5[,])
#remove phosphosite hits (the network naturally includes the protein as an
  #imputed feature, so no need to add the protein back for phos features)
k4_enet_net <- net[which(net$data_type != "phosphosite"),1:ncol(net)]
```

**Without network: Download and process data from folder syn29612385**
```{r data prep net}
## No network - only RNA, protein, and proteins in phospho data (without 
  #phosphosite)
raw_noNet <- read.table(synGet('syn30029402')$path,sep='\t',header=T) %>% 
  dplyr::rename(feature = Gene) %>%
  dplyr::select(feature, Cluster, alpha, data_type)

  #separate phospho and non-phospho data
no_net_nonphos <- raw_noNet[which(raw_noNet$data_type != "Phospho"),1:ncol(raw_noNet)]
no_net_phos <- raw_noNet[which(raw_noNet$data_type == "Phospho"),1:ncol(raw_noNet)]
  #remove phosphosite name from phospho data, just leaving protein
no_net_phos$feature <- gsub("(.*)-.*", "\\1", no_net_phos$feature)
  #recombine phos and non-phos data, remove duplicated proteins within clusters
no_net <- rbind(no_net_phos, no_net_nonphos)
k4_enet_noNet <- no_net[!duplicated(no_net[1:2]),]
k4_enet_noNet$Cluster <- as.double(k4_enet_noNet$Cluster)

glimpse(k4_enet_noNet)
glimpse(k4_enet_net)

cat("Without network:", nrow(k4_enet_noNet), "features. \nWith network: ", 
     nrow(k4_enet_net), "features.")
#ggplot(k4_enet_noNet, aes(x = Cluster, y = count(feature))) +
  #geom_bar()

#download necessary databases from LeapR
data("krbpaths")
```


```{r}
#Function copied directly from leapR
cluster_enrichment <- function(geneset, clusters, background=NA, sigfilter=0.05) {
  x = length(clusters)
  
  # The default background is all the genes that are in all the clusters
  # the user can submit their own background list if they want to do something different
  if (is.na(background))  background = sapply(1:x, function (n) unlist(clusters[[n]]))
  
  #this = sapply(1:x, function (i) list(enrichment_in_groups(geneset, clusters[[i]], background)))
  this = sapply(1:x, function (i) list(leapR(geneset=geneset, enrichment_method="enrichment_in_sets", 
                                             targets=clusters[[i]], background=background)))
  
  # if the sigfilter is set we'll only return those functions that have a p-value
  #   lower than the threshold
  if (is.na(sigfilter)) return(this)
  
  outlist = list()
  for (i in 1:x) {
    these = this[[i]]
    these = these[which(these[,"BH_pvalue"]<sigfilter),]
    outlist = c(outlist, list(these))
  }
  return(outlist)
}
```


Do enrichment in sets using krbpaths
```{r enrich in sets net}

  #Without network
modules <- as.numeric(k4_enet_noNet$Cluster)
names(modules) <- k4_enet_noNet$feature
clusters <- lapply(unique(modules),function(x) names(which(modules==x)))

enrich_sets_k4_enet_noNet_krbpaths = cluster_enrichment(geneset=krbpaths, 
                                                clusters=clusters, 
                                                background=names(modules), 
                                                sigfilter=0.05)
  #with network
modules <- k4_enet_net$Cluster
names(modules) <- k4_enet_net$feature
clusters <- lapply(unique(modules),function(x) names(which(modules==x)))

enrich_sets_k4_enet_net_krbpaths = cluster_enrichment(geneset=krbpaths, 
                                                clusters=clusters, 
                                                background=names(modules),
                                                sigfilter = 0.05)


# Run clusters individually to make sure it's accurate.
enrich_sets_k4_enet_net_krbpaths_byIndiv <- leapR_enrich_sets(k4_enet_net, krbpaths, 0.05)
enrich_sets_k4_enet_noNet_krbpaths_byIndiv <- leapR_enrich_sets(k4_enet_noNet, krbpaths, 0.05)

#count the number of enriched paths per dataset
count_noNet <- get_dim(enrich_sets_k4_enet_noNet_krbpaths)
count_net <- get_dim(enrich_sets_k4_enet_net_krbpaths)
count_noNet_i <- get_dim(enrich_sets_k4_enet_noNet_krbpaths_byIndiv)
count_net_i <- get_dim(enrich_sets_k4_enet_net_krbpaths_byIndiv)

#put all data frames into list
df_list <- list(count_noNet, count_net, count_noNet_i, count_net_i)

#merge all data frames in list
numPaths_k4_enet_krb <- df_list %>% reduce(full_join, by='Cluster')
#change column names to suit
colnames(numPaths_k4_enet_krb) <- c("Cluster", "noNet", "Net", 
                            "noNet_byClust", "Net_byClust")

numPaths_k4_enet_krb[,c(1,4,5)]

```



Identify pathways unique to each cluster within each dataset (net vs no net):
```{r}
  #get lists of enriched pathways per cluster, for net vs no net
pathInfo_k4_enet_noNet <- unique_paths_info(enrich_sets_k4_enet_noNet_krbpaths_byIndiv)
paths_k4_enet_noNet <- list_unique_paths(pathInfo_k4_enet_noNet)
pathInfo_k4_enet_net <- unique_paths_info(enrich_sets_k4_enet_net_krbpaths_byIndiv)
paths_k4_enet_net <- list_unique_paths(pathInfo_k4_enet_net)
  #view
paths_k4_enet_noNet
paths_k4_enet_net

```

```{r}
#Side-by-side comparison of pathways identified in cluster 1 (the largest cluster)
  # with and without a network

c1_shared <- intersect(paths_k4_enet_noNet[,1], paths_k4_enet_net[,1])
c1_noNet_unique <- setdiff(paths_k4_enet_noNet[,1], paths_k4_enet_net[,1])
c1_net_unique <- setdiff(paths_k4_enet_net[,1], paths_k4_enet_noNet[,1])
c1_noNet <- data.frame("paths" = c(c1_shared, c1_noNet_unique, rep(NA, length(c1_net_unique) - length(c1_noNet_unique))))
c1_net <- data.frame("paths" = c(c1_shared, c1_net_unique))
c1_compare <- cbind(c1_noNet, c1_net)

c1_compare

```

