---
title: "Deconvolution heatmaps"
author: "Sara Gosline"
date: "2/15/2022"
output: html_document
---


Here we use decomprolute to assess the deconvolution algorithm that best agrees with mRNA based approaches and use that to deconvolve cell type in the aML proteomics data.

```{r login}
library(amlresistancenetworks)
library(dplyr)
library(tidyr)
library(pheatmap)
library(RColorBrewer)

syn = synapseLogin()

#decontab <- syn$tableQuery('select * from syn27085832')$asDataFrame()
clusters <-read.table(syn$get('syn26642544')$path,sep='\t',header=T)

clin <- read.table(syn$get('syn26642974')$path,quote='"',header=T)%>%
  select(labId,FLT3_ITDCall,percentBlastsBM,percentBlastsPB)%>%
  distinct()

estimate <- syn$tableQuery('select * from syn27148720')$asDataFrame()%>%
  dplyr::rename(labId='sampleID')%>%
  dplyr::select(labId,TumorPurity)

##let's read in the AML subtypes as well
aml.res <- read.table('AML-tumor-prot-raw-xcell-AML.tsv',sep='\t',header=T,check.names = F,row.names = 1)

celltypes <- c("Tumor-combined GMP-like","Tumor-combined Myeloid-like",
  "Tumor-celltype HSC-like","Tumor-celltype Progenitor-like",
  "Tumor-celltype Promono-like","Tumor-celltype Monocyte-like",'LSC17') 

new.res <- aml.res%>%
    tibble::rownames_to_column('subType')%>%
    pivot_longer(cols=2:ncol(.),names_to='patient',values_to='cellPop')%>%
  subset(subType%in%celltypes)

lsc17.df<-subset(new.res,subType=='LSC17')%>%
  select(labId='patient',`LSC17 score`='cellPop')

##collect cluster membeership, FLT3 status
pat.annote <- clusters%>%
  dplyr::select(k.4,k.5,k.8)%>%
  mutate(k.4=as.factor(k.4),k.5=as.factor(k.5),k.8=as.factor(k.8))%>%
  tibble::rownames_to_column('labId')%>%full_join(clin)%>%
 # left_join(estimate)%>%
  mutate(FLT3=ifelse(FLT3_ITDCall=='Negative','Negative','Positive'))%>%
  left_join(lsc17.df)%>%
  tibble::column_to_rownames('labId')%>%
     # mutate(`Bone marrow blasts`=as.numeric(percentBlastsBM),
#                `Blood blasts`=as.numeric(percentBlastsPB))%>%
  dplyr::select(`K4 Cluster`=k.4,`K5 Cluster`=k.5,`K8 Cluster`=k.8,`FLT3 ITD`=FLT3,`LSC17 score`)
              #,TumorPurity)

new.res<-new.res%>%
  left_join(tibble::rownames_to_column(pat.annote,'patient'))%>%
  mutate(cellType=gsub('Tumor-','',subType))%>%
  mutate(cellType=gsub('celltype|combined ','',cellType))



k8.colors=  RColorBrewer::brewer.pal(8,'Dark2')
#c("#999999", "#E69F00", "#56B4E9", "#009E73", 
          #             "#F0E442", "#0072B2", "#D55E00", "#CC79A7")#rev(c("#fff7fb","#ece2f0","#d0d1e6","#a6bddb","#67a9cf","#3690c0","#02818a","#016450"))
names(k8.colors)=seq(1,8)
k5.colors=k8.colors[1:5]#c("#fef0d9","#fdcc8a","#fc8d59","#e34a33","#b30000")
names(k5.colors)=seq(1,5)

k4.colors=k8.colors[c(2,3,5,6)]
names(k4.colors)=seq(1,4)

flt3.colors<-c(Positive='#f0f0f0',Negative='#636363')

```


## Standard protein deconvolution

We found that LM7c matrix and xcell are the best in terms of mRNA/prot agreement, so we plot the deconvoltuion of AML samples on this.

```{r do plots, warning=F, message=F}


#algs = unique(decontab$algorithm)
#mats = unique(decontab$matrix)
a = 'xcell'
m = c('LM7c')
dats = c('prot','mrna')

lapply(dats,function(d){
    deconv = paste0('AML-tumor-',d,'-raw-',a,'-',m,'.tsv')
    print(deconv)
    fname = paste0(a,'-',m,'-',d,'.pdf')
    tmat <- read.table(deconv,sep='\t',header=T,row.names=1,check.names=F)
    
    
  ##K4 plot
  ann.colors=list(`K4 Cluster`=k4.colors,`FLT3 ITD`=flt3.colors)#,`Blood
    anns=c('K4 Cluster','FLT3 ITD','LSC17 score')

    pheatmap(tmat,labels_col=rep("",ncol(tmat)),
             annotation_col=pat.annote[,anns],
             annotation_colors=ann.colors,
             cellheight=12,cellwidth=2,
             clustering_distance_cols = 'correlation')

    f2name=paste0('k4-',fname) 

    pheatmap(tmat,filename=f2name,labels_col=rep("",ncol(tmat)),
             annotation_col=pat.annote[,anns],annotation_colors=ann.colors,
             cellheight=12,cellwidth=2,
             clustering_distance_cols = 'correlation',height=8)
    synapseStore(f2name,parentId='syn27092543')
  
    
  ##K5 plot
  ann.colors=list(`K5 Cluster`=k8.colors[1:5],`FLT3 ITD`=flt3.colors)#,`Blood
    anns=c('K5 Cluster','FLT3 ITD','LSC17 score')

    pheatmap(tmat,labels_col=rep("",ncol(tmat)),
             annotation_col=pat.annote[,anns],
             annotation_colors=ann.colors,
             cellheight=12,cellwidth=2,
             clustering_distance_cols = 'correlation')

    f2name=paste0('k5-',fname) 

    pheatmap(tmat,filename=f2name,labels_col=rep("",ncol(tmat)),
             annotation_col=pat.annote[,anns],annotation_colors=ann.colors,
             cellheight=12,cellwidth=2,
             clustering_distance_cols = 'correlation',height=8)
    synapseStore(f2name,parentId='syn27092543')
  
    ##k8 plot
    anns=c('K8 Cluster','FLT3 ITD','LSC17 score')
    ann.colors=list(`K8 Cluster`=k8.colors,`FLT3 ITD`=flt3.colors)#,`Blood blasts`=grey.colors(100),`Bone marrow blasts`=grey.colors(100))

    f2name=paste0('k8-',fname) 

  pheatmap(tmat,filename=f2name,labels_col=rep("",ncol(tmat)),annotation_col=pat.annote[,anns],annotation_colors=ann.colors,cellheight=12,cellwidth=2,clustering_distance_cols = 'correlation',height=8)
    synapseStore(f2name,parentId='syn27092543')
   
}) 
#    })
  
#})

```
The standard cell types do not cluster samples by subtype.

## AML subtypes


Now we want to visualize the AML cell types, which use a different matrix

```{r aml cell types}
ann.colors=list(`K5 Cluster`=k8.colors[1:5],`FLT3 ITD`=flt3.colors)#,`Blood blasts`=grey.colors(100),`Bone marrow blasts`=grey.colors(100))

anns=c('K4 Cluster','FLT3 ITD','LSC17 score')
    ann.colors=list(`K4 Cluster`=k4.colors,`FLT3 ITD`=flt3.colors)#,`Blood

fname = 'xcell-AML.pdf'
a = 'xcell'
m = c('AML')
dats = c('prot','mrna')

lapply(dats,function(d){
    deconv = paste0('AML-tumor-',d,'-raw-',a,'-',m,'.tsv')
    print(deconv)
    fname = paste0(a,'-',m,'-',d,'.pdf')
    tmat <- read.table(deconv,sep='\t',header=T,row.names=1,check.names=F)
    
   sc.res<-tmat[celltypes[1:6],]

pheatmap(sc.res,labels_col=rep("",ncol(sc.res)),
         annotation_col=pat.annote[,anns],annotation_colors=ann.colors,
         cellheight=12,cellwidth=2,clustering_distance_cols = 'correlation')
 f2name=paste0('k4-',fname) 

 pheatmap(sc.res,filename=f2name,labels_col=rep("",ncol(sc.res)),annotation_col=pat.annote[,anns],annotation_colors=ann.colors,cellheight=12,cellwidth=2,clustering_distance_cols = 'correlation',height=8)
    synapseStore(f2name,parentId='syn27092543')
})
```
Here there seem to be two distinct clusters - cluster 2-3 are in the tumor-like, and 1/4/5 are in the normal-like. Perhaps we can see how they are represented in boxplots instead.


```{r aml type box plot}
library(ggplot2)

other<-new.res%>%
  subset(subType!='LSC17')%>%
  separate('subType',into=c('sample','cell'),sep='-')%>%
  mutate(cell=tolower(cell))

ggplot(other,aes(x=cell,y=cellPop,fill=`K4 Cluster`))+
    geom_boxplot()+theme(axis.text.x = element_text(angle = 270, vjust = 0.0))+
    scale_fill_manual(values=ann.colors$`K4 Cluster`)
  scale_color_manual(values=ann.colors$`K4 Cluster`)
ggsave('scSubtypes_by_cluster.pdf')  
synapseStore('scSubtypes_by_cluster.pdf',parentId='syn27092543')

ggplot(other,aes(x=`K4 Cluster`,y=cellPop,fill=cell))+
    geom_boxplot()
ggsave('scSubtypes_all.pdf')
synapseStore('scSubtypes_all.pdf',parentId='syn27092543')

lsc17 <- new.res%>%
  subset(subType=='LSC17')

ggplot(lsc17,aes(x=subType,y=cellPop,fill=`K4 Cluster`))+geom_boxplot()+    scale_fill_manual(values=ann.colors$`K4 Cluster`)
ggsave('lsc17_by_cluster.pdf')
synapseStore('lsc17_by_cluster.pdf',parentId='syn27092543')


ggplot(lsc17,aes(x=`K4 Cluster`,y=cellPop,fill=subType))+geom_boxplot()
ggsave("lsc17_all.pdf")
synapseStore('lsc17_all.pdf',parentId='syn27092543')


```

now what does this mean? THE LSC17 signature is clearly highest in Cluster 2, which makes it the most stem-like, which agrees with the tumor HSC assignment in the single cell data. 




